### 2016-01-21 paper.tex
Add stuff from @weijianzhang

\documentclass[11pt, conference, , compsocconf]{IEEEtran}
\usepackage{tikz}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{bm}
\usepackage[linesnumbered]{algorithm2e}
\makeatletter
\renewcommand{\@algocf@capt@plain}{above}
\makeatother
\usepackage[colorlinks,urlcolor=blue,citecolor=hotpink,linkcolor=blue]{hyperref}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{arrows}
\usetikzlibrary{snakes}
\usetikzlibrary{calc}
\let\emptyset\varnothing
\definecolor{hotpink}{rgb}{0.9,0,0.5}
\definecolor{deepgray}{gray}{0.35}
\definecolor{deepgray2}{gray}{0.25}
\definecolor{deepgray3}{gray}{0.10}
\definecolor{lightgray}{gray}{0.95}
\definecolor{lightgray2}{gray}{0.85}
\definecolor{purple1}{RGB}{239,229,244}
\definecolor{purple2}{RGB}{216,191,216}
\definecolor{lightblue}{rgb}{0.73,0.33,0.83}
\definecolor{lightpurple}{rgb}{.8,.2,.8}
\definecolor{textcolor}{rgb}{0,0,5}
\definecolor{blue1}{RGB}{187,217,238}
\definecolor{blue2}{RGB}{235,244,250}
\definecolor{yellow1}{RGB}{255,249,234}
\definecolor{blue3}{RGB}{63,40,96}
\definecolor{red1}{RGB}{255, 102, 102}
\definecolor{green1}{RGB}{102, 255, 102} 
\newtheorem{lemma}{Lemma}
\newtheorem{thm}{Theorem}
\theoremstyle{definition}
\def\proof{\par{\bf Proof}. \ignorespaces}
\def\qedsymbol{\vbox{\hrule\hbox{%
                     \vrule height1.3ex\hskip0.8ex\vrule}\hrule}}
\def\endproof{\qquad\qedsymbol\medskip\par}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi

\usepackage{url}
% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{An Algorithm for Searching Evolving Graphs in Linear Algebraic Language}

% author names and affiliations
% use a multiple column layout for up to two different
% affiliations
% \and

\author{
\IEEEauthorblockN{Jiahao Chen}
\IEEEauthorblockA{ Computer Science and Artificial Intelligence Laboratory,\\
Massachusetts Institute of Technology,\\
Massachusetts, 02139-4307,USA\\
jiahao@mit.edu} 
\and
\IEEEauthorblockN{Weijian Zhang}
\IEEEauthorblockA{School of Mathematics,\\
University of Manchester,\\
Manchester, M13 9PL, England\\
weijian.zhang@manchester.ac.uk}
}

\maketitle

\begin{abstract}
  % A generalization of the breadth-first search (BFS) algorithm on evolving graphs is
  % discussed. 
  % The representation of this algorithm 
  % as algebraic operations is investigated.
  % The duality of the vector-matrix multiplication 
  % and a BFS step is extended to the fact that a block vector-matrix multiplication
  % is dual with the evolving-graph BFS. 
  % It is found that the algebraic representation allows for flexible forward and backward searching and can be used for mining citation networks.
A generalized breadth-first search (BFS) algorithm is presented for 
searching evolving graphs. 
The method is described in both graph and linear algebra notation and 
the two algorithms are shown to be equivalent. 
A preliminary discussion on mining citation networks is also presented.
\end{abstract}

\begin{IEEEkeywords}
~Evolving graph; complex network; breadth first search; data mining.
\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
%\begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}

Let's imagine a game played by three people $A$, $B$ and $C$.
Suppose we give a piece of information $X$ to $A$ and ask 
$A$ talks to $B$ first without $C$'s presence and then ask $B$ talks to $C$ without $A$'s presence.
In a different day, we give a piece of information $Y$ to $A$ again.
But this time, we ask $B$ talks to $C$ first and 
then ask $A$ talks to $B$ without $C$'s presence. 
We now ask the question ``Will $C$ find out what $X$ is ? Will $C$ find out
what $Y$ is?''

Evolving graphs are models to study this kind of scenario where the 
interrelations between entities are time-dependent.
% We assume that the reader is familiar with the basic graph terminology. 
% All the necessary material can be found in}.
We follow the notation of (static) graphs in  \cite{even12} and \cite{kegi11}.
In general, evolving graphs $G(t) = (V(t), E(t))$ are graphs that change as a function of time $t$. However, in different papers the term ``evolving graph'' has been
used to refer to slightly different models. 
 Flajolet et al. \cite{fkp89} use the term to mean the evolution of random graphs.
Bahmani et al. \cite{bkmu12} use the term to mean a sequence of graphs where
the difference between two consecutive graphs is small.
Similar models are also investigated using the term ``evolving network'' and 
``time-dependent graph''.

In this work, we consider a discretization of the timeline as $n$ distinct 
timestamps. We define an evolving graph $G_n$ to be 
a sequence of (static) graphs 
$G_n = \langle G^{[1]}, G^{[2]}, \ldots, G^{[n]} \rangle$. 
Depends on applications, the graph $G^{[i]}$ could mean either
a snapshot of $G(t)$ at timestamp $i$ or  the union of all the graphs
between timestamp $i-1$ and $i$ (excludes timestamp $i-1$).
Both the node set and edge set of $G^{[i]}$ could be different from
these of $G^{[j]}, j \ne i$. 
\begin{center}
\begin{tikzpicture}[snake=zigzag, line before snake = 5mm, line after snake = 5mm]
    % draw horizontal line   
    \draw (0,0) -- (2,0);
    \draw[snake] (2,0) -- (4,0);
    \draw (4,0) -- (5,0);
    \draw[snake] (5,0) -- (7,0);

    % draw vertical lines
    \foreach \x in {0,1,2,4,5,7}
      \draw (\x cm,3pt) -- (\x cm,-3pt);

    % draw nodes
    \draw (0,0) node[below=3pt] {$ 1 $} node[above=3pt] {$   $};
    \draw (1,0) node[below=3pt] {$ 2 $} node[above=3pt] {$ G^{[2]} $};
    \draw (2,0) node[below=3pt] {$ 3 $} node[above=3pt] {$  $};
    \draw (3,0) node[below=3pt] {$  $} node[above=3pt] {$  $};
    \draw (4,0) node[below=3pt] {$ 5 $} node[above=3pt] {$  $};
    \draw (5,0) node[below=3pt] {$ 6 $} node[above=3pt] {$ G^{[6]}$};
    \draw (6,0) node[below=3pt] {$  $} node[above=3pt] {$  $};
    \draw (7,0) node[below=3pt] {$ n $} node[above=3pt] {$ $};
  \end{tikzpicture}
\end{center}
Our model is the same as that of Tang et al. \cite{tmml10} though they
use slightly different notation.
Grindrod et al.'s model  \cite{gphe11} \cite{grihig13} is very similar to ours, 
where an evolving graph is considered as a sequence of adjacency matrices.
Our model can also be seen as a special case of 
the multilayer networks discussed by Kivel{\"a} et al. \cite{kabg14}, where 
there is only one set of layers for time.
% We present a preliminary discussion of using the model to mine citation networks.

% why BFS is important? 
Breadth-first search is probably the most basic graph searching algorithm and 
it is also the building block of other graph algorithms like minimum spanning tree 
and shortest path \cite{ckrs09}.
Given a graph and a source node $s$, 
the breadth-first search (BFS) systematically discover every node that is reachable from $s$. 
As the name suggested, BFS expands the frontier of discovered regions
uniformly across the breadth, i.e., the algorithm discovers all nodes
that are $k$ walks away from $s$ before discovering any nodes that are
$k+1$ walks away.

A temporal path is the basic element to 
 define almost  all important evolving graph algorithms.
Matrix-matrix multiplication between consecutive adjacency matrices
were commonly used as a means of determining temporal paths \cite{gphe11} \cite{grihig13}. However, we show
in one specific case it fails to include all the temporal paths and thus should not be used 
for searching evolving graphs in general.
Despite the fact that graph metrics \cite{ntmm13} \cite{tmml09} \cite{tsmm09} and 
various centrality measures (including Katz  \cite{gphe11} \cite{grihig13}, 
betweenness \cite{alhi15}, and closeness  \cite{tmml10}) 
have been generalized for evolving graphs,
a BFS for evolving graphs, which we shall call it \emph{evolving-graph BFS},
was never explicitly described or verified.

In this paper,  we present descriptions of an evolving-graph BFS algorithm, but its main concern is with its sparse matrix
representation and linear algebraic computations. 
We prove that the two algorithms (graph notation and algebra notation) are equivalent.

% GraphBlas related ideas

This paper is organized as follows. The evolving-graph BFS algorithm is described
and analyzed using graph notation in Section \ref{sec:breadth-first-search}. Section \ref{sec:evolv-graph-algor} considers the algebraic representation of the algorithm: 
A sparse matrix representation
of an evolving graph is presented in \ref{sec:representation}. 
The algorithm is presented in \ref{sec:breadth-first-search-1} and 
the properties of the search tree is discussed in \ref{sec:short-temp-path}.
The connection between a temporal path and a matrix-matrix multiplication is considered and discussed in \ref{sec:matr-matr-mult}. Finally, 
an application of the evolving-graph BFS in mining citation networks is described in Section \ref{sec:applications}.

\section{Breadth-First Search for Evolving Graphs}
\label{sec:breadth-first-search}

\subsection{Evolving-Graph BFS}
\label{sec:evolving-graph-bfs}

\begin{figure}[h]
 \begin{center}
    \begin{tikzpicture}[scale=.38, line width =1.5pt]
  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {2};
  \node[circle,draw=red1, minimum size=0.2cm] (n8) at (-10,5) {3};
  \node[circle,fill=green1, minimum size=0.2cm] (n10) at (-8,10) {1};

  \node[circle,draw=red1, minimum size=0.2cm] (n6) at (3,7) {2};
  \node[circle,fill=green1, minimum size=0.2cm] (n4) at (-2,5) {3};
  \node[circle,fill=green1, minimum size=0.2cm] (n1) at (0,10) {1};

   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {2};
  \node[circle,fill=green1, minimum size=0.2cm] (n12) at (6,5)  {3};
  \node[circle,draw=red1, minimum size=0.2cm] (n14) at (8,10) {1};

  \foreach \from/\to in {n10/n7, n1/n4, n11/n12}
   \draw[every edge,->] (\from) -- (\to);
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{ $t_2$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $t_1$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $t_3$};
    \end{tikzpicture}
\end{center}
\caption{An evolving graph with 3 timestamps $t_1$, $t_2$ and $t_3$.
At each timestamp, the evolving graph is represented as a graph. 
The green filled nodes in the figure are active nodes.}
\label{fig:eg_shortest_path}
\end{figure}

We show a small evolving graph $ \langle G^{[1]}, G^{[2]}, G^{[3]}\rangle$ with $3$ timestamps $t_1$, $t_2$ and $t_3$ 
in Figure \ref{fig:eg_shortest_path}, where each $G^{[i]}$ is a graph at timestamp $t_i$. 
For example,
at timestamp $t_1$, the graph $G^{[1]}$ contains three nodes $1$, $2$ and $3$ and $(1,2)$ is an edge of $G^{[1]}$. 
Though Figure \ref{fig:eg_shortest_path} presents a 
directed evolving graph,  the algorithms that we shall discuss work on both 
directed and undirected evolving graphs.
We shall use this example throughout the paper to illustrate various
notions.

Formally, an evolving graph $G_n $ is defined to be a sequence of graphs $\langle G^{[1]}, G^{[2]}, \ldots, G^{[n]}\rangle$,  where each $G^{[t]} = (V^{[t]}, E^{[t]})$ represents a graph at timestamp $t$. 
We shall use the notation $(v, t)$ to represent
a node $v$ at timestamp $t$. 
We say two nodes are \emph{connected} if there is an edge between them
and we say a node $(v,t)$ is \emph{active}, if $v$ is connected to at least a node other than itself at timestamp $t$. Otherwise we say it is \emph{inactive}.  
In Figure \ref{fig:eg_shortest_path}, node $(1,t_1)$ and $(2,t_2)$  are active but node $(3,t_1)$ is inactive.

In order to define the evolving-graph BFS, we need the concept of a 
\emph{temporal path}.
A temporal path on an evolving graph $G_n$ is a sequence of node-timestamp
pair $\langle (v_1, t_1), (v_2, t_2), \ldots, (v_m, t_m) \rangle$ such that 
$t_1 \leq t_2 \leq \cdots \leq t_m$ and $v_i = v_j$ if and only if 
$t_i \ne t_j$. The length of a temporal path is the length of the sequence.
For example, $\langle (1, t_1), (1, t_2), (3, t_2), (3, t_3) \rangle$ and 
$\langle (1, t_1), (2, t_1), (2, t_3), (3, t_3) \rangle$ 
are two temporal paths from $(1, t_1)$ to $(3, t_3)$
on the evolving graph in Figure \ref{fig:eg_shortest_path}.

We would like a BFS on evolving graphs to work similar to that on graphs: the search starts with some source node $(s,t)$, called \emph{root}. It constructs a tree through discovering all nodes that are reachable by 
temporal paths of length $k$ before moving to all nodes that are reachable by 
 temporal paths of length $k+1$.
By construction, the timestamp of an active node is always not later than 
that of its descendant. Suppose we 
start with node $(1, t_1)$ in Figure \ref{fig:eg_shortest_path}. 
After $1$ temporal path from $(1, t_1)$, we can reach 
$(2, t_1)$ and $(1, t_2)$. After $2$ temporal paths, we can reach 
$(2, t_1), (1, t_2), (2, t_2)$ and $(3, t_2)$. 
% This example motivates us to define the reachable set of 
% a node-timestamp pair.
Formally, we define the \emph{forward-neighbors} of a node $(v, t)$ to
be the set of all the reachable nodes after $1$ temporal path from
$(v, t)$, i.e., the nodes pointed by $v$ at timestamp $t$ and 
the active node $v$ itself at later timestamps. In Figure 
\ref{fig:eg_shortest_path}, 
the forward-neighbors of $(1, t_1)$  are $(2, t_1)$ and $(1, t_2)$ and 
the forward-neighbors of $(2, t_1)$ is $(2, t_3)$.
See Algorithm  \ref{alg:bfs} for the evolving-graph BFS algorithm.

\begin{algorithm}[h]
 \SetKwFor{For}{for}{}{end}
 \SetKwFor{While}{while}{}{end}
 \SetKwFor{If}{if}{}{end}
 \SetKwFunction{nei}{forward-neighbors}
 \SetKwFunction{enqueue}{enqueue}
 \SetKwProg{Fn}{function}{}{end}
 \Fn{BFS($G_n, (s,t)$)}{
  $dict.(s,t) = 0$ \\
  $i = 1$ \\
  $fronter = \emptyset$ \\
  \enqueue{$frontier, (s,t)$} \\
 \While{$frontier \ne \emptyset$}{
   $next = \emptyset$ \\
   \For{$(u,t') \in frontier$}{
     \For{$(v,t'') \in$\nei{$(u,t')$}}{
          \If{$(v,t'') \notin d$}{
              $dict.(v,t'') = i$\\
              \enqueue{$next, (v,t'')$}
              }
         }
       }
  $frontier = next$ \\
  $i = i+1$ \\
  }
  \textbf{return} $dict$
}
 \caption{Evolving-graph BFS. 
Let $G_n$ be an
evolving graph represented by adjacency lists and $(s, t)$ is a node 
of $G_n$.
For each node-timestamp pair $(v_i,t_i)$,
we use a dictionary $dict.(v_i,t_i)$ to store its distance (the length of temporal
paths) from $(s, t)$. We also use
two first-in, first-out queue $frontier$ and $next$, where $frontier$
stores the frontier nodes of the current discovered regions and $next$ stores
the frontier nodes of the next level. 
The algorithm returns $dict$ which stores all reachable node-timestamp pairs from $(s,t)$ such that if $dict.(v_i, t_i) = k$, 
the length of the shortest temporal path
from $(s,t)$ to $(v_i, t_i)$ is $k$.}
 \label{alg:bfs}
\end{algorithm}

\subsection{Analysis}
\label{sec:analysis}

For the graph $G^{[i]} = (V^{[i]}, E^{[i]})$ at timestamp $i$, 
we denote the active node set of $V^{[i]}$ as $\tilde V^{[i]}$, 
where we have $\tilde V^{[i]} \subseteq V^{[i]}$.
Step $10$ of  Algorithm \ref{alg:bfs} ensures each active node-timestamp pair in $G_n$ is enqueued in $next$ (and then $frontier$) at most once. 
And so \texttt{forward-neighbors} loops through each 
active node-timestamp pair at most once. 
The operations of enqueuing take $O(1)$ time.
Thus the total operations of enqueuing are 
$O(\sum_{i=1}^n|\tilde V^{[i]}|)$, where $|\cdot|$ represents the number of elements of a set.  Let $\sum_i |\tilde V^{[i]}| = |\tilde V|$ and $\sum_i |E^{[i]}| = |\hat E|$.
Then the time of scanning
\texttt{forward-neighbors} is 
\[
\sum_{i=1}^n\sum_{v \in \tilde V^{[i]}} \texttt{forward-neighbors}(v) = 
O(|\hat E|),
\]
by the definition of forward-neighbors.
Hence the total running time is $O(|\hat E|+ |\tilde V|)$.

Note that the number of adjacency lists is the same as the number
of active nodes in $G_n$. So for large $n$, this representation can use a large amount of memory.
%Also need to consider the operations for matrix lists.

\section{Traversing Evolving Graphs Using Linear Algebraic Operations}
\label{sec:evolv-graph-algor}

\subsection{Representation}
\label{sec:representation}
As before, let $G_n = \langle G^{[1]}, G^{[2]}, \ldots, G^{[n]}\rangle$ be an evolving graph with $n$ timestamps. For each graph $G^{[t]}$, we use $A^{[t]}$ to denote
the corresponding adjacency matrix.
\[
A^{[t]}(i,j) = 
\begin{cases}
 1 & \mbox{if $(i,j)$ is an edge of $G^{[t]}$,} \\
 0 & \mbox{otherwise.}
\end{cases}
\]
 Each $A^{[t]}$ can be represented by 
a $|V|\times |V|$ sparse matrix,
where $|V|$ is the total number of nodes in $G_n$.
We use a sequence of adjacency matrices $A_n = \langle A^{[1]}, A^{[2]}, \ldots, A^{n]}\rangle$ to represent $G_n$. 
For example, the evolving graph in Figure \ref{fig:eg_shortest_path}
can be represented as 
\[
\Big\langle
  \begin{bmatrix} 
    0 & 1 & 0 \\
    0 & 0 & 0 \\
    0 & 0 & 0
  \end{bmatrix},~
 \begin{bmatrix}
   0 & 0 & 1 \\
   0 & 0 & 0 \\
   0 & 0 & 0 
 \end{bmatrix},~
 \begin{bmatrix}
  0 & 0 & 0 \\
  0 & 0 & 1 \\
  0 & 0 & 0 
 \end{bmatrix}
\Big\rangle
\]
We assume $A_n$ is stored as an array of sparse matrices in compressed sparse column (CSC) format.

\subsection{Forward and Backward Searching}
\label{sec:breadth-first-search-1}

There exits a duality between vector matrix multiplication and the BFS
for graphs \cite{kegi11}. Let $G = (V, E)$ be a (directed or undirected) graph
and $k$ a node in the graph $G$. We could represent $G$ as an adjacency 
matrix $A$ and $k$ a column vector $b$ such that 
\[
b_i = 
\begin{cases}
1 & \mbox{if $i = k$,} \\
0 & \mbox{otherwise.}
\end{cases}
\]
Then $A^Tb$ is a column vector where the indices of the nonzero entries 
indicate the forward-neighbors of $k$. This relationship can be generalized to be
a duality between a block vector matrix multiplication and the 
evolving-graph BFS.

For the evolving case, we shall introduce a new binary operator $\odot$.
We define
\[
A \odot b = 
\begin{cases}
b & \mbox{if $Ab \ne 0$,} \\
0 & \mbox{otherwise,}
\end{cases}
\]
where $A$ is an adjacency matrix and $b$ a column vector defined as before.
Then it is easy to observe that 
the forward-neighbors of node $(k, 1)$ in $A_n = \langle A^{[1]}, A^{[2]}, \ldots, A^{n]}\rangle$ are the nonzero indices of the 
elements in the sequence
\begin{equation}
\label{eq:out_ne}
\big\langle (A^{[1]})^Tb, \quad  (A^{[2]})^T\odot b, \quad \ldots \quad (A^{[n]})^T\odot b \big\rangle.
\end{equation}
The forward-neighbors at timestamp $t$ are the nonzero indices of the $t$-th element of the above sequence.
Hence given $A_n$ and a vector $b$, we can use \eqref{eq:out_ne} to 
get a $1$-step  BFS tree. In Figure \ref{fig:eg_shortest_path}, the forward-neighbors
of node $(1, t_1)$ can be computed by
\[
\Big\langle
  \begin{bmatrix} 
    0 & 0 & 0 \\
    1 & 0 & 0 \\
    0 & 0 & 0
  \end{bmatrix} 
\begin{bmatrix}
1 \\
0 \\
0
\end{bmatrix}
,
 \begin{bmatrix}
   0 & 0 & 0 \\
   0 & 0 & 0 \\
   1 & 0 & 0 
 \end{bmatrix}
\odot
\begin{bmatrix}
1 \\
0 \\
0
\end{bmatrix}
,
 \begin{bmatrix}
  0 & 0 & 0 \\
  0 & 0 & 0 \\
  0 & 1 & 0 
 \end{bmatrix}
\odot
\begin{bmatrix}
1 \\
0 \\
0
\end{bmatrix}
\Big\rangle
\]
which is equal to 
\[
\Big\langle
\begin{bmatrix}
0 \\
1 \\
0
\end{bmatrix},~
\begin{bmatrix}
1 \\
0 \\
0
\end{bmatrix},~
\begin{bmatrix}
0 \\
0 \\
0
\end{bmatrix}
\Big\rangle
\]
This means node $(2,t_1)$ and $(1,t_2)$ are the forward-neighbors of $(1,t_1)$.

Let $(A^{[i]})^Tb = c_i$.
At the second step, we need to compute the following components
\begin{align}
  \label{eq:5}
 \big\langle (A^{[1]})^Tc_1,   (A^{[2]})^T\odot c_1, \ldots,  (A^{[n]})^T\odot c_1  \big\rangle  \\  
\big\langle (A^{[2]})^Tc_{2}, \ldots, (A^{[n]})^T\odot c_{2} \big\rangle \\
 \ldots  \\
  \big\langle (A^{[n]})^Tc_n \big\rangle 
\label{eq:6}
\end{align}
If we sum these resulting vectors according to timestamps, we obtain 
the forward-neighbors of the nodes computed at step $1$.
For $n =2$, we have 
\[
\Big\langle
(A^{[1]})^Tc_1,~
(A^{[2]})^T\odot c_1 + (A^{[2]})^T c_2
\Big\rangle.
\]
If we form a block triangular matrix $\bm{A_n}^T$ 
\[
\bm{A_n}^T = 
\begin{bmatrix}
(A^{[1]})^T &   0            & \cdots  &  & 0 \\
\sigma(A^{[2]})^T & (A^{[2]})^T & 0         &\cdots  & \vdots \\
\sigma(A^{[3]})^T & \sigma(A^{[3]})^T & (A^{[3]})^T & \cdots \\
   \vdots              &                              &      \vdots            &  &  \vdots \\
\sigma(A^{[n]})^T &                  &       \cdots                      &  & (A^{[n]})^T 
\end{bmatrix},
\]
where  $\sigma(A^{[i]})^Tb = (A^{[i]})^T\odot b$, and a block vector $\bm{b}$ 
\[
\bm{b}^T = [b^T, 0, \cdots, 0  ],
\]
then \eqref{eq:out_ne} is equivalent to $\bm{A_n}^T\bm{b}$ and 
summing \eqref{eq:5}-\eqref{eq:6} is equivalent to $\bm{A_n}^T(\bm{A_n}^T\bm{b})$. 
In general, we can define an evolving-graph BFS using 
block vector-matrix multiplications. The algorithm is presented in Algorithm
\ref{alg:bfsla}.

\begin{algorithm}[h]
 \SetKwFor{For}{for}{}{end}
 \SetKwFor{While}{while}{}{end}
 \SetKwFor{If}{if}{}{end}
 \SetKwFunction{nei}{forward-neighbors}
 \SetKwFunction{enqueue}{enqueue}
 \SetKwFunction{nz}{nonzeros}
 \SetKwFunction{intersect}{intersect}
 \SetKwFunction{map}{map}
 \SetKwProg{Fn}{function}{}{end}
 \Fn{ABFS($A_n, (s,t)$)}{
   Form $\bm{A}^T$ from $A_n$\\
   $m = A^{[1]}.dimension$ \\ 
   $b$ is a sparse vector of length $mn$ \\
   $b(s+mt) =1$ \\
   $i = 1$ \\
   $dict.(s,t) = 0$ \\
   \While{\nz{$b$} $\ne \emptyset$}{
       $\bm{b} = \bm{A}^T\bm{b}$ \\
       \For{$i$ in \nz{$b$}}{
         $node =$ \map{$i$} \\
         \If{$node$ in $dict$}{
           $b(i) = 0$ \\
         }
      }
      $nodes =$\map{$b$}\\
       \For{$node$ in $nodes$}{
           $dict.node = i$
         }

       $i = i + 1$
       }
  \textbf{return} $dict$ \\
}
 \caption{Algebraic evolving-graph BFS. 
Let $A_n$ be the corresponding adjacency matrices of $G_n$ and 
$(s,t)$ a node of $G_n$. As in Algorithm \ref{alg:bfs}, we use a dictionary $dict.(v_i, t_i)$
to store $(v_i, t_i)$'s distance from $(s,t)$. Function \texttt{nonzeros(v)} 
returns the nonzero indices of a given vector  $v$ and function \texttt{map(b)} 
maps a block vector's indices to their corresponding nodes. The algorithm  returns 
$dict$ which stores all the reachable active nodes and their distance from $(s,t)$.}
 \label{alg:bfsla}
\end{algorithm}

Representing evolving graphs as sparse matrices offers great flexibility 
on the search direction. If we define the \emph{backward-neighbors} of a 
node $(v,t)$ to be a set of nodes that precede $(v,t)$ if we travel backwards
in time. Then the backward-neighbors of $(v,t)$ can be computed by
\begin{equation}
\label{eq:backward}
(A^{[t]})b + (A^{[t+1]})\odot b + \cdots + (A^{[n]})\odot b.
\end{equation}
We shall get a backward traversal algorithm if we replace
the vector-matrix multiplication $\bm{A_n}^T\bm{b}$ 
by $\bm{A_n}\bm{b}$, where 
\[
\bm{A_n} = 
\begin{bmatrix}
A^{[1]} &   0            & \cdots  &  & 0 \\
\sigma(A^{[2]}) & A^{[2]} & 0         &\cdots  & \vdots \\
\sigma(A^{[3]}) & \sigma(A^{[3]}) & A^{[3]} & \cdots \\
   \vdots              &                              &      \vdots            &  &  \vdots \\
\sigma(A^{[n]}) &                  &       \cdots                      &  & A^{[n]} 
\end{bmatrix}.
\]
Having the ability to traverse backwards is important for many real-life applications. 
We shall show an example in Section \ref{sec:applications}.

Algorithm \ref{alg:bfsla} outlines the basic idea of an algebraic evolving-graph BFS.
But in practice we don't need to form $\bm{A}$ and $\bm{b}$ explicitly at step $4$ and $6$. By realizing the dependency between these sparse matrices, 
we can compute  the result in parallel.

\begin{thm}[Correctness of evolving-graph BFS]
Let an evolving graph $G_n$ and a starting node $(s,t_0)$ be given. 
The evolving-graph BFS discovers every active node that is reachable from the source $(s,t_0)$ and $d.(v, t)$ is the length of the shortest temporal path 
from $(s,t_0)$ to $(v,t)$. 
\end{thm}
\proof
We could map an evolving graph $G_n$ to a static graph $G$ in such a way that
if there is an edge from $(v, t)$ to $(u,t)$ in $G_n$, there is an edge from $v_t$ to $u_t$ in $G$.
For any two nodes $v_s$ and $v_t$ in $G$,  if $s < t$, we add an edge from $v_s$ to $v_t$. 
By construction, the forward-neighbors of an active node $(v, t)$ in $G_n$ 
are in one-to-one correspondence with the forward-neighbors of a node $v_t$ in $G$.  
Hence the correctness of the evolving graph BFS follows from that of a static graph.
\endproof

\begin{thm}
 Algorithm \ref{alg:bfs} and Algorithm \ref{alg:bfsla} are equivalent. 
\end{thm}
\proof
We shall check the two algorithms are equivalent in each module of 
the method.
The initialization steps are trivially equivalent.
By the construction of $\bm{A^T}$, 
step $9$ of Algorithm \ref{alg:bfsla} computes the forward-neighbors 
of all the frontier nodes, which is equivalent to step $8$-$9$ 
of Algorithm \ref{alg:bfs}. Step $10$-$15$ of Algorithm \ref{alg:bfsla} ensures
active nodes will not be visited twice, which is equivalent to step $10$
of Algorithm \ref{alg:bfs}. Finally, a dictionary $dict$ is used in both algorithms 
to keep record of the distance of each node from the source $(s,t)$.
\endproof

\subsection{Search Tree}
\label{sec:short-temp-path}

We say two nodes $(v_1,t_1)$ and $(v2, t_2)$ are \emph{temporally connected} if 
$(v_1, t_1)$ and $(v2,t_2)$ are connected by a temporal path. 
Notice this notion is symmetric, reflexive but not transitive. For example, suppose 
$t_1 < t_2 < t_3$ and $(v_1, t_1)$ and $(v_3, t_3)$ are temporally connected 
by a temporal path from $(v_1,t_1)$ to $(v_3, t_3)$. Also
$(v_3, t_3)$ and $(v_2, t_2)$ are temporally connected by a temporal path from
$(v_2,t_2)$ to $(v_3, t_3)$. But this does not imply $(v_1, t_1)$ and $(v_2, t_2)$ are temporal connected.

Given a root node $(s, t)$, during the execution the BFS discussed above builds a search tree, denoted by $T(s,t)$, 
where nodes are temporally connected. 
\begin{lemma}
 For a node active at different timestamps, say $(s, t_1)$ and $(s,t_2)$ such that $t_1 < t_2$,  we have $T(s, t_2) \subset T(s, t_1)$.
\end{lemma}
\proof
Suppose $(v, t)$ belongs to $T(s, t_2)$. Then there exits an temporal path
(constructed by the evolving graph BFS) from $(s,t_2)$ to $(v,t)$. Since both
$(s,t_1)$ and $(s,t_2)$ are active, $(s,t_2)$ is an out-neighbor of $(s,t_1)$ by definition, i.e., there exits an edge from $(s,t_1)$ to $(s,t_2)$.  Hence there exits an temporal path
from $(s,t_1)$ to $(v,t)$ and so $(v,t) \in T(s,t_1)$.
\endproof

% consider $(A+I)^Tb$, which records all the connected components.
% consider the huge matrix block and how to relate it.

% \subsection{Online Searching}
% \label{sec:online-searching}

% In this section, we consider ways of  updating new graphs. Suppose $A_n $ is 
% an evolving graph with $n$ timestamps. We received a new graph $A^{[n+1]}$
% at timestamp $n+1$. How to update the tree components while minimize the computation?
% The forward-neighbors at timestamp $$

\subsection{Matrix-Matrix Multiplication}
\label{sec:matr-matr-mult}

For adjacency matrices $A^{[s]}$ and $A^{[t]}$, where $s < t$, a nonzero entry
$(i,j)$ of $A^{[s]} A^{[t]}$ indicates a temporal path from $(i,s)$ to $(j,t)$ via
some node $k$. 
In general, a nonzero entry $(i,j)$ of the matrix $A^{[1]}A^{[2]}\cdots A^{[n]}$ 
indicates a temporal path $p$ that starts at $(i,1)$ and ends at  $(j,n)$. 
Notice $p$ must go through at least one edge at each timestamp between $1$ and $n$.
For the evolving graph in Figure \ref{fig:eg_shortest_path}, $A^{[1]}A^{[2]}$
is a zero matrix since there is no temporal path from timestamp $t_1$ to $t_2$ that
go through at least one edge at $t_1$. However, 
$\langle(1, t_1), (1, t_2), (3,t_2)\rangle$ is a temporal path of Figure \ref{fig:eg_shortest_path}.

% But it provides no information about the nodes in between. In addition, the temporal path must pass at least 
%  one node at each timestamps. We believe this requirement is too restrictive. For example,
% $A^{[t_1]}A^{[t_2]}A^{[t_3]}$ is a zero matrix. Let $\langle i,j,\ldots k \rangle$ represents the sets of all possible subsets of $\langle 1, 2, \ldots, n \rangle$, where $i<j<\cdots<k$, then 
% the nonzero entries of the matrix
% \[
% \sum_{i,j,\ldots,k}A^{[1]}A^{[i]}A^{[j]}\cdots A^{[k]}A^{[n]},
% \]
% indicates temporal paths  from temporal paths from timestamp $1$ to $n$.

\section{Application in Citation Networks}
\label{sec:applications}

We shall map a database of research papers to an evolving graph $G_n$
in such a way that if author $i$ cited author $j$ on a paper published
at  time $t$, the evolving graph $G_n$ contains a directed edge from
$i$ to $j$ at timestamp $t$. Then given an author $a$ at timestamp $t_0$,
we can use our evolving-graph BFS to construct $T(a, t_0)$ which is 
a set of all the authors that have been influenced by $a$'s work at  
time $t_0$. We define a \emph{community} to be a group of 
researchers that share the same information source. 
For example, given a paper published by $a$ at year $2016$, 
we can determine $a$'s community by first backward searching to 
find the source then forward searching to find the descendents.

To design such system, we shall use two parameters: 
 a parameter for 
determining the layers of the backward search tree which can be chosen by the user;
a second parameter for determining the number of roots as a means of 
filtering out unessential or overlapping trees. The subset of roots can be determined
by ranking algorithms like PageRank or machine-learning based preference ranking.

\section{Conclusion}

In this paper, we have introduced and analyzed the evolving-graph BFS in both 
graph and algebraic notation. 
In order to study evolving graphs and experiment with various graph types, we have developed EvolvingGraphs.jl \cite{zhang15}, a software package in the Julia language \cite{bkse12} for the creation, manipulation, and study of dynamic networks.

This paper leaves important extensions open for further research. First, it is 
of interest to extend the evolving-graph BFS algorithm for weighted evolving graphs, which can provide more valuable information when the degree of connectivity is available in the data. 
Second, efficient algebraic representations and algorithms need to be addressed. 
One major advantage of representing evolving graphs as sparse matrices
is the matrix representation is highly structured so it is easy to design
high performance parallel implementations. 
Third, depth-first search and other graph traversal algorithms can be 
extended for searching evolving graphs using similar ideas.

% % use section* for acknowledgement
% \section*{Acknowledgment}

% The authors would like to thank...
% more thanks here

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)

\bibliographystyle{myplain2-doi}

\bibliography{strings,njhigham,paper}

% that's all folks
\end{document}

### 2016-01-21 paper.tex
reorganize the paper

@@
 
-This paper is organized as follows. The evolving-graph BFS algorithm is described
-and analyzed using graph notation in Section \ref{sec:breadth-first-search}. Section \ref{sec:evolv-graph-algor} considers the algebraic representation of the algorithm: 
-A sparse matrix representation
-of an evolving graph is presented in \ref{sec:representation}. 
-The algorithm is presented in \ref{sec:breadth-first-search-1} and 
-the properties of the search tree is discussed in \ref{sec:short-temp-path}.
-The connection between a temporal path and a matrix-matrix multiplication is considered and discussed in \ref{sec:matr-matr-mult}. Finally, 
-an application of the evolving-graph BFS in mining citation networks is described in Section \ref{sec:applications}.
+% This paper is organized as follows. The evolving-graph BFS algorithm is described
+% and analyzed using graph notation in Section \ref{sec:breadth-first-search}. Section \ref{sec:evolv-graph-algor} considers the algebraic representation of the algorithm: 
+% A sparse matrix representation
+% of an evolving graph is presented in \ref{sec:representation}. 
+% The algorithm is presented in \ref{sec:breadth-first-search-1} and 
+% the properties of the search tree is discussed in \ref{sec:short-temp-path}.
+% The connection between a temporal path and a matrix-matrix multiplication is considered and discussed in \ref{sec:matr-matr-mult}. Finally, 
+% an application of the evolving-graph BFS in mining citation networks is described in Section \ref{sec:applications}.
 
@@
 
-\subsection{Analysis}
-\label{sec:analysis}
-
-For the graph $G^{[i]} = (V^{[i]}, E^{[i]})$ at timestamp $i$, 
-we denote the active node set of $V^{[i]}$ as $\tilde V^{[i]}$, 
-where we have $\tilde V^{[i]} \subseteq V^{[i]}$.
-Step $10$ of  Algorithm \ref{alg:bfs} ensures each active node-timestamp pair in $G_n$ is enqueued in $next$ (and then $frontier$) at most once. 
-And so \texttt{forward-neighbors} loops through each 
-active node-timestamp pair at most once. 
-The operations of enqueuing take $O(1)$ time.
-Thus the total operations of enqueuing are 
-$O(\sum_{i=1}^n|\tilde V^{[i]}|)$, where $|\cdot|$ represents the number of elements of a set.  Let $\sum_i |\tilde V^{[i]}| = |\tilde V|$ and $\sum_i |E^{[i]}| = |\hat E|$.
-Then the time of scanning
-\texttt{forward-neighbors} is 
-\[
-\sum_{i=1}^n\sum_{v \in \tilde V^{[i]}} \texttt{forward-neighbors}(v) = 
-O(|\hat E|),
-\]
-by the definition of forward-neighbors.
-Hence the total running time is $O(|\hat E|+ |\tilde V|)$.
-
-Note that the number of adjacency lists is the same as the number
-of active nodes in $G_n$. So for large $n$, this representation can use a large amount of memory.
 %Also need to consider the operations for matrix lists.
 
-\section{Traversing Evolving Graphs Using Linear Algebraic Operations}
-\label{sec:evolv-graph-algor}
-
-\subsection{Representation}
+\subsection{Algebraic Evolving-Graph BFS}
 \label{sec:representation}
@@
 \]
-We assume $A_n$ is stored as an array of sparse matrices in compressed sparse column (CSC) format.
-
-\subsection{Forward and Backward Searching}
-\label{sec:breadth-first-search-1}
 
@@
 
+\subsection{Analysis}
+\label{sec:analysis}
+
+% adjacency matrices
+Suppose $G_n$ is represented by adjacency lists.
+Then the number of adjacency lists is the same as the number
+of active nodes in $G_n$. So for large $n$, this representation 
+can use a large amount of memory.
+For the graph $G^{[i]} = (V^{[i]}, E^{[i]})$ at timestamp $i$, 
+we denote the active node set of $V^{[i]}$ as $\tilde V^{[i]}$, 
+where we have $\tilde V^{[i]} \subseteq V^{[i]}$.
+Step $10$ of  Algorithm \ref{alg:bfs} ensures each active node-timestamp pair in $G_n$ is enqueued in $next$ (and then $frontier$) at most once. 
+And so \texttt{forward-neighbors} loops through each 
+active node-timestamp pair at most once. 
+The operations of enqueuing take $O(1)$ time.
+Thus the total operations of enqueuing are 
+$O(\sum_{i=1}^n|\tilde V^{[i]}|)$, where $|\cdot|$ represents the number of elements of a set.  Let $\sum_i |\tilde V^{[i]}| = |\tilde V|$ and $\sum_i |E^{[i]}| = |\hat E|$.
+Then the time of scanning
+\texttt{forward-neighbors} is 
+\[
+\sum_{i=1}^n\sum_{v \in \tilde V^{[i]}} \texttt{forward-neighbors}(v) = 
+O(|\hat E|),
+\]
+by the definition of forward-neighbors.
+Hence the total running time is $O(|\hat E|+ |\tilde V|)$.
+
 \begin{thm}[Correctness of evolving-graph BFS]
@@
 
+\section{Node Activeness}
+\label{sec:matr-matr-mult}
+
+\subsection{Temporal Path}
+\label{sec:temporal-path}
+
+In this subsection, we compare with the ``wrong'' way of traveling an evolving graph
+and show the importance of node activeness.
+
 \subsection{Matrix-Matrix Multiplication}
-\label{sec:matr-matr-mult}
+\label{sec:matr-matr-mult-1}
 
@@
 % indicates temporal paths  from temporal paths from timestamp $1$ to $n$.
+
+\section{Implementation in Julia}
+\label{sec:implementation-julia}
+
+In order to study evolving graphs and experiment with various graph types, we have developed EvolvingGraphs.jl \cite{zhang15}, a software package in the Julia language \cite{bkse12} for the creation, manipulation, and study of dynamic networks.
+
+In Julia,  $A_n$ is stored as an array of sparse matrices in compressed sparse column (CSC) format.
 
@@
 graph and algebraic notation. 
-In order to study evolving graphs and experiment with various graph types, we have developed EvolvingGraphs.jl \cite{zhang15}, a software package in the Julia language \cite{bkse12} for the creation, manipulation, and study of dynamic networks.
-
 This paper leaves important extensions open for further research. First, it is 

### 2016-01-21 paper.tex
ignore auto/ and update julia implementation part

@@
 In order to study evolving graphs and experiment with various graph types, we have developed EvolvingGraphs.jl  \cite{zhang15}, a software package in the Julia language \cite{bkse12} for the creation, manipulation, and study of dynamic networks.
-
-In Julia,  $A_n$ is stored as an array of sparse matrices in compressed sparse column (CSC) format.
+Julia is a high-level, high-performance dynamic programming language for technical
+computing. When we implement EvolvingGraphs.jl, we make particular use of 
+the parameterizable type system and multiple dispatch in Julia. 
+We define a data type  \texttt{MatrixList}, which represents an evolving graph
+as a sequence of sparse matrices stored in compressed sparse column (CSC) format. 
+Many basic graph functions are defined for \texttt{MatrixList}, including 
+retrieving nodes, timestamps, and matrices.
 

### 2016-01-21 paper.tex
update introduction

@@
 
-Let's imagine a game played by three people $A$, $B$ and $C$.
-Suppose we give a piece of information $X$ to $A$ and ask 
-$A$ talks to $B$ first without $C$'s presence and then ask $B$ talks to $C$ without $A$'s presence.
-In a different day, we give a piece of information $Y$ to $A$ again.
-But this time, we ask $B$ talks to $C$ first and 
-then ask $A$ talks to $B$ without $C$'s presence. 
-We now ask the question ``Will $C$ find out what $X$ is ? Will $C$ find out
-what $Y$ is?''
+Let's imagine a game played by three people
+$A$, $B$, and $C$. Suppose each of them has a piece of information,
+say $X$, $Y$, and $Z$ respectively. The goal of the game is 
+to get information from other people as much as possible. 
+Suppose $A$ talks to $B$ first and then $B$ talks to $C$, 
+then $C$ may has all three pieces of information $X$, $Y$, and $Z$ even 
+if $C$ has never talked to $A$. 
+However, if $B$ talks to $C$ first and then $A$ talks to $B$, then
+no matter how smart $C$ is, $C$ can never get information 
+$X$.  Time is an important factor for information flow.
 
-Evolving graphs are models to study this kind of scenario where the 
-interrelations between entities are time-dependent.
+We can analyze the spread of information between $A$, $B$, and $C$
+as a process on a graph that changes over time.  
+Such graphs are known as evolving graphs.
+% Evolving graphs are models to study this kind of scenario where the 
+% interrelations between entities are time-dependent.
 % We assume that the reader is familiar with the basic graph terminology. 
@@
 We follow the notation of (static) graphs in  \cite{even12} and \cite{kegi11}.
-In general, evolving graphs $G(t) = (V(t), E(t))$ are graphs that change as a function of time $t$. However, in different papers the term ``evolving graph'' has been
+In general, a evolving graph $G(t) = (V(t), E(t))$ is  a graph where the node set
+and edge set are functions of time $t$.
+However, in different literatures the term ``evolving graph'' has been
 used to refer to slightly different models. 
@@
 In this work, we consider a discretization of the timeline as $n$ distinct 
-timestamps. We define an evolving graph $G_n$ to be 
+timestamps and define an evolving graph $G_n$ to be 
 a sequence of (static) graphs 
@@
 % why BFS is important?
-Breadth-first search is probably the most basic graph searching algorithm and 
-it is also the building block of other graph algorithms like minimum spanning tree 
-and shortest path \cite{ckrs09}.
-Given a graph and a source node $s$, 
-the breadth-first search (BFS) systematically discover every node that is reachable from $s$. 
-As the name suggested, BFS expands the frontier of discovered regions
-uniformly across the breadth, i.e., the algorithm discovers all nodes
-that are $k$ walks away from $s$ before discovering any nodes that are
-$k+1$ walks away.
 
+% Given a graph and a source node $s$, 
+% the breadth-first search (BFS) systematically discover every node that is reachable from $s$. 
+% As the name suggested, BFS expands the frontier of discovered regions
+% uniformly across the breadth, i.e., the algorithm discovers all nodes
+% that are $k$ walks away from $s$ before discovering any nodes that are
+% $k+1$ walks away. 
 A temporal path is the basic element to 
@@
 for searching evolving graphs in general.
+
+Breadth-first search is probably the most basic graph searching algorithm and 
+it is also the building block of other graph algorithms like minimum spanning tree 
+and shortest path \cite{ckrs09}.
 Despite the fact that graph metrics \cite{ntmm13} \cite{tmml09} \cite{tsmm09} and 
@@
 
-\section{Node Activeness}
+\section{Node Activeness and its Impact}
 \label{sec:matr-matr-mult}
+
+\subsection{Aggregated Graphs}
+\label{sec:aggregated-graphs}
+
+In this section, we answer the question why we need evolving graph.
+
+At the beginning of the paper, we set up a game played by three people $A$, $B$,
+and $C$ and we showed if
 
@@
 $\langle(1, t_1), (1, t_2), (3,t_2)\rangle$ is a temporal path of Figure \ref{fig:eg_shortest_path}.
+
+Introducing ones on the diagonal of the matrices will still get the wrong result.
 
@@
 
-\section{Application in Citation Networks}
+\section{Application to Citation Networks}
 \label{sec:applications}

### 2016-01-21 paper.tex
Minor cleanup of citations
Remove windows crap

@@
-\documentclass[11pt, conference, , compsocconf]{IEEEtran}
+\documentclass[11pt,conference,compsocconf]{IEEEtran}
 \usepackage{tikz}
@@
 % All the necessary material can be found in}.
-We follow the notation of (static) graphs in  \cite{even12} and \cite{kegi11}.
+We follow the notation of (static) graphs in \cite{even12,kegi11}.
 In general, a evolving graph $G(t) = (V(t), E(t))$ is  a graph where the node set
@@
 Matrix-matrix multiplication between consecutive adjacency matrices
-were commonly used as a means of determining temporal paths \cite{gphe11} \cite{grihig13}. However, we show
+were commonly used as a means of determining temporal paths \cite{gphe11,grihig13}. However, we show
 in one specific case it fails to include all the temporal paths and thus should not be used
@@
 and shortest path \cite{ckrs09}.
-Despite the fact that graph metrics \cite{ntmm13} \cite{tmml09} \cite{tsmm09} and 
-various centrality measures (including Katz  \cite{gphe11} \cite{grihig13}, 
+Despite the fact that graph metrics \cite{ntmm13,tmml09,tsmm09} and
+various centrality measures like Katz centrality~\cite{gphe11,grihig13},
 betweenness \cite{alhi15}, and closeness \cite{tmml10})

### 2016-01-21 paper.tex
Rewrite description of algebraic formulation

@@
 
-There exits a duality between vector matrix multiplication and the BFS
-for graphs \cite{kegi11}. Let $G = (V, E)$ be a (directed or undirected) graph
-and $k$ a node in the graph $G$. We could represent $G$ as an adjacency
-matrix $A$ and $k$ a column vector $b$ such that
+This algebraic formulation exploits a graphical interpretation of matrix--vector
+products involving the adjacency matrix~\cite{kegi11}.
+Let $G = (V, E)$ be a (directed or undirected) graph
+and $k$ be a node in the graph $G$. If $A$ is the adjacency matrix of $G$ and
+$b$ is the column vector
 \[
@@
 1 & \mbox{if $i = k$,} \\
-0 & \mbox{otherwise.}
+0 & \mbox{otherwise,}
 \end{cases}
 \]
-Then $A^Tb$ is a column vector where the indices of the nonzero entries
-indicate the forward-neighbors of $k$. This relationship can be generalized to be
-a duality between a block vector matrix multiplication and the
-evolving-graph BFS.
+then $A^Tb$ is a column vector where the indices of the nonzero entries
+indicate the forward-neighbors of $k$.
 
-For the evolving case, we shall introduce a new binary operator $\odot$.
-We define
+The evolving graph BFS can be thought of as a generalization of blocked
+matrix--vector products.
+Introduce a new operator $\odot$, a variation on the ordinary matrix--vector
+product such that
 \[
@@
 b & \mbox{if $Ab \ne 0$,} \\
-0 & \mbox{otherwise,}
+0 & \mbox{otherwise.}
 \end{cases}
 \]
-where $A$ is an adjacency matrix and $b$ a column vector defined as before.
-Then it is easy to observe that
-the forward-neighbors of node $(k, 1)$ in $A_n = \langle A^{[1]}, A^{[2]}, \ldots, A^{n]}\rangle$ are the nonzero indices of the
+Then, the forward-neighbors of node $(k, 1)$ in $A_n = \langle A^{[1]}, A^{[2]}, \ldots, A^{n]}\rangle$ are the nonzero indices of the
 elements in the sequence

### 2016-01-21 paper.tex
Run spell check

@@
 and edge set are functions of time $t$.
-However, in different literatures the term ``evolving graph'' has been
+However, in different literature the term ``evolving graph'' has been
 used to refer to slightly different models.
@@
 
-In this work, we consider a discretization of the timeline as $n$ distinct
-timestamps and define an evolving graph $G_n$ to be
+In this work, we consider a discretization of the time line as $n$ distinct
+time stamps and define an evolving graph $G_n$ to be
 a sequence of (static) graphs
@@
 Depends on applications, the graph $G^{[i]}$ could mean either
-a snapshot of $G(t)$ at timestamp $i$ or  the union of all the graphs
-between timestamp $i-1$ and $i$ (excludes timestamp $i-1$).
+a snapshot of $G(t)$ at time stamp $i$ or  the union of all the graphs
+between time stamp $i-1$ and $i$ (excludes time stamp $i-1$).
 Both the node set and edge set of $G^{[i]}$ could be different from
@@
 Our model can also be seen as a special case of
-the multilayer networks discussed by Kivel{\"a} et al. \cite{kabg14}, where
+the multi-layer networks discussed by Kivel{\"a} et al. \cite{kabg14}, where
 there is only one set of layers for time.
@@
 \end{center}
-\caption{An evolving graph with 3 timestamps $t_1$, $t_2$ and $t_3$.
-At each timestamp, the evolving graph is represented as a graph.
+\caption{An evolving graph with 3 time stamps $t_1$, $t_2$ and $t_3$.
+At each time stamp, the evolving graph is represented as a graph.
 The green filled nodes in the figure are active nodes.}
@@
 
-We show a small evolving graph $ \langle G^{[1]}, G^{[2]}, G^{[3]}\rangle$ with $3$ timestamps $t_1$, $t_2$ and $t_3$
-in Figure \ref{fig:eg_shortest_path}, where each $G^{[i]}$ is a graph at timestamp $t_i$.
+We show a small evolving graph $ \langle G^{[1]}, G^{[2]}, G^{[3]}\rangle$ with $3$ time stamps $t_1$, $t_2$ and $t_3$
+in Figure \ref{fig:eg_shortest_path}, where each $G^{[i]}$ is a graph at time stamp $t_i$.
 For example,
-at timestamp $t_1$, the graph $G^{[1]}$ contains three nodes $1$, $2$ and $3$ and $(1,2)$ is an edge of $G^{[1]}$.
+at time stamp $t_1$, the graph $G^{[1]}$ contains three nodes $1$, $2$ and $3$ and $(1,2)$ is an edge of $G^{[1]}$.
 Though Figure \ref{fig:eg_shortest_path} presents a
@@
 
-Formally, an evolving graph $G_n $ is defined to be a sequence of graphs $\langle G^{[1]}, G^{[2]}, \ldots, G^{[n]}\rangle$,  where each $G^{[t]} = (V^{[t]}, E^{[t]})$ represents a graph at timestamp $t$.
+Formally, an evolving graph $G_n $ is defined to be a sequence of graphs $\langle G^{[1]}, G^{[2]}, \ldots, G^{[n]}\rangle$,  where each $G^{[t]} = (V^{[t]}, E^{[t]})$ represents a graph at time stamp $t$.
 We shall use the notation $(v, t)$ to represent
-a node $v$ at timestamp $t$.
+a node $v$ at time stamp $t$.
 We say two nodes are \emph{connected} if there is an edge between them
-and we say a node $(v,t)$ is \emph{active}, if $v$ is connected to at least a node other than itself at timestamp $t$. Otherwise we say it is \emph{inactive}.
+and we say a node $(v,t)$ is \emph{active}, if $v$ is connected to at least a node other than itself at time stamp $t$. Otherwise we say it is \emph{inactive}.
 In Figure \ref{fig:eg_shortest_path}, node $(1,t_1)$ and $(2,t_2)$  are active but node $(3,t_1)$ is inactive.
@@
 \emph{temporal path}.
-A temporal path on an evolving graph $G_n$ is a sequence of node-timestamp
+A temporal path on an evolving graph $G_n$ is a sequence of node-time stamp
 pair $\langle (v_1, t_1), (v_2, t_2), \ldots, (v_m, t_m) \rangle$ such that
@@
  temporal paths of length $k+1$.
-By construction, the timestamp of an active node is always not later than
+By construction, the time stamp of an active node is always not later than
 that of its descendant. Suppose we
@@
 be the set of all the reachable nodes after $1$ temporal path from
-$(v, t)$, i.e., the nodes pointed by $v$ at timestamp $t$ and
-the active node $v$ itself at later timestamps. In Figure
+$(v, t)$, i.e., the nodes pointed by $v$ at time stamp $t$ and
+the active node $v$ itself at later time stamps. In Figure
 \ref{fig:eg_shortest_path},
@@
 of $G_n$.
-For each node-timestamp pair $(v_i,t_i)$,
+For each node-time stamp pair $(v_i,t_i)$,
 we use a dictionary $dict.(v_i,t_i)$ to store its distance (the length of temporal
@@
 the frontier nodes of the next level.
-The algorithm returns $dict$ which stores all reachable node-timestamp pairs from $(s,t)$ such that if $dict.(v_i, t_i) = k$,
+The algorithm returns $dict$ which stores all reachable node-time stamp pairs from $(s,t)$ such that if $dict.(v_i, t_i) = k$,
 the length of the shortest temporal path
@@
 \label{sec:representation}
-As before, let $G_n = \langle G^{[1]}, G^{[2]}, \ldots, G^{[n]}\rangle$ be an evolving graph with $n$ timestamps. For each graph $G^{[t]}$, we use $A^{[t]}$ to denote
+As before, let $G_n = \langle G^{[1]}, G^{[2]}, \ldots, G^{[n]}\rangle$ be an evolving graph with $n$ time stamps. For each graph $G^{[t]}$, we use $A^{[t]}$ to denote
 the corresponding adjacency matrix.
@@
 \]
-then $A^Tb$ is a column vector where the indices of the nonzero entries
-indicate the forward-neighbors of $k$.
+then $A^Tb$ is a column vector whose nonzero entries have indices that are the
+forward-neighbors of $k$.
 
@@
 \end{equation}
-The forward-neighbors at timestamp $t$ are the nonzero indices of the $t$-th element of the above sequence.
+The forward-neighbors at time stamp $t$ are the nonzero indices of the $t$-th element of the above sequence.
 Hence given $A_n$ and a vector $b$, we can use \eqref{eq:out_ne} to
@@
 \end{align}
-If we sum these resulting vectors according to timestamps, we obtain
+If we sum these resulting vectors according to time stamps, we obtain
 the forward-neighbors of the nodes computed at step $1$.
@@
 can use a large amount of memory.
-For the graph $G^{[i]} = (V^{[i]}, E^{[i]})$ at timestamp $i$,
+For the graph $G^{[i]} = (V^{[i]}, E^{[i]})$ at time stamp $i$,
 we denote the active node set of $V^{[i]}$ as $\tilde V^{[i]}$,
 where we have $\tilde V^{[i]} \subseteq V^{[i]}$.
-Step $10$ of  Algorithm \ref{alg:bfs} ensures each active node-timestamp pair in $G_n$ is enqueued in $next$ (and then $frontier$) at most once.
+Step $10$ of  Algorithm \ref{alg:bfs} ensures each active node-time stamp pair in $G_n$ is enqueued in $next$ (and then $frontier$) at most once.
 And so \texttt{forward-neighbors} loops through each
-active node-timestamp pair at most once.
+active node-time stamp pair at most once.
 The operations of enqueuing take $O(1)$ time.
@@
 \begin{lemma}
- For a node active at different timestamps, say $(s, t_1)$ and $(s,t_2)$ such that $t_1 < t_2$,  we have $T(s, t_2) \subset T(s, t_1)$.
+ For a node active at different time stamps, say $(s, t_1)$ and $(s,t_2)$ such that $t_1 < t_2$,  we have $T(s, t_2) \subset T(s, t_1)$.
 \end{lemma}
@@
 indicates a temporal path $p$ that starts at $(i,1)$ and ends at  $(j,n)$.
-Notice $p$ must go through at least one edge at each timestamp between $1$ and $n$.
+Notice $p$ must go through at least one edge at each time stamp between $1$ and $n$.
 For the evolving graph in Figure \ref{fig:eg_shortest_path}, $A^{[1]}A^{[2]}$
-is a zero matrix since there is no temporal path from timestamp $t_1$ to $t_2$ that
+is a zero matrix since there is no temporal path from time stamp $t_1$ to $t_2$ that
 go through at least one edge at $t_1$. However,
@@
 Many basic graph functions are defined for \texttt{MatrixList}, including
-retrieving nodes, timestamps, and matrices.
+retrieving nodes, time stamps, and matrices.
 
@@
 at  time $t$, the evolving graph $G_n$ contains a directed edge from
-$i$ to $j$ at timestamp $t$. Then given an author $a$ at timestamp $t_0$,
+$i$ to $j$ at time stamp $t$. Then given an author $a$ at time stamp $t_0$,
 we can use our evolving-graph BFS to construct $T(a, t_0)$ which is
@@
 we can determine $a$'s community by first backward searching to
-find the source then forward searching to find the descendents.
+find the source then forward searching to find the descendants.
 

### 2016-01-22 paper.tex
Fix formatting of the algebraic example
Also put in A^T in the definition of ocirc
Fix typo: exits -> exists

@@
 \[
-A \odot b =
+A^T \odot b =
 \begin{cases}
-b & \mbox{if $Ab \ne 0$,} \\
+b & \mbox{if $A^Tb \ne 0$,} \\
 0 & \mbox{otherwise.}
@@
 of node $(1, t_1)$ can be computed by
-\[
-\Big\langle
+
+\begin{align*}
+\left\langle
   \begin{bmatrix}
@@
     0 & 0 & 0
-  \end{bmatrix}
+  \end{bmatrix}\!\!\!\!
 \begin{bmatrix}
@@
    1 & 0 & 0
- \end{bmatrix}
-\odot
+ \end{bmatrix}\!\!
+\odot\!\!
 \begin{bmatrix}
@@
   0 & 1 & 0
- \end{bmatrix}
-\odot
+ \end{bmatrix}\!\!
+\odot\!\!
 \begin{bmatrix}
@@
 \end{bmatrix}
-\Big\rangle
-\]
-which is equal to
-\[
-\Big\langle
+\right\rangle
+\\
+=
+\left\langle
 \begin{bmatrix}
@@
 \end{bmatrix}
-\Big\rangle
-\]
+\right\rangle
+\end{align*}
 This means node $(2,t_1)$ and $(1,t_2)$ are the forward-neighbors of $(1,t_1)$.
@@
 \begin{lemma}
- For a node active at different time stamps, say $(s, t_1)$ and $(s,t_2)$ such that $t_1 < t_2$,  we have $T(s, t_2) \subset T(s, t_1)$.
+For a node active at different time stamps, say $(s, t_1)$ and $(s,t_2)$ such that $t_1 < t_2$,  we have $T(s, t_2) \subseteq T(s, t_1)$.
 \end{lemma}
 \proof
-Suppose $(v, t)$ belongs to $T(s, t_2)$. Then there exits an temporal path
+Suppose $(v, t)$ belongs to $T(s, t_2)$. Then there exists an temporal path
 (constructed by the evolving graph BFS) from $(s,t_2)$ to $(v,t)$. Since both
-$(s,t_1)$ and $(s,t_2)$ are active, $(s,t_2)$ is an out-neighbor of $(s,t_1)$ by definition, i.e., there exits an edge from $(s,t_1)$ to $(s,t_2)$.  Hence there exits an temporal path
+$(s,t_1)$ and $(s,t_2)$ are active, $(s,t_2)$ is an out-neighbor of $(s,t_1)$ by definition, i.e., there exists an edge from $(s,t_1)$ to $(s,t_2)$.  Hence there exists a temporal path
 from $(s,t_1)$ to $(v,t)$ and so $(v,t) \in T(s,t_1)$.

### 2016-01-22 paper.tex
Fix braces
add temporal path figure

@@
 We follow the notation of (static) graphs in \cite{even12,kegi11}.
-In general, a evolving graph $G(t) = (V(t), E(t))$ is  a graph where the node set
+In general, an evolving graph $G(t) = (V(t), E(t))$ is  a graph where the node set
 and edge set are functions of time $t$.
@@
 \[
-\Big\langle
+\left\langle
   \begin{bmatrix}
@@
  \end{bmatrix}
-\Big\rangle
+\right\rangle
 \]
@@
 
-\subsection{Aggregated Graphs}
-\label{sec:aggregated-graphs}
+% \subsection{Aggregated Graphs}
+% \label{sec:aggregated-graphs}
 
-In this section, we answer the question why we need evolving graph.
-
-At the beginning of the paper, we set up a game played by three people $A$, $B$,
-and $C$ and we showed if
+% % In this section, we answer the question why we need evolving graph.
+% % 
+% At the beginning of the paper, we set up a game played by three people $A$, $B$,
+% and $C$ and we showed if $A$ and $B$
 
@@
 \label{sec:temporal-path}
+
+\begin{figure}[h]
+ \begin{center}
+    \begin{tikzpicture}[scale=.36, line width =1.4pt]
+  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {2};
+  \node[circle,draw=red1, minimum size=0.2cm] (n8) at (-10,5) {3};
+  \node[circle,fill=green1, minimum size=0.2cm] (n10) at (-8,10) {1};
+
+  \node[circle,draw=red1, minimum size=0.2cm] (n6) at (3,7) {2};
+  \node[circle,fill=green1, minimum size=0.2cm] (n4) at (-2,5) {3};
+  \node[circle,fill=green1, minimum size=0.2cm] (n1) at (0,10) {1};
+
+   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {2};
+  \node[circle,fill=green1, minimum size=0.2cm] (n12) at (6,5)  {3};
+  \node[circle,draw=red1, minimum size=0.2cm] (n14) at (8,10) {1};
+
+  \foreach \from/\to in {n10/n7, n11/n12}
+   \draw[every edge,->] (\from) -- (\to);
+     \draw[dotted,->](n1) -- (n4);
+  \draw[dotted,->](n10) -- (n1);
+  \draw[dotted,->](n4) -- (n12);
+\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
+(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{ $t_2$};
+\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
+(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $t_1$};
+\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
+(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $t_3$};
+    \end{tikzpicture}
+   \begin{tikzpicture}[scale=.36, line width =1.4pt]
+  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {2};
+  \node[circle,draw=red1, minimum size=0.2cm] (n8) at (-10,5) {3};
+  \node[circle,fill=green1, minimum size=0.2cm] (n10) at (-8,10) {1};
+
+  \node[circle,draw=red1, minimum size=0.2cm] (n6) at (3,7) {2};
+  \node[circle,fill=green1, minimum size=0.2cm] (n4) at (-2,5) {3};
+  \node[circle,fill=green1, minimum size=0.2cm] (n1) at (0,10) {1};
+
+   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {2};
+  \node[circle,fill=green1, minimum size=0.2cm] (n12) at (6,5)  {3};
+  \node[circle,draw=red1, minimum size=0.2cm] (n14) at (8,10) {1};
+
+  \foreach \from/\to in {n1/n4}
+   \draw[every edge,->] (\from) -- (\to);
+     \draw[dotted,->](n7) to[out=70, in=40] (n11);
+  \draw[dotted,->](n10) -- (n7);
+   \draw[dotted,->](n11) -- (n12);
+\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
+(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{ $t_2$};
+\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
+(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $t_1$};
+\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
+(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $t_3$};
+    \end{tikzpicture}
+\end{center}
+\caption{An evolving graph with 3 time stamps $t_1$, $t_2$ and $t_3$.
+The back dashed lines indicate a temporal path from $(1, t_1)$ to $(3,t_3)$.}
+\label{fig:active}
+\end{figure}
 

### 2016-01-22 paper.tex
update lemma 1

@@
 \begin{lemma}
-For a node active at different time stamps, say $(s, t_1)$ and $(s,t_2)$ such that $t_1 < t_2$,  we have $T(s, t_2) \subseteq T(s, t_1)$.
+For a node active at different time stamps, say $(s, t_1)$ and $(s,t_2)$ such that $t_1 \leq t_2$,  we have $T(s, t_2) \subseteq T(s, t_1)$, where $T(s,t_1) = T(s,t_2)$ if and only if 
+$t_1 = t_2$.
 \end{lemma}
 \proof
-Suppose $(v, t)$ belongs to $T(s, t_2)$. Then there exists an temporal path
+Trivially, $t_1 =t_2$ if and only if $(s,t_1) = (s,t_2)$ if and only if 
+$T(s,t_1) = T(s,t_2)$. Now suppose $t_1 < t_2$.
+Let $(v, t)$ be an arbitrary active node that belongs to $T(s, t_2)$. Then there exists an temporal path
 (constructed by the evolving graph BFS) from $(s,t_2)$ to $(v,t)$. Since both
@@
 \caption{An evolving graph with 3 time stamps $t_1$, $t_2$ and $t_3$.
-The back dashed lines indicate a temporal path from $(1, t_1)$ to $(3,t_3)$.}
+The back dashed lines of the two figures indicate two temporal paths from $(1, t_1)$ to $(3,t_3)$.}
 \label{fig:active}
 \end{figure}
+
+The concept of node activeness is important for temporal path on evolving graphs.
+The
 

### 2016-01-22 paper.tex
out-neighbor -> forward-neighbor

@@
 $T(s,t_1) = T(s,t_2)$. Now suppose $t_1 < t_2$.
-Let $(v, t)$ be an arbitrary active node that belongs to $T(s, t_2)$. Then there exists an temporal path
+Let $(v, t)$ be an arbitrary active node that belongs to $T(s, t_2)$. Then there exists a temporal path
 (constructed by the evolving graph BFS) from $(s,t_2)$ to $(v,t)$. Since both
-$(s,t_1)$ and $(s,t_2)$ are active, $(s,t_2)$ is an out-neighbor of $(s,t_1)$ by definition, i.e., there exists an edge from $(s,t_1)$ to $(s,t_2)$.  Hence there exists a temporal path
+$(s,t_1)$ and $(s,t_2)$ are active, $(s,t_2)$ is a forward-neighbor of $(s,t_1)$ by definition, i.e., there exists an edge from $(s,t_1)$ to $(s,t_2)$.  Hence there exists a temporal path
 from $(s,t_1)$ to $(v,t)$ and so $(v,t) \in T(s,t_1)$.

### 2016-01-22 paper.tex
update node activeness section

@@
 
-\subsection{Temporal Path}
-\label{sec:temporal-path}
-
 \begin{figure}[h]
@@
 
-The concept of node activeness is important for temporal path on evolving graphs.
-The
-
-In this subsection, we compare with the ``wrong'' way of traveling an evolving graph
-and show the importance of node activeness.
-
-\subsection{Matrix-Matrix Multiplication}
-\label{sec:matr-matr-mult-1}
+Node activeness is the key concept for defining temporal paths which 
+shapes how to traverse an evolving graph. 
+Recall an active node $(v,t)$ is a node $v$ at timestamp $t$
+that connects to at least another node other that itself.   
+It means the entity at that specific timestamp either
+receives recommendations or gives recommendations to others (or possibly both).
+This implies the entity is ``alive''. Here ``alive'' does not necessary mean physically 
+alive. For the case of citation network that we shall discuss later, node activeness more or less means spiritually alive in the sense that the idea or knowledge of the entity is alive.
 
@@
 $(i,j)$ of $A^{[s]} A^{[t]}$ indicates a temporal path from $(i,s)$ to $(j,t)$ via
-some node $k$.
+some node $k$ (which is unknown).
 In general, a nonzero entry $(i,j)$ of the matrix $A^{[1]}A^{[2]}\cdots A^{[n]}$
@@
 Notice $p$ must go through at least one edge at each time stamp between $1$ and $n$.
-For the evolving graph in Figure \ref{fig:eg_shortest_path}, $A^{[1]}A^{[2]}$
+For the evolving graph in Figure \ref{fig:active}, $A^{[1]}A^{[2]}$
 is a zero matrix since there is no temporal path from time stamp $t_1$ to $t_2$ that
 go through at least one edge at $t_1$. However,
-$\langle(1, t_1), (1, t_2), (3,t_2)\rangle$ is a temporal path of Figure \ref{fig:eg_shortest_path}.
-
-Introducing ones on the diagonal of the matrices will still get the wrong result.
+$\langle(1, t_1), (1, t_2), (3,t_2)\rangle$ is a temporal path of Figure \ref{fig:active}.
+Notice also if we introduce ones on the diagonal of the matrices $\{A^{[i]}\}$,
+each node of the resulting evolving graph will become 
+active at all timestamp. Hence matrix-matrix multiplication between consecutive
+adjacency matrices should be avoid  for computing temporal paths.
 

### 2016-01-22 paper.tex
add more details to node activeness section

@@
 alive. For the case of citation network that we shall discuss later, node activeness more or less means spiritually alive in the sense that the idea or knowledge of the entity is alive.
+We traverse an evolving graph only on the active nodes.
 
@@
 indicates a temporal path $p$ that starts at $(i,1)$ and ends at  $(j,n)$.
-Notice $p$ must go through at least one edge at each time stamp between $1$ and $n$.
-For the evolving graph in Figure \ref{fig:active}, $A^{[1]}A^{[2]}$
+Notice the fact that $p$ must go through at least one edge at each time stamp between $1$ and $n$.
+For the evolving graph in Figure \ref{fig:eg_shortest_path}, $A^{[1]}A^{[2]}$
 is a zero matrix since there is no temporal path from time stamp $t_1$ to $t_2$ that
 go through at least one edge at $t_1$. However,
-$\langle(1, t_1), (1, t_2), (3,t_2)\rangle$ is a temporal path of Figure \ref{fig:active}.
-Notice also if we introduce ones on the diagonal of the matrices $\{A^{[i]}\}$,
+$\langle(1, t_1), (1, t_2), (3,t_2)\rangle$ is a temporal path as shown in Figure \ref{fig:active}.
+Notice also that if we introduce ones on the diagonal of the matrices $\{A^{[i]}\}$,
 each node of the resulting evolving graph will become 
-active at all timestamp. Hence matrix-matrix multiplication between consecutive
+active at all timestamps. 
+Hence matrix-matrix multiplication between consecutive
 adjacency matrices should be avoid  for computing temporal paths.

### 2016-01-22 paper.tex
update the introduction part

@@
 
-Breadth-first search is probably the most basic graph searching algorithm and
+Breadth-first search (BFS) is probably the most basic graph searching algorithm and
 it is also the building block of other graph algorithms like minimum spanning tree
@@
 was never explicitly described or verified.
+Therefore, there is a need for clear presentations of the algorithm.
 
@@
 Let an evolving graph $G_n$ and a starting node $(s,t_0)$ be given.
-The evolving-graph BFS discovers every active node that is reachable from the source $(s,t_0)$ and $d.(v, t)$ is the length of the shortest temporal path
+The evolving-graph BFS discovers every active node that is reachable from the source $(s,t_0)$ and $dict.(v, t)$ is the length of the shortest temporal path
 from $(s,t_0)$ to $(v,t)$.

### 2016-01-22 paper.tex
walk through Algorithm 1

@@
 the forward-neighbors of $(2, t_1)$ is $(2, t_3)$.
-See Algorithm  \ref{alg:bfs} for the evolving-graph BFS algorithm.
+
+Given an evolving graph $G_n$ and a starting active node $(s,t)$, Algorithm \ref{alg:bfs} 
+returns all the reachable active nodes and their distances from $(s,t)$. 
+For Figure \ref{fig:eg_shortest_path}, suppose we start with $(1, t_2)$, then
+$frontier$ is $\langle(1,t_2)\rangle$ after step $5$.  Since the only forward-neighbor of 
+$(1,t_2)$ is $(3,t_2)$, $dict.(3,t_2) =1$ at step $11$ and $next$ is enqueued to be 
+$\langle (3,t_2) \rangle$ at step $12$. After that $frontier$ is equal to $next$. Since 
+$frontier$ is not empty, we go through step $7$-$12$ again. This time 
+the forward-neighbor of $(3,t_2)$ is $(3,t_3)$, so $dict.(3,t_3)=2$ and 
+$next = \langle (3,t_3) \rangle$ and so on. The loop terminates when $frontier$ is empty.
 

### 2016-01-22 paper.tex
add acknowledgement

@@
  \Fn{ABFS($A_n, (s,t)$)}{
-   Form $\bm{A}^T$ from $A_n$\\
-   $m = A^{[1]}.dimension$ \\
+   Form $\bm{A}^T$ from $A_n$.\\
+   Let $m$ be the dimension of  $A^{[1]}$. \\
    $b$ is a sparse vector of length $mn$ \\
@@
 Algorithm \ref{alg:bfsla} outlines the basic idea of an algebraic evolving-graph BFS.
-But in practice we don't need to form $\bm{A}$ and $\bm{b}$ explicitly at step $4$ and $6$. By realizing the dependency between these sparse matrices,
+But in practice we don't need to form $\bm{A}$ explicitly at step $2$. By realizing the dependency between these sparse matrices,
 we can compute  the result in parallel.
@@
 
-% % use section* for acknowledgement
-% \section*{Acknowledgment}
+\section*{Acknowledgment}
 
-% The authors would like to thank...
-% more thanks here
+The authors are grateful to Professor Nicholas J. Higham and Doctor Vedran \u Sego for suggestions.
+This work was completed while the second author visited MIT CSAIL. He thanks 
+Professor Alan Edelman for making his stay so pleasant.
 

### 2016-01-22 paper.tex
update theorem 1

@@
 For any two nodes $v_s$ and $v_t$ in $G$,  if $s < t$, we add an edge from $v_s$ to $v_t$.
-By construction, the forward-neighbors of an active node $(v, t)$ in $G_n$
+By construction, an active node $(v,t)$ in $G_n$ is in one-to-one correspondence with
+a node $v_t$ in $G$.
+The forward-neighbors of an active node $(v, t)$ in $G_n$
 are in one-to-one correspondence with the forward-neighbors of a node $v_t$ in $G$.
@@
 active nodes will not be visited twice, which is equivalent to step $10$
-of Algorithm \ref{alg:bfs}. Finally, a dictionary $dict$ is used in both algorithms
+of Algorithm \ref{alg:bfs}. Finally, a dictionary $dict$ is used by both algorithms
 to keep record of the distance of each node from the source $(s,t)$.

### 2016-01-22 paper.tex
various minor changes

@@
 
-Given a root node $(s, t)$, during the execution the BFS discussed above builds a search tree, denoted by $T(s,t)$,
-where nodes are temporally connected.
+Given a root node $(s, t)$, during the execution the evolving-graph BFS builds a search tree, denoted by $T(s,t)$,
+where nodes in $T(s,t)$ are temporally connected.
 \begin{lemma}
@@
 Hence matrix-matrix multiplication between consecutive
-adjacency matrices should be avoid  for computing temporal paths.
+adjacency matrices can not discover all possible temporal paths and should be avoided 
+when temporal paths are the essential parts of the algorithm.
 
@@
 
-In order to study evolving graphs and experiment with various graph types, we have developed EvolvingGraphs.jl  \cite{zhang15}, a software package in the Julia language \cite{bkse12} for the creation, manipulation, and study of dynamic networks.
+To study evolving graphs and experiment with various graph types, we have developed EvolvingGraphs.jl  \cite{zhang15}, a software package in the Julia language \cite{bkse12} for the creation, manipulation, and study of dynamic networks.
 Julia is a high-level, high-performance dynamic programming language for technical

### 2016-01-22 paper.tex
update implementation section

@@
 
-To study evolving graphs and experiment with various graph types, we have developed EvolvingGraphs.jl  \cite{zhang15}, a software package in the Julia language \cite{bkse12} for the creation, manipulation, and study of dynamic networks.
-Julia is a high-level, high-performance dynamic programming language for technical
-computing. When we implement EvolvingGraphs.jl, we make particular use of
-the parameterizable type system and multiple dispatch in Julia.
-We define a data type  \texttt{MatrixList}, which represents an evolving graph
+To study evolving graphs and experiment with various graph types, we have developed EvolvingGraphs.jl  \cite{zhang15}, a software package for the creation, manipulation, and study of evolving graphs written in Julia \cite{bkse12}.
+The implementation of the evolving-graph BFS is
+available online at 
+\begin{center}
+\url{https://github.com/weijianzhang/EvolvingGraphs.jl}.
+\end{center}
+%Julia is a high-level, high-performance dynamic programming language for technical
+%computing. 
+% We make particular use of
+% the parameterizable type system and multiple dispatch in Julia.
+ \texttt{MatrixList}, a data type in EvolvingGraphs.jl,  represents an evolving graph
 as a sequence of sparse matrices stored in compressed sparse column (CSC) format.
-Many basic graph functions are defined for \texttt{MatrixList}, including
-retrieving nodes, time stamps, and matrices.
+Many basic graph functions are implemented for \texttt{MatrixList}.
 

### 2016-01-22 paper.tex
update application section

@@
 time $t_0$. We define a \emph{community} to be a group of
-researchers that share the same information source.
+researchers that have been influenced by the same authors.
 For example, given a paper published by $a$ at year $2016$,
 we can determine $a$'s community by first backward searching to
-find the source then forward searching to find the descendants.
+find the root authors then forward searching to find the descendants.
 
-To design such system, we shall use two parameters:
- a parameter for
-determining the layers of the backward search tree which can be chosen by the user;
-a second parameter for determining the number of roots as a means of
-filtering out unessential or overlapping trees. The subset of roots can be determined
-by ranking algorithms like PageRank or machine-learning based preference ranking.
+To design such information retrieval system, two parameters will be needed:
+ one for determining the layers of the backward search tree which can be 
+adjusted by the user;
+the other for determining root authors for forward searching 
+so that we can 
+filter out unessential or overlapped trees, which can be computed
+by ranking algorithms like PageRank \cite{pbmw99} or machine-learning based preference ranking.
 

### 2016-01-22 paper.tex
update conclusion

@@
 In this paper, we have introduced and analyzed the evolving-graph BFS in both
-graph and algebraic notation.
+graph and algebraic notation. The relationship between active nodes and 
+evolving graph algorithms is discussed.
 This paper leaves important extensions open for further research. First, it is
@@
 is the matrix representation is highly structured so it is easy to design
-high performance parallel implementations.
+high performance parallel implementation.
 Third, depth-first search and other graph traversal algorithms can be
-extended for searching evolving graphs using similar ideas.
+extended using similar ideas.
 

### 2016-01-22 paper.tex
Begin to restructure introduction

@@
 
-Let's imagine a game played by three people
-$A$, $B$, and $C$. Suppose each of them has a piece of information,
-say $X$, $Y$, and $Z$ respectively. The goal of the game is
-to get information from other people as much as possible.
-Suppose $A$ talks to $B$ first and then $B$ talks to $C$,
-then $C$ may has all three pieces of information $X$, $Y$, and $Z$ even
-if $C$ has never talked to $A$.
-However, if $B$ talks to $C$ first and then $A$ talks to $B$, then
-no matter how smart $C$ is, $C$ can never get information
-$X$.  Time is an important factor for information flow.
+Let's imagine a game played by three people, numbered $1$, $2$, and $3$,
+each of whom has a message, labeled $a$, $b$, and $c$ respectively.
+At each turn, one particular player is allowed to talk to one other player,
+who must in turn convey all the messages in his or her possession.
+The goal of the game is to collect all the messages.
+Suppose $1$ talks to $2$ first, and $2$ in turn talks to $3$.
+Then, $3$ can collect all the messages even without talking to $1$ directly.
+However, if $2$ talks to $3$ before $1$ talks to $2$, then
+$3$ can never get $a$.
 
-We can analyze the spread of information between $A$, $B$, and $C$
-as a process on a graph that changes over time.
-Such graphs are known as evolving graphs.
-% Evolving graphs are models to study this kind of scenario where the
-% interrelations between entities are time-dependent.
-% We assume that the reader is familiar with the basic graph terminology.
-% All the necessary material can be found in}.
+We can use analyze the spread of information between the players using graph
+theory. In this process, the time ordering of events matters, which leads us to
+consider a graph that changes over time, i.e.\ a graph $G(t) = (V(t), E(t))$ whose
+vertex set $V$ and edge set $E$ evolve over on time, $t$.
+Such a graph is called an ``evolving graph''~\cite{fkp89,bkmu12},
+``evolving network'' or %XXX WHO?
+``time-dependent graph''. %XXX WHO?
+
+Treatments of evolving graphs vary in their generality and focus.
+Kivel{\"a} \textit{et al.}~\cite{kabg14} treats time dependence as a special
+case of families of graphs with multiple interrelationships.
+Others like Flajolet \textit{et al.}~\cite{fkp89} use time to index the family
+of related graph, but are not concerned with explicit time-dependent processes.
+
+Our model can also be seen as a special case of
+the multi-layer networks discussed by Kivel{\"a} \textit{et al.}~\cite{kabg14}, where
+there is only one set of layers for time.
+
+%XXX is the first sentence of this restatement true?
+Previous work on evolving graphs tends to focus on special cases, such as
+random graphs~\cite{fkp89} or sequences of graphs whose successive differences
+are small~\cite{bkmu12}.  use the term to mean the evolution of random graphs.
+Bahmani et al. \cite{bkmu12} use the term to mean a sequence of graphs where
+successive differences are small.
+
+Here, we are interested in extensions of graph algorithms that generalize for evolving graphs that work in full
+generality.
+
+%ARE THIS RELATED TO THE PREVIOUS WORK HOW
+Work in this area include Despite the fact that graph metrics \cite{ntmm13,tmml09,tsmm09} and
+various centrality measures like Katz centrality~\cite{gphe11,grihig13},
+betweenness \cite{alhi15}, and closeness \cite{tmml10})
+have been generalized for evolving graphs,
+a BFS for evolving graphs, which we shall call it \emph{evolving-graph BFS},
+was never explicitly described or verified.
+%XXX THIS LEADS US TO CONSIDER ALGEBRAIC FORMULATIONS ALSO SINCE...
+
+In this paper, we present a generalization of the breadth first search (BFS)
+algorithm to evolving graphs. BFS is arguably the most fundamental graph search
+algorithm, which leads to other graphical computations
+such as the minimum spanning tree and shortest path \cite{ckrs09}.
+We also demonstrate how BFS on evolving graphs can be expressed in the language
+of linear algebra, and how our formulation helps analyze citation networks.
+
+\section{Breadth-First Search for Evolving Graphs}
+\label{sec:breadth-first-search}
+
+\subsection{Temporal Paths over Active Nodes}
+
+\begin{figure}[h]
+ \begin{center}
+    \begin{tikzpicture}[scale=.38, line width =1.5pt]
+  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {2};
+  \node[circle,draw=red1, minimum size=0.2cm] (n8) at (-10,5) {3};
+  \node[circle,fill=green1, minimum size=0.2cm] (n10) at (-8,10) {1};
+
+  \node[circle,draw=red1, minimum size=0.2cm] (n6) at (3,7) {2};
+  \node[circle,fill=green1, minimum size=0.2cm] (n4) at (-2,5) {3};
+  \node[circle,fill=green1, minimum size=0.2cm] (n1) at (0,10) {1};
+
+   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {2};
+  \node[circle,fill=green1, minimum size=0.2cm] (n12) at (6,5)  {3};
+  \node[circle,draw=red1, minimum size=0.2cm] (n14) at (8,10) {1};
+
+  \foreach \from/\to in {n10/n7, n1/n4, n11/n12}
+   \draw[every edge,->] (\from) -- (\to);
+\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
+(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{ $t_2$};
+\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
+(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $t_1$};
+\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
+(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $t_3$};
+    \end{tikzpicture}
+\end{center}
+\caption{An evolving graph with 3 time stamps $t_1$, $t_2$ and $t_3$.
+At each time stamp, the evolving graph is represented as a graph.
+The green filled nodes in the figure are active nodes.}
+\label{fig:eg_shortest_path}
+\end{figure}
+
+The key new idea in generalizing BFS to evolving graphs is to be able to compute
+traversals that span multiple times. We call some, but not all,
+such traversals \textit{temporal paths}.
+
+Figure~\ref{fig:eg_shortest_path} shows a small example of a evolving
+directed graph, $G_3 = \langle G^{[1]}, G^{[2]}, G^{[3]}\rangle$, consisting of
+a sequence of three graphs $G^{[i]}$, each bearing a time stamp $t_i$.
+There are directed edges $1\rightarrow2$ at time $t_1$,
+$1\rightarrow3$ at time $t_2$, and $2\rightarrow3$ at time $t_3$.
+Each edge exists only at a particular discrete time and the nodes connected
+by edges are considered \textit{active} at that time.
+
+Temporal paths connect only active nodes in ways that respect time ordering.
+Thus the sequences
+$\langle (1, t_1), (1, t_2), (3, t_2), (3, t_3) \rangle$ and
+$\langle (1, t_1), (2, t_1), (2, t_3), (3, t_3) \rangle$
+are both examples of \textit{temporal paths} from $(1, t_1)$ to $(3, t_3)$,
+which are drawn as dotted lines with arrowheads in Figure~\ref{fig:active}.
+However, $\langle (1, t_1), (1, t_2), (2, t_2), (3, t_2), (3, t_3) \rangle$
+is not a temporal path because node 2 is inactive at time $t_2$.
+Similarly,
+
+\begin{figure}[h]
+ \begin{center}
+    \begin{tikzpicture}[scale=.36, line width =1.4pt]
+  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {2};
+  \node[circle,draw=red1, minimum size=0.2cm] (n8) at (-10,5) {3};
+  \node[circle,fill=green1, minimum size=0.2cm] (n10) at (-8,10) {1};
+
+  \node[circle,draw=red1, minimum size=0.2cm] (n6) at (3,7) {2};
+  \node[circle,fill=green1, minimum size=0.2cm] (n4) at (-2,5) {3};
+  \node[circle,fill=green1, minimum size=0.2cm] (n1) at (0,10) {1};
+
+   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {2};
+  \node[circle,fill=green1, minimum size=0.2cm] (n12) at (6,5)  {3};
+  \node[circle,draw=red1, minimum size=0.2cm] (n14) at (8,10) {1};
+
+  \foreach \from/\to in {n10/n7, n11/n12}
+   \draw[every edge,->] (\from) -- (\to);
+     \draw[dotted,->](n1) -- (n4);
+  \draw[dotted,->](n10) -- (n1);
+  \draw[dotted,->](n4) -- (n12);
+\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
+(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{ $t_2$};
+\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
+(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $t_1$};
+\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
+(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $t_3$};
+    \end{tikzpicture}
+   \begin{tikzpicture}[scale=.36, line width =1.4pt]
+  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {2};
+  \node[circle,draw=red1, minimum size=0.2cm] (n8) at (-10,5) {3};
+  \node[circle,fill=green1, minimum size=0.2cm] (n10) at (-8,10) {1};
+
+  \node[circle,draw=red1, minimum size=0.2cm] (n6) at (3,7) {2};
+  \node[circle,fill=green1, minimum size=0.2cm] (n4) at (-2,5) {3};
+  \node[circle,fill=green1, minimum size=0.2cm] (n1) at (0,10) {1};
+
+   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {2};
+  \node[circle,fill=green1, minimum size=0.2cm] (n12) at (6,5)  {3};
+  \node[circle,draw=red1, minimum size=0.2cm] (n14) at (8,10) {1};
+
+  \foreach \from/\to in {n1/n4}
+   \draw[every edge,->] (\from) -- (\to);
+     \draw[dotted,->](n7) to[out=70, in=40] (n11);
+  \draw[dotted,->](n10) -- (n7);
+   \draw[dotted,->](n11) -- (n12);
+\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
+(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{ $t_2$};
+\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
+(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $t_1$};
+\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
+(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $t_3$};
+    \end{tikzpicture}
+\end{center}
+\caption{An evolving graph with 3 time stamps $t_1$, $t_2$ and $t_3$.
+The black dashed lines of the two figures indicate two temporal paths from $(1, t_1)$ to $(3,t_3)$.}
+\label{fig:active}
+\end{figure}
+
+The restriction that temporal paths may only traverse active nodes reflects
+underlying structure in many real world applications, such as analyzing influence
+of nodes over social networks, and furthermore leads to interesting and nontrivial
+adaptations of graph algorithms such as BFS
+
+\subsection{Evolving-Graph BFS}
+\label{sec:evolving-graph-bfs}
+
+This example shows a directed
+evolving graph
+For example,
+at time stamp $t_1$, the graph $G^{[1]}$ contains three nodes $1$, $2$ and $3$ and $(1,2)$ is an edge of $G^{[1]}$.
+Though Figure \ref{fig:eg_shortest_path} presents a
+directed evolving graph,  the algorithms that we shall discuss work on both
+directed and undirected evolving graphs.
+We shall use this example throughout the paper to illustrate various
+notions.
+
 We follow the notation of (static) graphs in \cite{even12,kegi11}.
-In general, an evolving graph $G(t) = (V(t), E(t))$ is  a graph where the node set
-and edge set are functions of time $t$.
-However, in different literature the term ``evolving graph'' has been
-used to refer to slightly different models.
- Flajolet et al. \cite{fkp89} use the term to mean the evolution of random graphs.
-Bahmani et al. \cite{bkmu12} use the term to mean a sequence of graphs where
-the difference between two consecutive graphs is small.
-Similar models are also investigated using the term ``evolving network'' and
-``time-dependent graph''.
-
 In this work, we consider a discretization of the time line as $n$ distinct
@@
 where an evolving graph is considered as a sequence of adjacency matrices.
-Our model can also be seen as a special case of
-the multi-layer networks discussed by Kivel{\"a} et al. \cite{kabg14}, where
-there is only one set of layers for time.
 % We present a preliminary discussion of using the model to mine citation networks.
@@
 
-\section{Breadth-First Search for Evolving Graphs}
-\label{sec:breadth-first-search}
-
-\subsection{Evolving-Graph BFS}
-\label{sec:evolving-graph-bfs}
-
-\begin{figure}[h]
- \begin{center}
-    \begin{tikzpicture}[scale=.38, line width =1.5pt]
-  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {2};
-  \node[circle,draw=red1, minimum size=0.2cm] (n8) at (-10,5) {3};
-  \node[circle,fill=green1, minimum size=0.2cm] (n10) at (-8,10) {1};
-
-  \node[circle,draw=red1, minimum size=0.2cm] (n6) at (3,7) {2};
-  \node[circle,fill=green1, minimum size=0.2cm] (n4) at (-2,5) {3};
-  \node[circle,fill=green1, minimum size=0.2cm] (n1) at (0,10) {1};
-
-   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {2};
-  \node[circle,fill=green1, minimum size=0.2cm] (n12) at (6,5)  {3};
-  \node[circle,draw=red1, minimum size=0.2cm] (n14) at (8,10) {1};
-
-  \foreach \from/\to in {n10/n7, n1/n4, n11/n12}
-   \draw[every edge,->] (\from) -- (\to);
-\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
-(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{ $t_2$};
-\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
-(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $t_1$};
-\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
-(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $t_3$};
-    \end{tikzpicture}
-\end{center}
-\caption{An evolving graph with 3 time stamps $t_1$, $t_2$ and $t_3$.
-At each time stamp, the evolving graph is represented as a graph.
-The green filled nodes in the figure are active nodes.}
-\label{fig:eg_shortest_path}
-\end{figure}
-
-We show a small evolving graph $ \langle G^{[1]}, G^{[2]}, G^{[3]}\rangle$ with $3$ time stamps $t_1$, $t_2$ and $t_3$
-in Figure \ref{fig:eg_shortest_path}, where each $G^{[i]}$ is a graph at time stamp $t_i$.
-For example,
-at time stamp $t_1$, the graph $G^{[1]}$ contains three nodes $1$, $2$ and $3$ and $(1,2)$ is an edge of $G^{[1]}$.
-Though Figure \ref{fig:eg_shortest_path} presents a
-directed evolving graph,  the algorithms that we shall discuss work on both
-directed and undirected evolving graphs.
-We shall use this example throughout the paper to illustrate various
-notions.
-
 Formally, an evolving graph $G_n $ is defined to be a sequence of graphs $\langle G^{[1]}, G^{[2]}, \ldots, G^{[n]}\rangle$,  where each $G^{[t]} = (V^{[t]}, E^{[t]})$ represents a graph at time stamp $t$.
@@
 $t_i \ne t_j$. The length of a temporal path is the length of the sequence.
-For example, $\langle (1, t_1), (1, t_2), (3, t_2), (3, t_3) \rangle$ and
-$\langle (1, t_1), (2, t_1), (2, t_3), (3, t_3) \rangle$
-are two temporal paths from $(1, t_1)$ to $(3, t_3)$
-on the evolving graph in Figure \ref{fig:eg_shortest_path}.
 
@@
 
-\begin{figure}[h]
- \begin{center}
-    \begin{tikzpicture}[scale=.36, line width =1.4pt]
-  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {2};
-  \node[circle,draw=red1, minimum size=0.2cm] (n8) at (-10,5) {3};
-  \node[circle,fill=green1, minimum size=0.2cm] (n10) at (-8,10) {1};
-
-  \node[circle,draw=red1, minimum size=0.2cm] (n6) at (3,7) {2};
-  \node[circle,fill=green1, minimum size=0.2cm] (n4) at (-2,5) {3};
-  \node[circle,fill=green1, minimum size=0.2cm] (n1) at (0,10) {1};
-
-   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {2};
-  \node[circle,fill=green1, minimum size=0.2cm] (n12) at (6,5)  {3};
-  \node[circle,draw=red1, minimum size=0.2cm] (n14) at (8,10) {1};
-
-  \foreach \from/\to in {n10/n7, n11/n12}
-   \draw[every edge,->] (\from) -- (\to);
-     \draw[dotted,->](n1) -- (n4);
-  \draw[dotted,->](n10) -- (n1);
-  \draw[dotted,->](n4) -- (n12);
-\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
-(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{ $t_2$};
-\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
-(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $t_1$};
-\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
-(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $t_3$};
-    \end{tikzpicture}
-   \begin{tikzpicture}[scale=.36, line width =1.4pt]
-  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {2};
-  \node[circle,draw=red1, minimum size=0.2cm] (n8) at (-10,5) {3};
-  \node[circle,fill=green1, minimum size=0.2cm] (n10) at (-8,10) {1};
-
-  \node[circle,draw=red1, minimum size=0.2cm] (n6) at (3,7) {2};
-  \node[circle,fill=green1, minimum size=0.2cm] (n4) at (-2,5) {3};
-  \node[circle,fill=green1, minimum size=0.2cm] (n1) at (0,10) {1};
-
-   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {2};
-  \node[circle,fill=green1, minimum size=0.2cm] (n12) at (6,5)  {3};
-  \node[circle,draw=red1, minimum size=0.2cm] (n14) at (8,10) {1};
-
-  \foreach \from/\to in {n1/n4}
-   \draw[every edge,->] (\from) -- (\to);
-     \draw[dotted,->](n7) to[out=70, in=40] (n11);
-  \draw[dotted,->](n10) -- (n7);
-   \draw[dotted,->](n11) -- (n12);
-\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
-(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{ $t_2$};
-\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
-(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $t_1$};
-\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
-(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $t_3$};
-    \end{tikzpicture}
-\end{center}
-\caption{An evolving graph with 3 time stamps $t_1$, $t_2$ and $t_3$.
-The back dashed lines of the two figures indicate two temporal paths from $(1, t_1)$ to $(3,t_3)$.}
-\label{fig:active}
-\end{figure}
-
 Node activeness is the key concept for defining temporal paths which

### 2016-01-22 paper.tex
Finish clean up of introduction

@@
 Treatments of evolving graphs vary in their generality and focus.
-Kivel{\"a} \textit{et al.}~\cite{kabg14} treats time dependence as a special
+Kivel\"a \textit{et al.}~\cite{kabg14} treats time dependence as a special
 case of families of graphs with multiple interrelationships.
 Others like Flajolet \textit{et al.}~\cite{fkp89} use time to index the family
-of related graph, but are not concerned with explicit time-dependent processes.
+of related graphs, but are not concerned with explicit time-dependent processes.
+Yet others focus on incremental updates to large graphs~\cite{bkmu12}.
+Here, we describe evolving graphs as a time-ordered sequence of graphs, similar
+to the study of metrized graphs by Tang and coworkers~\cite{ntmm13,tmml09,tsmm09,tmml10}
+the dynamics of communities by Grindrod, Higham and coworkers~\cite{gphe11,grihig13}.
 
-Our model can also be seen as a special case of
-the multi-layer networks discussed by Kivel{\"a} \textit{et al.}~\cite{kabg14}, where
-there is only one set of layers for time.
+The game described at the beginning can be encoded in terms of analyzing the
+microscopic dynamics of an evolving graph. The spread of information to the
+winner can be described in terms of discrete paths through the evolving graph,
+and can be interpreted naturally as a graph traveral problem. Traversals of an
+ordinary (static) graph may be computed using well known methods such as the
+breadth-first search (BFS). However, a proper description of the dynamics
+requires paths that also move monotonically forward in time, thus necessating
+a generalization of the ordinary BFS algorithm. An informal description of such
+a generalization was provided by \cite{tmml09}.
 
-%XXX is the first sentence of this restatement true?
-Previous work on evolving graphs tends to focus on special cases, such as
-random graphs~\cite{fkp89} or sequences of graphs whose successive differences
-are small~\cite{bkmu12}.  use the term to mean the evolution of random graphs.
-Bahmani et al. \cite{bkmu12} use the term to mean a sequence of graphs where
-successive differences are small.
-
-Here, we are interested in extensions of graph algorithms that generalize for evolving graphs that work in full
-generality.
-
-%ARE THIS RELATED TO THE PREVIOUS WORK HOW
-Work in this area include Despite the fact that graph metrics \cite{ntmm13,tmml09,tsmm09} and
-various centrality measures like Katz centrality~\cite{gphe11,grihig13},
-betweenness \cite{alhi15}, and closeness \cite{tmml10})
-have been generalized for evolving graphs,
-a BFS for evolving graphs, which we shall call it \emph{evolving-graph BFS},
-was never explicitly described or verified.
-%XXX THIS LEADS US TO CONSIDER ALGEBRAIC FORMULATIONS ALSO SINCE...
-
-In this paper, we present a generalization of the breadth first search (BFS)
-algorithm to evolving graphs. BFS is arguably the most fundamental graph search
-algorithm, which leads to other graphical computations
-such as the minimum spanning tree and shortest path \cite{ckrs09}.
-We also demonstrate how BFS on evolving graphs can be expressed in the language
-of linear algebra, and how our formulation helps analyze citation networks.
+In this paper, we present a complete description of the BFS algorithm for evolving graphs.
+The algebraic formulation of this algorithm makes explicit several subtleties in the correct
+determination of relevant trajectories, which we term \textit{temporal paths},
+which in turn can have important consequences for the computation and interpretation
+of quantities which can be expressed as sums over paths, e.g.\ Katz centrality.
 

### 2016-01-22 paper.tex
Clean up section on temporal paths

@@
 The key new idea in generalizing BFS to evolving graphs is to be able to compute
-traversals that span multiple times. We call some, but not all,
-such traversals \textit{temporal paths}.
+paths that traverse multiple times. We call the relevant paths
+\textit{temporal paths}.
 
@@
 The restriction that temporal paths may only traverse active nodes reflects
-underlying structure in many real world applications, such as analyzing influence
-of nodes over social networks, and furthermore leads to interesting and nontrivial
-adaptations of graph algorithms such as BFS
+underlying structure in many real world applications, such as analyzing the
+influence of nodes over social networks. We will also show later in
+%XXX FIX X-REF TO SECTION
+that the resulting structure of allowable temporal paths leads to nontrivial
+subtitled in the generalization of algorithms and concepts from ordinary
+(static) graphs.
 

### 2016-01-22 paper.tex
Formalize definitions

@@
 \newtheorem{thm}{Theorem}
+\newtheorem{defn}{Definition}
 \theoremstyle{definition}
@@
 
-\section{Breadth-First Search for Evolving Graphs}
+\section{Breadth-First Search over Evolving Graphs}
 \label{sec:breadth-first-search}
@@
 that the resulting structure of allowable temporal paths leads to nontrivial
-subtitled in the generalization of algorithms and concepts from ordinary
+subtleties in the generalization of algorithms and concepts from ordinary
 (static) graphs.
 
-\subsection{Evolving-Graph BFS}
+\subsection{Breadth-First Traversal Over Temporal Paths}
 \label{sec:evolving-graph-bfs}
 
-This example shows a directed
-evolving graph
-For example,
-at time stamp $t_1$, the graph $G^{[1]}$ contains three nodes $1$, $2$ and $3$ and $(1,2)$ is an edge of $G^{[1]}$.
-Though Figure \ref{fig:eg_shortest_path} presents a
-directed evolving graph,  the algorithms that we shall discuss work on both
-directed and undirected evolving graphs.
-We shall use this example throughout the paper to illustrate various
-notions.
+The example presented above in Section~\ref{sec:breadth-first-search}
+demonstrates how the notion of active nodes restrict the set of temporal paths
+that need to be considered when traversing an evolving graph.
 
-We follow the notation of (static) graphs in \cite{even12,kegi11}.
-In this work, we consider a discretization of the time line as $n$ distinct
-time stamps and define an evolving graph $G_n$ to be
-a sequence of (static) graphs
-$G_n = \langle G^{[1]}, G^{[2]}, \ldots, G^{[n]} \rangle$.
-Depends on applications, the graph $G^{[i]}$ could mean either
-a snapshot of $G(t)$ at time stamp $i$ or  the union of all the graphs
-between time stamp $i-1$ and $i$ (excludes time stamp $i-1$).
-Both the node set and edge set of $G^{[i]}$ could be different from
-these of $G^{[j]}, j \ne i$.
+We now give a general description of the BFS algorithm over evolving graphs,
+both directed and undirected, which correctly takes into account the structure
+of temporal paths.
+Our notation generalizes the standard notations for static graphs presented in
+\cite{even12,kegi11}.
+
+\begin{defn}
+An \textit{evolving graph} $G_n$ is a sequence of
+(static) graphs
+$G_n = \langle G^{[1]}, G^{[2]}, \ldots, G^{[n]} \rangle$ with associated
+time labels $t_1, t_2, \ldots, t_n$ respectively.
+Each $G^{[t]} = (V^{[t]}, E^{[t]})$ represents a (static) graph labeled by a
+time $t$.
+\end{defn}
+
+Intuitively, an evolving graph is some discretization of the continuous-time family $G(t)$:
+
 \begin{center}
@@
 \end{center}
-Our model is the same as that of Tang et al. \cite{tmml10} though they
-use slightly different notation.
+
+We assume no particular relation between the node and edge sets for each static
+graph $G^{[t]} = (V^{[t]}, E^{[t]})$.
+
+\begin{defn}
+A \textit{temporal node} is a pair $(v, t)$, where $v\in V^{[t]}$ is a node at
+a time $t$.
+\end{defn}
+
+\begin{defn}
+An \textit{active node} is a temporal node $(v,t)$ if there exists at least one
+edge $e\in E^{[t]}$ that connects $v\in V^{[t]}$ to another node $w\in V^{[t]}$,
+$w\ne v$.
+
+An \textit{inactive node} is a temporal node that is not an active node.
+\end{defn}
+
+In Figure~\ref{fig:eg_shortest_path}, the temporal nodes $(1,t_1)$ and $(2,t_2)$
+are active nodes, whereas the temporal node $(3,t_1)$ is an inactive node.
+
+\begin{defn}
+A \textit{temporal path on an evolving graph $G_n$ from temporal node $(v_1, t_1)$
+to temporal node $(v_m, t_m)$} is a time-ordered sequence of temporal notes,
+$\langle (v_1, t_1), (v_2, t_2), \ldots, (v_m, t_m) \rangle$.
+Here, time ordering means that $t_1 \leq t_2 \leq \cdots \leq t_m$ and
+$v_i = v_j$ iff $t_i \ne t_j$.
+The length of a temporal path is the length of the sequence.
+\end{defn}
+
+This definition of a temporal path is essentially that of the dynamic walk
+in \cite{gphe11,grihig13}.
+
 Grindrod et al.'s model  \cite{gphe11} \cite{grihig13} is very similar to ours,
@@
 % an application of the evolving-graph BFS in mining citation networks is described in Section \ref{sec:applications}.
-
-Formally, an evolving graph $G_n $ is defined to be a sequence of graphs $\langle G^{[1]}, G^{[2]}, \ldots, G^{[n]}\rangle$,  where each $G^{[t]} = (V^{[t]}, E^{[t]})$ represents a graph at time stamp $t$.
-We shall use the notation $(v, t)$ to represent
-a node $v$ at time stamp $t$.
-We say two nodes are \emph{connected} if there is an edge between them
-and we say a node $(v,t)$ is \emph{active}, if $v$ is connected to at least a node other than itself at time stamp $t$. Otherwise we say it is \emph{inactive}.
-In Figure \ref{fig:eg_shortest_path}, node $(1,t_1)$ and $(2,t_2)$  are active but node $(3,t_1)$ is inactive.
-
-In order to define the evolving-graph BFS, we need the concept of a
-\emph{temporal path}.
-A temporal path on an evolving graph $G_n$ is a sequence of node-time stamp
-pair $\langle (v_1, t_1), (v_2, t_2), \ldots, (v_m, t_m) \rangle$ such that
-$t_1 \leq t_2 \leq \cdots \leq t_m$ and $v_i = v_j$ if and only if
-$t_i \ne t_j$. The length of a temporal path is the length of the sequence.
 

### 2016-01-23 paper.tex
Clean up definitions leading up to the description of BFS

@@
 \begin{defn}
-An \textit{evolving graph} $G_n$ is a sequence of
+An \textbf{evolving graph} $G_n$ is a sequence of
 (static) graphs
@@
 We assume no particular relation between the node and edge sets for each static
-graph $G^{[t]} = (V^{[t]}, E^{[t]})$.
+graph $G^{[t]} = (V^{[t]}, E^{[t]})$. In particular, we allow the node sets to
+change over time, so that each $V^{[t]}$ may be different. Changing
+node sets happen naturally in citation networks, where nodes may appear or
+disappear from the citation network over time.
+The addition, removal, or relabeling of nodes can be expressed in terms
+of a map $\Pi^{[t,t^\prime]} : V^{[t]} \rightarrow V^{[t^\prime]}$ that expresses
+the appropriate permutations and/or projections.
 
 \begin{defn}
-A \textit{temporal node} is a pair $(v, t)$, where $v\in V^{[t]}$ is a node at
+A \textbf{temporal node} is a pair $(v, t)$, where $v\in V^{[t]}$ is a node at
 a time $t$.
@@
 \begin{defn}
-An \textit{active node} is a temporal node $(v,t)$ if there exists at least one
+An \textbf{active node} is a temporal node $(v,t)$ if there exists at least one
 edge $e\in E^{[t]}$ that connects $v\in V^{[t]}$ to another node $w\in V^{[t]}$,
@@
 
-An \textit{inactive node} is a temporal node that is not an active node.
+An \textbf{inactive node} is a temporal node that is not an active node.
 \end{defn}
@@
 \begin{defn}
-A \textit{temporal path on an evolving graph $G_n$ from temporal node $(v_1, t_1)$
-to temporal node $(v_m, t_m)$} is a time-ordered sequence of temporal notes,
+A \textbf{temporal path} of length $m$ on an evolving graph $G_n$
+from temporal node $(v_1, t_1)$ to temporal node $(v_m, t_m)$} is
+a time-ordered sequence of active nodes,
 $\langle (v_1, t_1), (v_2, t_2), \ldots, (v_m, t_m) \rangle$.
@@
 $v_i = v_j$ iff $t_i \ne t_j$.
-The length of a temporal path is the length of the sequence.
 \end{defn}
 
-This definition of a temporal path is essentially that of the dynamic walk
-in \cite{gphe11,grihig13}.
-
-Grindrod et al.'s model  \cite{gphe11} \cite{grihig13} is very similar to ours,
-where an evolving graph is considered as a sequence of adjacency matrices.
-% We present a preliminary discussion of using the model to mine citation networks.
-
-% why BFS is important?
-
-% Given a graph and a source node $s$,
-% the breadth-first search (BFS) systematically discover every node that is reachable from $s$.
-% As the name suggested, BFS expands the frontier of discovered regions
-% uniformly across the breadth, i.e., the algorithm discovers all nodes
-% that are $k$ walks away from $s$ before discovering any nodes that are
-% $k+1$ walks away.
-A temporal path is the basic element to
- define almost  all important evolving graph algorithms.
-Matrix-matrix multiplication between consecutive adjacency matrices
-were commonly used as a means of determining temporal paths \cite{gphe11,grihig13}. However, we show
-in one specific case it fails to include all the temporal paths and thus should not be used
-for searching evolving graphs in general.
-
-Breadth-first search (BFS) is probably the most basic graph searching algorithm and
-it is also the building block of other graph algorithms like minimum spanning tree
-and shortest path \cite{ckrs09}.
-Despite the fact that graph metrics \cite{ntmm13,tmml09,tsmm09} and
-various centrality measures like Katz centrality~\cite{gphe11,grihig13},
-betweenness \cite{alhi15}, and closeness \cite{tmml10})
-have been generalized for evolving graphs,
-a BFS for evolving graphs, which we shall call it \emph{evolving-graph BFS},
-was never explicitly described or verified.
-Therefore, there is a need for clear presentations of the algorithm.
-
-In this paper,  we present descriptions of an evolving-graph BFS algorithm, but its main concern is with its sparse matrix
-representation and linear algebraic computations.
-We prove that the two algorithms (graph notation and algebra notation) are equivalent.
-
-% GraphBlas related ideas
-
-% This paper is organized as follows. The evolving-graph BFS algorithm is described
-% and analyzed using graph notation in Section \ref{sec:breadth-first-search}. Section \ref{sec:evolv-graph-algor} considers the algebraic representation of the algorithm:
-% A sparse matrix representation
-% of an evolving graph is presented in \ref{sec:representation}.
-% The algorithm is presented in \ref{sec:breadth-first-search-1} and
-% the properties of the search tree is discussed in \ref{sec:short-temp-path}.
-% The connection between a temporal path and a matrix-matrix multiplication is considered and discussed in \ref{sec:matr-matr-mult}. Finally,
-% an application of the evolving-graph BFS in mining citation networks is described in Section \ref{sec:applications}.
-
-We would like a BFS on evolving graphs to work similar to that on graphs: the search starts with some source node $(s,t)$, called \emph{root}.
-It constructs a tree through discovering all nodes that are reachable by
-temporal paths of length $k$ before moving to all nodes that are reachable by
- temporal paths of length $k+1$.
-By construction, the time stamp of an active node is always not later than
-that of its descendant. Suppose we
-start with node $(1, t_1)$ in Figure \ref{fig:eg_shortest_path}.
-After $1$ temporal path from $(1, t_1)$, we can reach
-$(2, t_1)$ and $(1, t_2)$. After $2$ temporal paths, we can reach
-$(2, t_1), (1, t_2), (2, t_2)$ and $(3, t_2)$.
-% This example motivates us to define the reachable set of
-% a node-timestamp pair.
-Formally, we define the \emph{forward-neighbors} of a node $(v, t)$ to
-be the set of all the reachable nodes after $1$ temporal path from
-$(v, t)$, i.e., the nodes pointed by $v$ at time stamp $t$ and
-the active node $v$ itself at later time stamps. In Figure
-\ref{fig:eg_shortest_path},
+This definition of a temporal path differs from that of the dynamic walk
+in \cite{gphe11,grihig13} in its explicit requirement of containing only active
+nodes. The definition implies that if either or both end points of a temporal
+path are inactive, then the entire temporal path must be the empty sequence
+$\langle \rangle$. Keeping track explicitly of the time labels of each temporal
+node allows greater generality to cases where the node sets change over time.
+Furthermore, we shall show later in
+%XXX X-REF
+that the explicit bookkeeping of the time labels is essential for correctly
+generalizing the BFS to evolving graphs.
+
+The following definition of \textbf{forward neighbors} generalizes the notion of
+neighbors and reachability in static graphs.
+
+\begin{defn}
+The \textbf{$k$-forward neighbors} of a temporal node $(v, t)$ are the temporal nodes
+that are the $k+1$st temporal node in every temporal path of length $k+1$ starting
+from $(v, t)$. The \textbf{forward neighbors} of a temporal node $(v, t)$ are its
+ \textbf{1-forward neighbors}.
+\end{defn}
+
+In Figure~\ref{fig:eg_shortest_path},
 the forward-neighbors of $(1, t_1)$ are $(2, t_1)$ and $(1, t_2)$ and
 the forward-neighbors of $(2, t_1)$ is $(2, t_3)$.
-
-Given an evolving graph $G_n$ and a starting active node $(s,t)$, Algorithm \ref{alg:bfs}
-returns all the reachable active nodes and their distances from $(s,t)$.
+The 2-forward neighbors of $(1, t_1)$ are
+$(2, t_1), (1, t_2), (2, t_2)$ and $(3, t_2)$.
+By construction, time stamp of every forward neighbor of an active node $(v, t)$
+must be no earlier than $t$.
+
+\begin{defn}
+The distance from a temporal node $(v, t)$ to a temporal node $(w, s)$ is
+the $k$ for which $(w, s)$ is a $k$-forward neighbor of $(v, t)$.
+\end{defn}
+
+Note that this notion of distance is not a metric, since the distance from $(v, t)$
+to $(w, s)$ will in general differ from the distance of $(v, t)$ from $(w, s)$.
+
+\begin{defn}
+A temporal node $(w, s)$ is \textbf{reachable} from a temporal node $(v, t)$ if
+the distance to $(w, s)$ from a temporal node $(v, t)$ there exists some finite
+integer $k$ for which $(w, s)$ is a $k$-forward neighbor of $(v, t)$.
+\end{defn}
+
+The BFS on evolving graphs is described in Algorithm \ref{alg:bfs}.
+Given an evolving graph $G_n$ and a root $(s,t)$,
+Algorithm~\ref{alg:bfs} returns all temporal notes reachable from the root and
+their distances from the root.
+The BFS constructs a tree inductively by discovering all $k$-forward neighbors
+of the root before proceeding to all $(k+1)$-forward neighbors of the root.
+
 For Figure \ref{fig:eg_shortest_path}, suppose we start with $(1, t_2)$, then
@@
 }
- \caption{Evolving-graph BFS.
-Let $G_n$ be an
-evolving graph represented by adjacency lists and $(s, t)$ is a node
+ \caption{Breadth-first search (BFS) on an evolving graph: given
+$G_n$, an evolving graph represented by adjacency lists, and a root $(s, t)$,
+
+ is a node
 of $G_n$.

### 2016-01-23 paper.tex
Start to redescribe BFS

@@
 A \textbf{temporal path} of length $m$ on an evolving graph $G_n$
-from temporal node $(v_1, t_1)$ to temporal node $(v_m, t_m)$} is
+from temporal node $(v_1, t_1)$ to temporal node $(v_m, t_m)$ is
 a time-ordered sequence of active nodes,
@@
 
+For each temporal node $(v_i,t_i)$, we use a dictionary $dict.(v_i,t_i)$
+to store its distance from $(s, t)$. We also use
+two FIFO queues $frontier$ and $next$, where $frontier$
+stores the frontier nodes of the current discovered regions and $next$ stores
+the frontier nodes of the next level.
+
 \begin{algorithm}[h]
@@
 \caption{Breadth-first search (BFS) on an evolving graph: given
-$G_n$, an evolving graph represented by adjacency lists, and a root $(s, t)$,
-
- is a node
-of $G_n$.
-For each node-time stamp pair $(v_i,t_i)$,
-we use a dictionary $dict.(v_i,t_i)$ to store its distance (the length of temporal
-paths) from $(s, t)$. We also use
-two first-in, first-out queue $frontier$ and $next$, where $frontier$
-stores the frontier nodes of the current discovered regions and $next$ stores
-the frontier nodes of the next level.
+$G_n$, an evolving graph represented by adjacency lists, and a root $(s, t)$.
 The algorithm returns $dict$ which stores all reachable node-time stamp pairs from $(s,t)$ such that if $dict.(v_i, t_i) = k$,
 the length of the shortest temporal path
-from $(s,t)$ to $(v_i, t_i)$ is $k$.}
+from $(s,t)$ to $(v_i, t_i)$ is $k$.
+}
 \label{alg:bfs}

### 2016-01-23 paper.tex
add adjacency matrices subsection

@@
 
+\subsection{Temporal Paths and Adjacency Matrices}
+\label{sec:temp-paths-adjac}
+
+%walks on graphs
+Let $A$ be the corresponding adjacency matrix of a graph $G$. 
+It is a well known result that the $(i,j)$ entry of $A^k$ counts the number of 
+walks of length exactly $k$ between node $i$ and node $j$. 
+It is natural to consider a generalization of this relationship to 
+the case of evolving graphs. 
+One may conclude that  the $(i,j)$ entry of the sum of all the product 
+\begin{equation}
+\label{eq:sum}
+A^{[1]}A^{[t_i]}A^{[t_j]} \cdots A^{t_k]}A^{[n]},  \quad t_i \leq t_j \leq \cdots \leq t_k,
+\end{equation}
+counts the number of temporal paths from $(i,1)$ to $(j,n)$.
+However, if we look at it carefully. We observe some temporal paths are 
+missed. 
+In Figure \ref{fig:eg_shortest_path}, the $(1,3)$ entry of the sum of the product
+\[
+A^{[1]}A^{[2]}A^{[3]} + A^{[1]}A^{[3]}
+\]
+is equal to $1$ even though
+there are two temporal paths from $(1, t_1)$ to $(3, t_3)$ as shown in Figure \ref{fig:active}.
+% Indeed, Grindrod et al. \cite{gphe11} observe
+% that the $(i,j)$ entry of $A^{[1]}A^{[2]}\cdots A^{[n]}$ counts the number of 
+% temporal paths of length $n$ from $(i, 1)$ to $(j, n)$, which
+% go through exactly one edge at each time stamp between 
+% $t_1$ and $t_n$. 
+% Intuitively, it seems that the $(i,j)$ entry of 
+% the summation of all the products of the form
+% counts the number of all temporal paths between $i$ and $j$ at all 
+% time stamps.  
+% However, this is not true since not all temporal paths go through
+% exactly one edge at each time stamp. 
+% In Figure , the summation of all the products of adjacency matrices
+% $A^{[1]}$, $A^{[2]}$, and $A^{[3]}$ 
+% \[
+% A^{[1]}A^{[2]}A^{[3]}  = 0
+% \]
+% The $(1,3)$ entry is equal to $1$.  But there are two temporal
+% paths from $(1, t_1)$ to $(3, t_3)$ as shown in Figure \ref{fig:active}. 
+% missing components call it E
+% For adjacency matrices $A^{[s]}$ and $A^{[t]}$, where $s < t$, a nonzero entry
+% $(i,j)$ of $A^{[s]} A^{[t]}$ indicates a temporal path from
+ $(i,s)$ to $(j,t)$ via
+% some node $k$ (which is unknown).
+% In general, a nonzero entry $(i,j)$ of the matrix $A^{[1]}A^{[2]}\cdots A^{[n]}$
+% indicates a temporal path $p$ that starts at $(i,1)$ and ends at  $(j,n)$.
+% Notice the fact that $p$ must go through at least one edge at each time stamp between $1$ and $n$.
+Notice $A^{[1]}A^{[2]} = 0$ since 
+there is no temporal path from time stamp $t_1$ to $t_2$ that
+go through at least one edge at $t_1$. However,
+\[
+\langle(1, t_1), (1, t_2), (3,t_2)\rangle
+\]
+is a temporal path as shown in Figure \ref{fig:active}, suggesting we miss to count a temporal path from $(1, t_1)$ to $(3, t_3)$.
+Notice also that if we introduce ones on the diagonal of the matrices $\{A^{[i]}\}$,
+each node of the resulting evolving graph will become
+active at all timestamps.
+Hence matrix-matrix multiplication between consecutive
+adjacency matrices can not discover all possible temporal paths and should be avoided
+when temporal paths are the essential parts of the algorithm.
+
 Node activeness is the key concept for defining temporal paths which

### 2016-01-23 paper.tex
Clean up description of Alg 1

@@
 is not a temporal path because node 2 is inactive at time $t_2$.
-Similarly,
 
@@
 from $(v, t)$. The \textbf{forward neighbors} of a temporal node $(v, t)$ are its
- \textbf{1-forward neighbors}.
+1-forward neighbors.
 \end{defn}
@@
 In Figure~\ref{fig:eg_shortest_path},
-the forward-neighbors of $(1, t_1)$ are $(2, t_1)$ and $(1, t_2)$ and
-the forward-neighbors of $(2, t_1)$ is $(2, t_3)$.
+the forward neighbors of $(1, t_1)$ are $(2, t_1)$ and $(1, t_2)$ and
+the forward neighbors of $(2, t_1)$ is $(2, t_3)$.
 The 2-forward neighbors of $(1, t_1)$ are
@@
 Note that this notion of distance is not a metric, since the distance from $(v, t)$
-to $(w, s)$ will in general differ from the distance of $(v, t)$ from $(w, s)$.
+to $(w, s)$ will in general differ from the distance of $(v, t)$ from $(w, s)$
+owing to time ordering.
 
@@
 
-The BFS on evolving graphs is described in Algorithm \ref{alg:bfs}.
-Given an evolving graph $G_n$ and a root $(s,t)$,
+The BFS on evolving graphs is described in Algorithm~\ref{alg:bfs}.
+Given an evolving graph $G_n$ and a root $(v_1, t_1)$,
 Algorithm~\ref{alg:bfs} returns all temporal notes reachable from the root and
 their distances from the root.
+$reached$ is a dictionary from temporal nodes to integers whose key set represents
+all visited temporal nodes and whose value set are the corresponding distances
+from the root.
+
 The BFS constructs a tree inductively by discovering all $k$-forward neighbors
 of the root before proceeding to all $(k+1)$-forward neighbors of the root.
-
-For Figure \ref{fig:eg_shortest_path}, suppose we start with $(1, t_2)$, then
-$frontier$ is $\langle(1,t_2)\rangle$ after step $5$.  Since the only forward-neighbor of
-$(1,t_2)$ is $(3,t_2)$, $dict.(3,t_2) =1$ at step $11$ and $next$ is enqueued to be
-$\langle (3,t_2) \rangle$ at step $12$. After that $frontier$ is equal to $next$. Since
-$frontier$ is not empty, we go through step $7$-$12$ again. This time
-the forward-neighbor of $(3,t_2)$ is $(3,t_3)$, so $dict.(3,t_3)=2$ and
-$next = \langle (3,t_3) \rangle$ and so on. The loop terminates when $frontier$ is empty.
-
-For each temporal node $(v_i,t_i)$, we use a dictionary $dict.(v_i,t_i)$
-to store its distance from $(s, t)$. We also use
-two FIFO queues $frontier$ and $next$, where $frontier$
-stores the frontier nodes of the current discovered regions and $next$ stores
-the frontier nodes of the next level.
+Within the outermost loop, the algorithm iterates over $frontier$, a list of all
+temporal nodes of distance $k$ from the root. The $nextfrontier$ list is populated
+with all temporal nodes that are forward neighbors of any temporal node in the
+$frontier$ list which have not yet been reached by the algorithm.
+
+As a simple example, consider the BFS on the example graph in
+Figure~\ref{fig:eg_shortest_path} starting from the root $(1, t_2)$.
+The $frontier$ list is first initialized to $\{(1,t_2)\}$.
+Since the only forward neighbor of $(1,t_2)$ is $(3,t_2)$,
+iteration $k=1$ produces $reached[(3,t_2)]=1$ and $nextfrontier = \{(3,t_2)\}$.
+In the next iteration $k=2$, the only forward neighbor of $(3,t_2)$ is $(3,t_3)$,
+so $reached[(3,t_3)]=2$ and $nextfrontier = \{(3,t_3)\}$ and so on.
+The algorithm terminates after $k=3$ after verifying that $(3,t_3)$ has no
+forward neighbors.
 
@@
  \SetKwFor{If}{if}{}{end}
- \SetKwFunction{nei}{forward-neighbors}
- \SetKwFunction{enqueue}{enqueue}
+ \SetKwFunction{nei}{forwardneighbors}
  \SetKwProg{Fn}{function}{}{end}
- \Fn{BFS($G_n, (s,t)$)}{
-  $dict.(s,t) = 0$ \\
-  $i = 1$ \\
-  $fronter = \emptyset$ \\
-  \enqueue{$frontier, (s,t)$} \\
+ \Fn{BFS($G_n, (v_0,t_0)$)}{
+  $reached[(v_0,t_0)] = 0$ \\
+  $frontier = \{(v_0,t_0)\}$ \\
+  $k = 1$ \\
   \While{$frontier \ne \emptyset$}{
-   $next = \emptyset$ \\
-   \For{$(u,t') \in frontier$}{
-     \For{$(v,t'') \in$\nei{$(u,t')$}}{
-          \If{$(v,t'') \notin d$}{
-              $dict.(v,t'') = i$\\
-              \enqueue{$next, (v,t'')$}
-              }
-         }
-       }
-  $frontier = next$ \\
-  $i = i+1$ \\
-  }
-  \textbf{return} $dict$
-}
-\caption{Breadth-first search (BFS) on an evolving graph: given
-$G_n$, an evolving graph represented by adjacency lists, and a root $(s, t)$.
-The algorithm returns $dict$ which stores all reachable node-time stamp pairs from $(s,t)$ such that if $dict.(v_i, t_i) = k$,
-the length of the shortest temporal path
-from $(s,t)$ to $(v_i, t_i)$ is $k$.
+   $nextfrontier = \emptyset$ \\
+   \For{$(v,t) \in frontier$}{
+     \For{$(v',t') \in$\nei{$(v,t)$}}{
+          \If{$(v',t') \notin d$}{
+              $reached[(v',t')] = k$\\
+              $nextfrontier = nextfrontier \cup \{(v',t')\}$ \\
+              }
+         }
+       }
+  $frontier = nextfrontier$ \\
+  $k = k+1$ \\
+  }
+  \textbf{return} $reached$
+}
+\caption{Breadth-first search (BFS) on an evolving graph $G_n$ starting from a
+root $(v_1, t_1)$. The return value, $reached$, is a dictionary mapping all
+reachable temporal notes from the root to their distances from the root. At the
+end of each iteration $k$, the $frontier$ set contains all temporal nodes of
+distance $k$ fro the root.
 }

### 2016-01-23 paper.tex
Fix fencepost error in root
Reorganize the analysis section

@@
  \SetKwProg{Fn}{function}{}{end}
- \Fn{BFS($G_n, (v_0,t_0)$)}{
-  $reached[(v_0,t_0)] = 0$ \\
-  $frontier = \{(v_0,t_0)\}$ \\
+ \Fn{BFS($G_n, (v_1,t_1)$)}{
+  $reached[(v_1,t_1)] = 0$ \\
+  $frontier = \{(v_1,t_1)\}$ \\
   $k = 1$ \\
@@
 end of each iteration $k$, the $frontier$ set contains all temporal nodes of
-distance $k$ fro the root.
+distance $k$ from the root.
 }
@@
 
-%Also need to consider the operations for matrix lists.
-
-\subsection{Algebraic Evolving-Graph BFS}
+\begin{thm}[Correctness of the evolving graph BFS]
+Algorithm~\ref{alg:bfs} discovers every active node that is reachable from the
+root $(v_1, t_1)$, and $reached[(v, t)]$ is the distance from $(v_1, t_1)$ to
+$(v, t)$.
+\end{thm}
+
+\proof
+Define a static graph $G=(V,E)$ with node set $V = $V^{[1]}$ $\{t\}, t\in\{t_1, t_2, \dots, t_n\}$  such that We could map an evolving graph $G_n$ to a static graph $G$ in such a way that
+if there is an edge from $(v, t)$ to $(u,t)$ in $G_n$, there is an edge from $v_t$ to $u_t$ in $G$.
+For any two nodes $v_s$ and $v_t$ in $G$, if $s < t$, we add an edge from $v_s$ to $v_t$.
+By construction, an active node $(v,t)$ in $G_n$ is in one-to-one correspondence with
+a node $v_t$ in $G$.
+The forward-neighbors of an active node $(v, t)$ in $G_n$
+are in one-to-one correspondence with the forward-neighbors of a node $v_t$ in $G$.
+Hence the correctness of the evolving graph BFS follows from that of a static graph.
+\endproof
+
+\subsection{Analysis}
+\label{sec:analysis}
+
+As presented, the BFS over evolving graphs makes no assumptions about how the% adjacency matrices
+Suppose $G_n$ is represented by adjacency lists.
+Then the number of adjacency lists is the same as the number
+of active nodes in $G_n$. So for large $n$, this representation
+can use a large amount of memory.
+For the graph $G^{[i]} = (V^{[i]}, E^{[i]})$ at time stamp $i$,
+we denote the active node set of $V^{[i]}$ as $\tilde V^{[i]}$,
+where we have $\tilde V^{[i]} \subseteq V^{[i]}$.
+Step $10$ of  Algorithm \ref{alg:bfs} ensures each active node-time stamp pair in $G_n$ is enqueued in $next$ (and then $frontier$) at most once.
+And so \texttt{forward-neighbors} loops through each
+active node-time stamp pair at most once.
+The operations of enqueuing take $O(1)$ time.
+Thus the total operations of enqueuing are
+$O(\sum_{i=1}^n|\tilde V^{[i]}|)$, where $|\cdot|$ represents the number of elements of a set.  Let $\sum_i |\tilde V^{[i]}| = |\tilde V|$ and $\sum_i |E^{[i]}| = |\hat E|$.
+Then the time of scanning
+\texttt{forward-neighbors} is
+\[
+\sum_{i=1}^n\sum_{v \in \tilde V^{[i]}} \texttt{forward-neighbors}(v) =
+O(|\hat E|),
+\]
+by the definition of forward-neighbors.
+Hence the total running time is $O(|\hat E|+ |\tilde V|)$.
+
+\section{Formulating the Evolving Graph BFS with Linear Algebra}
 \label{sec:representation}
@@
 
-\subsection{Analysis}
-\label{sec:analysis}
-
-% adjacency matrices
-Suppose $G_n$ is represented by adjacency lists.
-Then the number of adjacency lists is the same as the number
-of active nodes in $G_n$. So for large $n$, this representation
-can use a large amount of memory.
-For the graph $G^{[i]} = (V^{[i]}, E^{[i]})$ at time stamp $i$,
-we denote the active node set of $V^{[i]}$ as $\tilde V^{[i]}$,
-where we have $\tilde V^{[i]} \subseteq V^{[i]}$.
-Step $10$ of  Algorithm \ref{alg:bfs} ensures each active node-time stamp pair in $G_n$ is enqueued in $next$ (and then $frontier$) at most once.
-And so \texttt{forward-neighbors} loops through each
-active node-time stamp pair at most once.
-The operations of enqueuing take $O(1)$ time.
-Thus the total operations of enqueuing are
-$O(\sum_{i=1}^n|\tilde V^{[i]}|)$, where $|\cdot|$ represents the number of elements of a set.  Let $\sum_i |\tilde V^{[i]}| = |\tilde V|$ and $\sum_i |E^{[i]}| = |\hat E|$.
-Then the time of scanning
-\texttt{forward-neighbors} is
-\[
-\sum_{i=1}^n\sum_{v \in \tilde V^{[i]}} \texttt{forward-neighbors}(v) =
-O(|\hat E|),
-\]
-by the definition of forward-neighbors.
-Hence the total running time is $O(|\hat E|+ |\tilde V|)$.
-
-\begin{thm}[Correctness of evolving-graph BFS]
-Let an evolving graph $G_n$ and a starting node $(s,t_0)$ be given.
-The evolving-graph BFS discovers every active node that is reachable from the source $(s,t_0)$ and $dict.(v, t)$ is the length of the shortest temporal path
-from $(s,t_0)$ to $(v,t)$.
-\end{thm}
-\proof
-We could map an evolving graph $G_n$ to a static graph $G$ in such a way that
-if there is an edge from $(v, t)$ to $(u,t)$ in $G_n$, there is an edge from $v_t$ to $u_t$ in $G$.
-For any two nodes $v_s$ and $v_t$ in $G$,  if $s < t$, we add an edge from $v_s$ to $v_t$.
-By construction, an active node $(v,t)$ in $G_n$ is in one-to-one correspondence with
-a node $v_t$ in $G$.
-The forward-neighbors of an active node $(v, t)$ in $G_n$
-are in one-to-one correspondence with the forward-neighbors of a node $v_t$ in $G$.
-Hence the correctness of the evolving graph BFS follows from that of a static graph.
-\endproof
-
 \begin{thm}
@@
 \endproof
-
-\subsection{Search Tree}
-\label{sec:short-temp-path}
-
-We say two nodes $(v_1,t_1)$ and $(v2, t_2)$ are \emph{temporally connected} if
-$(v_1, t_1)$ and $(v2,t_2)$ are connected by a temporal path.
-Notice this notion is symmetric, reflexive but not transitive. For example, suppose
-$t_1 < t_2 < t_3$ and $(v_1, t_1)$ and $(v_3, t_3)$ are temporally connected
-by a temporal path from $(v_1,t_1)$ to $(v_3, t_3)$. Also
-$(v_3, t_3)$ and $(v_2, t_2)$ are temporally connected by a temporal path from
-$(v_2,t_2)$ to $(v_3, t_3)$. But this does not imply $(v_1, t_1)$ and $(v_2, t_2)$ are temporal connected.
-
-Given a root node $(s, t)$, during the execution the evolving-graph BFS builds a search tree, denoted by $T(s,t)$,
-where nodes in $T(s,t)$ are temporally connected.
-\begin{lemma}
-For a node active at different time stamps, say $(s, t_1)$ and $(s,t_2)$ such that $t_1 \leq t_2$,  we have $T(s, t_2) \subseteq T(s, t_1)$, where $T(s,t_1) = T(s,t_2)$ if and only if
-$t_1 = t_2$.
-\end{lemma}
-\proof
-Trivially, $t_1 =t_2$ if and only if $(s,t_1) = (s,t_2)$ if and only if
-$T(s,t_1) = T(s,t_2)$. Now suppose $t_1 < t_2$.
-Let $(v, t)$ be an arbitrary active node that belongs to $T(s, t_2)$. Then there exists a temporal path
-(constructed by the evolving graph BFS) from $(s,t_2)$ to $(v,t)$. Since both
-$(s,t_1)$ and $(s,t_2)$ are active, $(s,t_2)$ is a forward-neighbor of $(s,t_1)$ by definition, i.e., there exists an edge from $(s,t_1)$ to $(s,t_2)$.  Hence there exists a temporal path
-from $(s,t_1)$ to $(v,t)$ and so $(v,t) \in T(s,t_1)$.
-\endproof
-
-% consider $(A+I)^Tb$, which records all the connected components.
-% consider the huge matrix block and how to relate it.
-
-% \subsection{Online Searching}
-% \label{sec:online-searching}
-
-% In this section, we consider ways of  updating new graphs. Suppose $A_n $ is
-% an evolving graph with $n$ timestamps. We received a new graph $A^{[n+1]}$
-% at timestamp $n+1$. How to update the tree components while minimize the computation?
-% The forward-neighbors at timestamp $$
 
@@
 
+% \subsection{Search Tree}
+% \label{sec:short-temp-path}
+%
+% We say two nodes $(v_1,t_1)$ and $(v2, t_2)$ are \emph{temporally connected} if
+% $(v_1, t_1)$ and $(v2,t_2)$ are connected by a temporal path.
+% Notice this notion is symmetric, reflexive but not transitive. For example, suppose
+% $t_1 < t_2 < t_3$ and $(v_1, t_1)$ and $(v_3, t_3)$ are temporally connected
+% by a temporal path from $(v_1,t_1)$ to $(v_3, t_3)$. Also
+% $(v_3, t_3)$ and $(v_2, t_2)$ are temporally connected by a temporal path from
+% $(v_2,t_2)$ to $(v_3, t_3)$. But this does not imply $(v_1, t_1)$ and $(v_2, t_2)$ are temporal connected.
+%
+% Given a root node $(s, t)$, during the execution the evolving-graph BFS builds a search tree, denoted by $T(s,t)$,
+% where nodes in $T(s,t)$ are temporally connected.
+% \begin{lemma}
+% For a node active at different time stamps, say $(s, t_1)$ and $(s,t_2)$ such that $t_1 \leq t_2$,  we have $T(s, t_2) \subseteq T(s, t_1)$, where $T(s,t_1) = T(s,t_2)$ if and only if
+% $t_1 = t_2$.
+% \end{lemma}
+% \proof
+% Trivially, $t_1 =t_2$ if and only if $(s,t_1) = (s,t_2)$ if and only if
+% $T(s,t_1) = T(s,t_2)$. Now suppose $t_1 < t_2$.
+% Let $(v, t)$ be an arbitrary active node that belongs to $T(s, t_2)$. Then there exists a temporal path
+% (constructed by the evolving graph BFS) from $(s,t_2)$ to $(v,t)$. Since both
+% $(s,t_1)$ and $(s,t_2)$ are active, $(s,t_2)$ is a forward-neighbor of $(s,t_1)$ by definition, i.e., there exists an edge from $(s,t_1)$ to $(s,t_2)$.  Hence there exists a temporal path
+% from $(s,t_1)$ to $(v,t)$ and so $(v,t) \in T(s,t_1)$.
+% \endproof
+
+% consider $(A+I)^Tb$, which records all the connected components.
+% consider the huge matrix block and how to relate it.
+
+% \subsection{Online Searching}
+% \label{sec:online-searching}
+
+% In this section, we consider ways of  updating new graphs. Suppose $A_n $ is
+% an evolving graph with $n$ timestamps. We received a new graph $A^{[n+1]}$
+% at timestamp $n+1$. How to update the tree components while minimize the computation?
+% The forward-neighbors at timestamp $$
+
 \section{Conclusion}

### 2016-01-23 paper.tex
Provide proper proof of Theorem 1

@@
 \begin{thm}[Correctness of the evolving graph BFS]
-Algorithm~\ref{alg:bfs} discovers every active node that is reachable from the
-root $(v_1, t_1)$, and $reached[(v, t)]$ is the distance from $(v_1, t_1)$ to
-$(v, t)$.
+Let $G_n$ be an evolving graph and $(v_1, t_1)$ be an active node of $G_n$.
+Then Algorithm~\ref{alg:bfs} discovers every active node that is reachable from
+the root $(v_1, t_1)$, and $reached[(v, t)]$ is the distance from $(v_1, t_1)$
+to $(v, t)$.
 \end{thm}
 
-\proof
-Define a static graph $G=(V,E)$ with node set $V = $V^{[1]}$ $\{t\}, t\in\{t_1, t_2, \dots, t_n\}$  such that We could map an evolving graph $G_n$ to a static graph $G$ in such a way that
-if there is an edge from $(v, t)$ to $(u,t)$ in $G_n$, there is an edge from $v_t$ to $u_t$ in $G$.
-For any two nodes $v_s$ and $v_t$ in $G$, if $s < t$, we add an edge from $v_s$ to $v_t$.
-By construction, an active node $(v,t)$ in $G_n$ is in one-to-one correspondence with
-a node $v_t$ in $G$.
-The forward-neighbors of an active node $(v, t)$ in $G_n$
-are in one-to-one correspondence with the forward-neighbors of a node $v_t$ in $G$.
-Hence the correctness of the evolving graph BFS follows from that of a static graph.
+\IEEEproof
+Define the set of temporal nodes
+$\tilde V_L^{[t]} = \{(v_1)_t = (v_1, t) | (v_1, v_2) \in E^{[t]}\}$, which
+consists of the active nodes at time $t$ which participate on the left side of an edge.
+Similarly, $\tilde V_R^{[t]} = \{(v_2)_t = (v_2, t) | (v_1, v_2) \in E^{[t]}\}$
+are the corresponding active nodes on the right side of an edge. Then
+$\tilde V^{[t]} = \tilde V_L^{[t]} \cup \tilde V_R^{[t]}$ is the set of active
+nodes at time $t$, and $V = \bigcup_t \tilde V^{[t]}$ is the set of all
+active nodes in $G_n$.
+
+Similarly, define an edge set $E' = \{(u_s,v_t)|u_s=(u,s)\in V,
+v_t=(v,t)\in V, v=u, s<t\}$, which consists of temporal nodes that connect
+active nodes sharing the same node at different times. Each edge in $E'$ is then
+in 1-1 correspondence with a temporal path of length 2,
+$\langle (v,s), (v,t) \rangle$.
+Define also $\tilde E^{[t]} = \{(e, t)|e\in E^{[t]}\}$, which are simply the edge
+sets in $G_n$ with time labels. Then $E = \bigcup_t \tilde E^{[t]} \cup E'$ is
+the set of all edges representing all allowed temporal paths of length 2.
+
+The node set $V$ and edge set $E$ now define a static graph $G=(V,E)$ that is in
+1-1 correspondence with the evolving graph $G_n$. The node set $V$ of $G$ is in
+1-1 correspondence with active nodes of $G_n$ while the edge set $E$ is in 1-1
+correspondence with all temporal paths of length 2 on $G_n$.
+
+We now establish a similar 1-1 correspondence of forward neighbors of an active
+node with a subset of $G$.
+By definition, the forward neighbors of some active node $(v, t) \in G_n$ are
+active nodes of either the form $(v, t')$ for some $t' > t$ or $(u, t)$ for some
+$u\ne v$. Clearly, the former are elements of $E' \subseteq E$ while the latter
+are elements of $E^{[t]} \subseteq \bigcup_s \tilde E^{[s]} \subseteq E$.
+Thus each forward neighbor of an active node $(v, t) \in G_n$ is in 1-1
+correspondence with a node in $V$ that is a neighbor of $v_t \in V$.
+
+The correctness of BFS on the evolving graph $G_n$ now follows from the
+correctness of BFS on the static graph $G$, since we have also established a 1-1
+correspondence for every intermediate quantity in Algorithm~\ref{alg:bfs}.
 \endproof
@@
 \end{thm}
-\proof
+
+\IEEEproof
 We shall check the two algorithms are equivalent in each module of

### 2016-01-23 paper.tex
Simplify complexity analysis
Given all the trouble we went through to prove Theorem 1, the same
machinery now gives us the complexity essentially for free.

@@
 to $(v, t)$.
+\label{thm:correct}
 \end{thm}
@@
 
-\subsection{Analysis}
-\label{sec:analysis}
-
-As presented, the BFS over evolving graphs makes no assumptions about how the% adjacency matrices
-Suppose $G_n$ is represented by adjacency lists.
-Then the number of adjacency lists is the same as the number
-of active nodes in $G_n$. So for large $n$, this representation
-can use a large amount of memory.
-For the graph $G^{[i]} = (V^{[i]}, E^{[i]})$ at time stamp $i$,
-we denote the active node set of $V^{[i]}$ as $\tilde V^{[i]}$,
-where we have $\tilde V^{[i]} \subseteq V^{[i]}$.
-Step $10$ of  Algorithm \ref{alg:bfs} ensures each active node-time stamp pair in $G_n$ is enqueued in $next$ (and then $frontier$) at most once.
-And so \texttt{forward-neighbors} loops through each
-active node-time stamp pair at most once.
-The operations of enqueuing take $O(1)$ time.
-Thus the total operations of enqueuing are
-$O(\sum_{i=1}^n|\tilde V^{[i]}|)$, where $|\cdot|$ represents the number of elements of a set.  Let $\sum_i |\tilde V^{[i]}| = |\tilde V|$ and $\sum_i |E^{[i]}| = |\hat E|$.
-Then the time of scanning
-\texttt{forward-neighbors} is
-\[
-\sum_{i=1}^n\sum_{v \in \tilde V^{[i]}} \texttt{forward-neighbors}(v) =
-O(|\hat E|),
-\]
-by the definition of forward-neighbors.
-Hence the total running time is $O(|\hat E|+ |\tilde V|)$.
+As presented, the BFS over evolving graphs makes no assumptions about how the
+evolving graph $G_n$ is represented. Suppose it is represented by a collection of
+adjacency lists, one for each active node in $G_n$. Then we have that the
+asympotic complexity of BFS on $G_n$ is the same as that for BFS on $G$, using
+the 1-1 construction of $G$ from $G_n$.
+
+\begin{thm}[Complexity of the evolving graph BFS]
+Let $G_n$ be an evolving graph represented using adjacency lists,
+$(v_1, t_1)$ be an active node of $G_n$, and $G=(V,E)$ be the static graph
+constructed from $G_n$ using the 1-1 correspondences defined in the proof of
+Theorem~\ref{thm:correct}. Then the asymptotic complexity of
+Algorithm~\ref{alg:bfs} is $O(|E| + |V|)$.
+\end{thm}
+
+\IEEEproof
+Any edge in any edge set of $G_n$ can be accessed in constant time in random
+access memory. By construction, BFS on $G_n$ is in 1-1 correspondence with BFS
+on the static graph $G$. The number of operations of BFS on $G$ is $O(|E| + |V|)$,
+and so the result follows.
+\endproof
 

### 2016-01-23 paper.tex
Clean up presentation of algebraic BFS

@@
 
+The preceding example illustrates the fact that $G^{[1]}$ plays no part in the
+BFS traversal of $G_n$ starting from $(1, t_2)$. In general, all $G^{[t]}$ with
+time stamps $t < t'$ for a starting node $(v, t')$ are irrelevant to the BFS
+traversal. Hence without loss of generality we may assume that BFS is always
+computed with a root at time $t_1$, the earliest time stamp in $G_n$.
+
 \begin{algorithm}[h]
@@
 \label{sec:representation}
-As before, let $G_n = \langle G^{[1]}, G^{[2]}, \ldots, G^{[n]}\rangle$ be an evolving graph with $n$ time stamps. For each graph $G^{[t]}$, we use $A^{[t]}$ to denote
-the corresponding adjacency matrix.
+
+For each static graph $G^{[t]} = (V^{[t]}, E^{[t]})$ that constitutes the
+evolving graph $G_n$, define its corresponding
+$\left\vert V^{[t]}\right\vert \times \left\vert V^{[t]}\right\vert$
+adjacency matrix with elements
 \[
-A^{[t]}(i,j) =
+A^{[t]}_{ij} =
 \begin{cases}
- 1 & \mbox{if $(i,j)$ is an edge of $G^{[t]}$,} \\
+ 1 & \mbox{if $(i,j)\in E^{[t]}$,} \\
  0 & \mbox{otherwise.}
@@
 \]
- Each $A^{[t]}$ can be represented by
-a $|V|\times |V|$ sparse matrix,
-where $|V|$ is the total number of nodes in $G_n$.
-We use a sequence of adjacency matrices $A_n = \langle A^{[1]}, A^{[2]}, \ldots, A^{n]}\rangle$ to represent $G_n$.
-For example, the evolving graph in Figure \ref{fig:eg_shortest_path}
-can be represented as
+
+We can then represent $G_n$ using the sequence of adjacency matrices
+$A_n = \langle A^{[1]}, A^{[2]}, \ldots, A^{n]}\rangle$. The example in
+Figure~\ref{fig:eg_shortest_path} can be represented as
 \[
@@
  \end{bmatrix}
-\right\rangle
+\right\rangle.
 \]
 
-This algebraic formulation exploits a graphical interpretation of matrix--vector
-products involving the adjacency matrix~\cite{kegi11}.
-Let $G = (V, E)$ be a (directed or undirected) graph
-and $k$ be a node in the graph $G$. If $A$ is the adjacency matrix of $G$ and
-$b$ is the column vector
+This algebraic representation allows us to exploit a graphical interpretation of
+matrix--vector products involving the adjacency matrix~\cite{kegi11}.
+If $A$ is the adjacency matrix of a (static) graph $G$ and $e_k$ is $k$th
+elementary unit vector, then the nonzero entries of $A^T e_k$ have indices that
+are neighbors of $k$.
+The algebraic formulation of BFS on evolving graphs follows similarly, but
+requires a new kind of matrix--vector product, $\odot$, defined by
 \[
-b_i =
-\begin{cases}
-1 & \mbox{if $i = k$,} \\
-0 & \mbox{otherwise,}
-\end{cases}
-\]
-then $A^Tb$ is a column vector whose nonzero entries have indices that are the
-forward-neighbors of $k$.
-
-The evolving graph BFS can be thought of as a generalization of blocked
-matrix--vector products.
-Introduce a new operator $\odot$, a variation on the ordinary matrix--vector
-product such that
-\[
-A^T \odot b =
+A^T \!\odot b =
 \begin{cases}
@@
 \]
-Then, the forward-neighbors of node $(k, 1)$ in $A_n = \langle A^{[1]}, A^{[2]}, \ldots, A^{n]}\rangle$ are the nonzero indices of the
-elements in the sequence
+The forward neighbors of a temporal node $(k, t_1)$ in $A_n$ are then be determined
+from the indices and time stamps of the nonzero elements in the sequence
 \begin{equation}
 \label{eq:out_ne}
-\big\langle (A^{[1]})^Tb, \quad  (A^{[2]})^T\odot b, \quad \ldots \quad (A^{[n]})^T\odot b \big\rangle.
+\big\langle (A^{[1]})^Te_k, \; (A^{[2]})^T\!\odot e_k, \; \ldots \; (A^{[n]})^T\!\odot e_k\big\rangle.
 \end{equation}
-The forward-neighbors at time stamp $t$ are the nonzero indices of the $t$-th element of the above sequence.
-Hence given $A_n$ and a vector $b$, we can use \eqref{eq:out_ne} to
-get a $1$-step BFS tree. In Figure \ref{fig:eg_shortest_path}, the forward-neighbors
-of node $(1, t_1)$ can be computed by
-
+The nonzero entries of the first vector represent forward neighbors that are on
+the same time stamp $t_1$, whereas all other nonzero entries represent forward
+neighbors that are advanced in time but remain on the same node $k$.
+The quantity \eqref{eq:out_ne} therefore encodes a BFS tree of depth 2,
+as its nonzero entries are labeled by all temporal nodes of distance 1 from
+$(k, t_1)$.
+
+Referring back to the example of Figure~\ref{fig:eg_shortest_path}, the forward
+neighbors of node $(1, t_1)$ can be computed by
 \begin{align*}
@@
 0
-\end{bmatrix}
-,
+\end{bmatrix}\!,
  \begin{bmatrix}
@@
 0
-\end{bmatrix}
+\end{bmatrix}\!
 \right\rangle
@@
 \end{align*}
-This means node $(2,t_1)$ and $(1,t_2)$ are the forward-neighbors of $(1,t_1)$.
+From this computation, we can deduce that $(2,t_1)$ and $(1,t_2)$ are the forward
+neighbors of $(1,t_1)$.
 

### 2016-01-23 paper.tex
Simply discussion of backward time travel

@@
 \end{algorithm}
-
-Representing evolving graphs as sparse matrices offers great flexibility
-on the search direction. If we define the \emph{backward-neighbors} of a
-node $(v,t)$ to be a set of nodes that precede $(v,t)$ if we travel backwards
-in time. Then the backward-neighbors of $(v,t)$ can be computed by
-\begin{equation}
-\label{eq:backward}
-(A^{[t]})b + (A^{[t+1]})\odot b + \cdots + (A^{[n]})\odot b.
-\end{equation}
-We shall get a backward traversal algorithm if we replace
-the vector-matrix multiplication $\bm{A_n}^T\bm{b}$
-by $\bm{A_n}\bm{b}$, where
-\[
-\bm{A_n} =
-\begin{bmatrix}
-A^{[1]} &   0            & \cdots  &  & 0 \\
-\sigma(A^{[2]}) & A^{[2]} & 0         &\cdots  & \vdots \\
-\sigma(A^{[3]}) & \sigma(A^{[3]}) & A^{[3]} & \cdots \\
-   \vdots              &                              &      \vdots            &  &  \vdots \\
-\sigma(A^{[n]}) &                  &       \cdots                      &  & A^{[n]}
-\end{bmatrix}.
-\]
-Having the ability to traverse backwards is important for many real-life applications.
-We shall show an example in Section \ref{sec:applications}.
 
@@
 
+Tracing influence through citation networks naturally lead to considering paths
+through evolving graphs that traverse time backwards. While we have until now
+only considered forward evolution in time, the treatment for reverse evolution
+follows naturally simply by reversing the time labels, e.g.\ by the transformation
+$t\rightarrrow -t$.
+
+% Representing evolving graphs as sparse matrices offers great flexibility
+% on the search direction. If we define the \emph{backward-neighbors} of a
+% node $(v,t)$ to be a set of nodes that precede $(v,t)$ if we travel backwards
+% in time. Then the backward-neighbors of $(v,t)$ can be computed by
+% \begin{equation}
+% \label{eq:backward}
+% (A^{[t]})b + (A^{[t+1]})\odot b + \cdots + (A^{[n]})\odot b.
+% \end{equation}
+% We shall get a backward traversal algorithm if we replace
+% the vector-matrix multiplication $\bm{A_n}^T\bm{b}$
+% by $\bm{A_n}\bm{b}$, where
+% \[
+% \bm{A_n} =
+% \begin{bmatrix}
+% A^{[1]} &   0            & \cdots  &  & 0 \\
+% \sigma(A^{[2]}) & A^{[2]} & 0         &\cdots  & \vdots \\
+% \sigma(A^{[3]}) & \sigma(A^{[3]}) & A^{[3]} & \cdots \\
+%    \vdots              &                              &      \vdots            &  &  \vdots \\
+% \sigma(A^{[n]}) &                  &       \cdots                      &  & A^{[n]}
+% \end{bmatrix}.
+% \]
+% Having the ability to traverse backwards is important for many real-life applications.
+% We shall show an example in Section \ref{sec:applications}.
+
 % \subsection{Search Tree}

### 2016-01-23 paper.tex
Simplify description of implementation.
Take out future work from conclusion.

Better acks.

@@
 
-To study evolving graphs and experiment with various graph types, we have developed EvolvingGraphs.jl  \cite{zhang15}, a software package for the creation, manipulation, and study of evolving graphs written in Julia \cite{bkse12}.
-The implementation of the evolving-graph BFS is
-available online at
+To study evolving graphs and experiment with various graph types, we have
+developed EvolvingGraphs.jl~\cite{zhang15}, a software package for the creation,
+manipulation, and study of evolving graphs written in Julia~\cite{bkse12}.
+The implementation of the evolving graph BFS is
+freely available online at
 \begin{center}
-\url{https://github.com/weijianzhang/EvolvingGraphs.jl}.
+\url{https://github.com/weijianzhang/EvolvingGraphs.jl}
 \end{center}
+and available with the MIT "Expat" license.
+
 %Julia is a high-level, high-performance dynamic programming language for technical
@@
 as a sequence of sparse matrices stored in compressed sparse column (CSC) format.
-Many basic graph functions are implemented for \texttt{MatrixList}.
+Many basic graph functions are also implemented for \texttt{MatrixList}.
 
@@
 
-We shall map a database of research papers to an evolving graph $G_n$
-in such a way that if author $i$ cited author $j$ on a paper published
-at  time $t$, the evolving graph $G_n$ contains a directed edge from
-$i$ to $j$ at time stamp $t$. Then given an author $a$ at time stamp $t_0$,
-we can use our evolving-graph BFS to construct $T(a, t_0)$ which is
-a set of all the authors that have been influenced by $a$'s work at
-time $t_0$. We define a \emph{community} to be a group of
+In this section, we show that evolving graph formalism presented above can be
+used to capture the structure of citation networks. Consider
+the evolving graph $G_n=\langle G^{[t]} \rangle_t$ such that $G^{[t]}$ has
+node set corresponding to authors active at time $t$ and directed edge set
+$E^{[t]} \ni (i, j)$ representing a citation of author $j$ by author $i$ in a
+publication at time $t$.
+
+Then given an author $a$ at time $t_1$, the evolving graph BFS described above
+can compute $T(a, t_1)$, the set of all the authors that have been influenced by
+$a$'s work at time $t_1$.
+Define also a \emph{community} to be a group of
 researchers that have been influenced by the same authors.
-For example, given a paper published by $a$ at year $2016$,
-we can determine $a$'s community by first backward searching to
-find the root authors then forward searching to find the descendants.
-
-To design such information retrieval system, two parameters will be needed:
- one for determining the layers of the backward search tree which can be
-adjusted by the user;
-the other for determining root authors for forward searching
-so that we can
-filter out unessential or overlapped trees, which can be computed
-by ranking algorithms like PageRank \cite{pbmw99} or machine-learning based preference ranking.
-
-Tracing influence through citation networks naturally lead to considering paths
-through evolving graphs that traverse time backwards. While we have until now
-only considered forward evolution in time, the treatment for reverse evolution
-follows naturally simply by reversing the time labels, e.g.\ by the transformation
-$t\rightarrrow -t$.
+For example, given a paper published by $a$ at time $t$,
+we can determine $a$'s community by searching backward in time to find
+$T^-1(a, t_1)$, the authors that influenced $a$ at time $t$, and also searching
+forward to find $T(a, t_1)$.
+The backward search in time follows straightforwardly from the forward time traveral
+presented above simply by reversing the time labels, e.g.\ by the transformation
+$t\rightarrow -t$.
 
@@
 
-In this paper, we have introduced and analyzed the evolving-graph BFS in both
-graph and algebraic notation. The relationship between active nodes and
-evolving graph algorithms is discussed.
-This paper leaves important extensions open for further research. First, it is
-of interest to extend the evolving-graph BFS algorithm for weighted evolving graphs, which can provide more valuable information when the degree of connectivity is available in the data.
-Second, efficient algebraic representations and algorithms need to be addressed.
-One major advantage of representing evolving graphs as sparse matrices
-is the matrix representation is highly structured so it is easy to design
-high performance parallel implementation.
-Third, depth-first search and other graph traversal algorithms can be
-extended using similar ideas.
-
-\section*{Acknowledgment}
-
-The authors are grateful to Professor Nicholas J. Higham and Doctor Vedran \u Sego for suggestions.
-This work was completed while the second author visited MIT CSAIL. He thanks
-Professor Alan Edelman for making his stay so pleasant.
+This paper introduced the generalization of BFS to evolving graphs and presented
+an equivalent formulation using linear algebra. The concept temporal paths
+connecting active nodes turns out to be critical for the correct generalization
+of BFS to evolving graphs. An efficient implementation of BFS on evolving
+graphs is now available as a open source Julia package.
+
+% This paper leaves important extensions open for further research. First, extensions
+% of BFS to weighted evolving graphs can provide more valuable information when the degree of connectivity is available in the data.
+% Second, efficient algebraic representations and algorithms need to be addressed.
+% One major advantage of representing evolving graphs as sparse matrices
+% is the matrix representation is highly structured so it is easy to design
+% high performance parallel implementation.
+% Third, depth-first search and other graph traversal algorithms can be
+% extended using similar ideas.
+
+\section*{Acknowledgments}
+
+We thank N. J. Higham and V. \u Sego (Manchester) for helpful suggestions.
+W.\ Z.\ thanks the School of Mathematics at U. Manchester for travel funding and
+A. Edelman for arranging for a fruitful visit to MIT.
+
+%XXX ANY OTHER FUNDING
 

### 2016-01-23 paper.tex
Clean up presentation of linear algebra

@@
 
+\subsection{Evolving graphs as sequences of adjacency matrices}
+
 For each static graph $G^{[t]} = (V^{[t]}, E^{[t]})$ that constitutes the
@@
 matrix--vector products involving the adjacency matrix~\cite{kegi11}.
-If $A$ is the adjacency matrix of a (static) graph $G$ and $e_k$ is $k$th
+If $A$ is the adjacency matrix of a (static) graph $G$ and $e_k$ is the $k$th
 elementary unit vector, then the nonzero entries of $A^T e_k$ have indices that
@@
 \]
-The forward neighbors of a temporal node $(k, t_1)$ in $A_n$ are then be determined
+The forward neighbors of a temporal node $(k, t_1)$ in $A_n$ can then be determined
 from the indices and time stamps of the nonzero elements in the sequence
@@
 
-Let $(A^{[i]})^Tb = c_i$.
-At the second step, we need to compute the following components
+\subsection{Evolving graphs as a blocked adjacency matrix}
+
+The proof of Theorem~\ref{thm:correct} provided a construction for representing
+an evolving graph $G_n$ by a static graph $G$ with nodes corresponding to active
+nodes of $G_n$. The adjacency matrix $\mathbf A$ of $G$ then has block structure
+which is useful for understanding the nature of the $\odot$ operation.
+
+Consider two iterations of the BFS on $G_n$ with root $(k, t_1)$. The second
+iteration requires computing the sequences
+
+\begin{subequations}
 \begin{align}
@@
 \end{align}
-If we sum these resulting vectors according to time stamps, we obtain
-the forward-neighbors of the nodes computed at step $1$.
-For $n =2$, we have
-\[
-\Big\langle
-(A^{[1]})^Tc_1,~
-(A^{[2]})^T\odot c_1 + (A^{[2]})^T c_2
-\Big\rangle.
-\]
-If we form a block triangular matrix $\bm{A_n}^T$
+\end{subequations}
+%
+where $c_i = (A^{[i]})^Te_k$. Summing resultant vectors that share the same
+time stamp, we obtain vectors whose nonzero elements have indexes labeled by
+the forward neighbors of the nodes computed at step $1$.
+
+Define the block triangular adjacency matrix
 \[
@@
 \]
-where  $\sigma(A^{[i]})^Tb = (A^{[i]})^T\odot b$, and a block vector $\bm{b}$
+where $\sigma(A^{[i]})^Tb = (A^{[i]})^T\odot b$, and the block vector
 \[
-\bm{b}^T = [b^T, 0, \cdots, 0  ],
+\bm{b}^T = [b^T, 0, \cdots, 0  ].
 \]
-then \eqref{eq:out_ne} is equivalent to $\bm{A_n}^T\bm{b}$ and
-summing \eqref{eq:5}-\eqref{eq:6} is equivalent to $\bm{A_n}^T(\bm{A_n}^T\bm{b})$.
-In general, we can define an evolving-graph BFS using
-block vector-matrix multiplications. The algorithm is presented in Algorithm
-\ref{alg:bfsla}.
+Then, \eqref{eq:out_ne} is equivalent to the matrix--vector product
+$\bm{A_n}^T\bm{b}$ and the summation over \eqref{eq:5}-\eqref{eq:6} is equivalent
+to $\bm{A_n}^T(\bm{A_n}^T\bm{b})$.
+
+\subsection{The algebraic formulation of BFS on evolving graphs}
+
+The blocked matrix--vector products introduced in the previous section allows us
+to write down an elegant algebraic formulation of BFS on evolving graphs, as
+presented in Algorithm~\ref{alg:bfsla}. Note that $\bm A^T$ need never be formed
+explicitly, as only its matrix--vector product is required in the algorithm.
 
@@
  \SetKwProg{Fn}{function}{}{end}
- \Fn{ABFS($A_n, (s,t)$)}{
+ \Fn{ABFS($A_n, (v_1,t_1)$)}{
    Form $\bm{A}^T$ from $A_n$.\\
-   Let $m$ be the dimension of  $A^{[1]}$. \\
-   $b$ is a sparse vector of length $mn$ \\
-   $b(s+mt) =1$ \\
-   $i = 1$ \\
-   $dict.(s,t) = 0$ \\
+   $\bm b = e_{v_1}$ \\
+   $k = 1$ \\
+   $reached[(v_1,t_1)] = 0$ \\
    \While{\nz{$b$} $\ne \emptyset$}{
        $\bm{b} = \bm{A}^T\bm{b}$ \\
-       \For{$i$ in \nz{$b$}}{
-         $node =$ \map{$i$} \\
-         \If{$node$ in $dict$}{
-           $b(i) = 0$ \\
+       \For{$i \in$ \nz{$b$}}{
+         \If{\map{$k$} $\in reached$}{
+           $b_k = 0$ \\
          }
@@
       $nodes =$\map{$b$}\\
-       \For{$node$ in $nodes$}{
-           $dict.node = i$
-         }
-
-       $i = i + 1$
-       }
-  \textbf{return} $dict$ \\
-}
- \caption{Algebraic evolving-graph BFS.
-Let $A_n$ be the corresponding adjacency matrices of $G_n$ and
-$(s,t)$ a node of $G_n$. As in Algorithm \ref{alg:bfs}, we use a dictionary $dict.(v_i, t_i)$
-to store $(v_i, t_i)$'s distance from $(s,t)$. Function \texttt{nonzeros(v)}
-returns the nonzero indices of a given vector  $v$ and function \texttt{map(b)}
-maps a block vector's indices to their corresponding nodes. The algorithm  returns
-$dict$ which stores all the reachable active nodes and their distance from $(s,t)$.}
+       \For{$node \in reached$}{
+           $reached[node] = k$ \\
+         }
+       $k = k + 1$
+       }
+  \textbf{return} $reached$ \\
+}
+\caption{An algebraic formulation of BFS on evolving graphs.
+Given $A_n$, the adjacency matrix representation of $G_n$ and $(v_1,t_1)$, a
+node of $G_n$, returns $reached$ as defined in Algorithm~\ref{alg:bfs}.
+The function \texttt{nonzeros(v)} returns the nonzero indices of the vector $v$,
+and the function \texttt{map(b)} maps a block vector's indices to their corresponding
+nodes.}
 \label{alg:bfsla}
 \end{algorithm}
-
-Algorithm \ref{alg:bfsla} outlines the basic idea of an algebraic evolving-graph BFS.
-But in practice we don't need to form $\bm{A}$ explicitly at step $2$. By realizing the dependency between these sparse matrices,
-we can compute  the result in parallel.
 
@@
 \IEEEproof
-We shall check the two algorithms are equivalent in each module of
-the method.
 The initialization steps are trivially equivalent.
-By the construction of $\bm{A^T}$,
-step $9$ of Algorithm \ref{alg:bfsla} computes the forward-neighbors
-of all the frontier nodes, which is equivalent to step $8$-$9$
-of Algorithm \ref{alg:bfs}. Step $10$-$15$ of Algorithm \ref{alg:bfsla} ensures
-active nodes will not be visited twice, which is equivalent to step $10$
-of Algorithm \ref{alg:bfs}. Finally, a dictionary $dict$ is used by both algorithms
-to keep record of the distance of each node from the source $(s,t)$.
+At the beginning of iteration $k$, the block vector $\bm b$ represents the
+frontier nodes encoding the $frontier$ set of Algorithm~\ref{alg:bfs}.
+The matrix--vector product $\bm A^T \bm b$ encodes the forward neighbors of all
+the frontier nodes. Subequently, active nodes that have already been visited in
+previous iterations are zeroed out of the new $\bm b$.
 \endproof

### 2016-01-23 paper.tex
update algebraic BFS algorithm

@@
    Form $\bm{A}^T$ from $A_n$.\\
-   $\bm b = e_{v_1}$ \\
+   $\bm{b}$ is a sparse vector of length $mn$. \\
+   $\bm{b}_s = 1$ \\
    $k = 1$ \\
    $reached[(v_1,t_1)] = 0$ \\
-   \While{\nz{$b$} $\ne \emptyset$}{
+   \While{\nz{$\bm{b}$} $\ne \emptyset$}{
        $\bm{b} = \bm{A}^T\bm{b}$ \\
-       \For{$i \in$ \nz{$b$}}{
+       \For{$k \in$ \nz{$\bm{b}$}}{
          \If{\map{$k$} $\in reached$}{
-           $b_k = 0$ \\
-         }
-      }
-      $nodes =$\map{$b$}\\
-       \For{$node \in reached$}{
-           $reached[node] = k$ \\
+           $\bm{b}_k = 0$ \\
+         }
+      }
+      $activeNodes =$\map{$\bm{b}$}\\
+       \For{$activeNode \in activeNodes$}{
+           $reached[activeNode] = k$ \\
          }
@@
 The function \texttt{nonzeros(v)} returns the nonzero indices of the vector $v$,
-and the function \texttt{map(b)} maps a block vector's indices to their corresponding
-nodes.}
+and the function \texttt{map(b)} maps a block vector's index (or indices) to their corresponding
+active node (or active nodes).}
 \label{alg:bfsla}

### 2016-01-23 paper.tex
Simplify description of adjacency matrix chains

@@
 
+\subsection{Not All Temporal Paths Can Be Expressed by Products of Adjacency Matrices}
+\label{sec:temp-paths-adjac}
+
+For a static graph $G$ with adjacency matrix $A$, $(A^k)_{ij}$ counts the number
+of paths of length $k$ between node $i$ and node $j$. Na\"ively, one might want
+to generalize this result to evolving graphs by postulating that the $(i,j)$th
+entry of the discrete path sum
+\begin{align}
+S^{[t_n]} = A^{[t_1]}A^{[t_n]} + \sum_{t_1 \le t \le t_n} A^{[t_1]}A^{[t]}A^{[t_n]}
++ \dots \nonumber \\
++ \sum_{t_1 \le t \le t' \le \dots \le t_n}
+A^{[t_1]}A^{[t]}A^{[t']}\cdots A^{[t_n]}
+\label{eq:sum}
+\end{align}
+counts the number of temporal paths from $(i,t_1)$ to $(j,t_n)$.
+However, this postulate is incorrect.
+In the example of Figure~\ref{fig:eg_shortest_path},
+\[
+(S^{[t_3]})_{13} = \left(A^{[t_1]}A^{[t_2]}A^{[t_3]} + A^{[t_1]}A^{[t_3]}\right)_{13} = 1
+\]
+even though there are clearly two temporal paths from
+$(1, t_1)$ to $(3, t_3)$ as shown in Figure~\ref{fig:active}.
+% Indeed, Grindrod et al. \cite{gphe11} observe
+% that the $(i,j)$ entry of $A^{[1]}A^{[2]}\cdots A^{[n]}$ counts the number of
+% temporal paths of length $n$ from $(i, 1)$ to $(j, n)$, which
+% go through exactly one edge at each time stamp between
+% $t_1$ and $t_n$.
+% Intuitively, it seems that the $(i,j)$ entry of
+% the summation of all the products of the form
+% counts the number of all temporal paths between $i$ and $j$ at all
+% time stamps.
+% However, this is not true since not all temporal paths go through
+% exactly one edge at each time stamp.
+% In Figure , the summation of all the products of adjacency matrices
+% $A^{[1]}$, $A^{[2]}$, and $A^{[3]}$
+% \[
+% A^{[1]}A^{[2]}A^{[3]}  = 0
+% \]
+% The $(1,3)$ entry is equal to $1$.  But there are two temporal
+% paths from $(1, t_1)$ to $(3, t_3)$ as shown in Figure \ref{fig:active}.
+% missing components call it E
+% For adjacency matrices $A^{[s]}$ and $A^{[t]}$, where $s < t$, a nonzero entry
+% $(i,j)$ of $A^{[s]} A^{[t]}$ indicates a temporal path from
+% $(i,s)$ to $(j,t)$ via
+% some node $k$ (which is unknown).
+% In general, a nonzero entry $(i,j)$ of the matrix $A^{[1]}A^{[2]}\cdots A^{[n]}$
+% indicates a temporal path $p$ that starts at $(i,1)$ and ends at  $(j,n)$.
+% Notice the fact that $p$ must go through at least one edge at each time stamp between $1$ and $n$.
+
+The first term in the sum $S^{[t_3]}$ vanishes since $A^{[t_1]}A^{[t_2]} = 0$.
+Furthermore, the vanishing of $S^{[t_2]} = A^{[t_1]}A^{[t_2]}$ itself reflects the absence of any temporal path
+from $t_1$ to $t_2$ that goes through at least one edge at $t_1$. However,
+\begin{equation}
+\langle(1, t_1), (1, t_2), (3, t_2)\rangle~\label{eq:tpathex}
+\end{equation}
+is a clearly a valid temporal path as shown in Figure~\ref{fig:active} which
+cannot be expressed by a product of adjacency matrices.
+
+Sums $S^{[t]}$ of the form \eqref{eq:sum} produce an incorrect count of temporal
+paths because they do not capture temporal paths with subpaths of the form
+$\langle (v, s), (v, t) \rangle$, $s < t$. The temporal path \eqref{eq:tpathex}
+is counted only by the matrix product $M^{[t_1]} A^{[t_2]}$, where
+\[
+    M^{[t_1]} = \begin{pmatrix}1 & 0 & 0 \\ 0 & 1 & 0 \\ 0 & 0 & 0\end{pmatrix}.
+\]
+$M^{[t_1]}$ describes the forward time propagation of active nodes. It counts
+temporal paths with subpaths $\langle (i, t_1), (i, t_2) \rangle$
+for all active nodes $(i, t_1)$ at time $t_1$, but not paths containing the
+subsequence $\langle (3, t_1), (3, t_2) \rangle$, which cannot exist in temporal
+paths since $(3, t_1)$ is an inactive node.
+
+One might attempt to amend the sums $S^{[t]}$ in \eqref{eq:sum} by redefining
+the adjacency matrices to include ones along the diagonal. However, the resulting
+sum is still incorrect, as it counts paths with subsequences $\langle (3, t_1), (3, t_2) \rangle$
+and are hence not temporal paths.
+
+The simple example of Figure~\ref{fig:eg_shortest_path} provides several
+counterexamples which demonstrate why sums over products of adjacency matrices of
+the form \eqref{eq:sum} do not count temporal paths correctly. The interpretation
+from considering powers of $A$ for static graphs therefore does not extend to
+evolving graphs, and sums like \eqref{eq:sum} cannot form the basis of correct
+analyses of evolving graphs.
+
+% But it provides no information about the nodes in between. In addition, the temporal path must pass at least
+%  one node at each timestamps. We believe this requirement is too restrictive. For example,
+% $A^{[t_1]}A^{[t_2]}A^{[t_3]}$ is a zero matrix. Let $\langle i,j,\ldots k \rangle$ represents the sets of all possible subsets of $\langle 1, 2, \ldots, n \rangle$, where $i<j<\cdots<k$, then
+% the nonzero entries of the matrix
+% \[
+% \sum_{i,j,\ldots,k}A^{[1]}A^{[i]}A^{[j]}\cdots A^{[k]}A^{[n]},
+% \]
+% indicates temporal paths  from temporal paths from timestamp $1$ to $n$.
+
 \subsection{Evolving graphs as sequences of adjacency matrices}
@@
 
-\section{Node Activeness and its Impact}
-\label{sec:matr-matr-mult}
-
-% \subsection{Aggregated Graphs}
-% \label{sec:aggregated-graphs}
-
-% % In this section, we answer the question why we need evolving graph.
-% %
-% At the beginning of the paper, we set up a game played by three people $A$, $B$,
-% and $C$ and we showed if $A$ and $B$
-
-\subsection{Temporal Paths and Adjacency Matrices}
-\label{sec:temp-paths-adjac}
-
-%walks on graphs
-Let $A$ be the corresponding adjacency matrix of a graph $G$.
-It is a well known result that the $(i,j)$ entry of $A^k$ counts the number of
-walks of length exactly $k$ between node $i$ and node $j$.
-It is natural to consider a generalization of this relationship to
-the case of evolving graphs.
-One may conclude that  the $(i,j)$ entry of the sum of all the product
-\begin{equation}
-\label{eq:sum}
-A^{[1]}A^{[t_i]}A^{[t_j]} \cdots A^{t_k]}A^{[n]},  \quad t_i \leq t_j \leq \cdots \leq t_k,
-\end{equation}
-counts the number of temporal paths from $(i,1)$ to $(j,n)$.
-However, if we look at it carefully. We observe some temporal paths are
-missed.
-In Figure \ref{fig:eg_shortest_path}, the $(1,3)$ entry of the sum of the product
-\[
-A^{[1]}A^{[2]}A^{[3]} + A^{[1]}A^{[3]}
-\]
-is equal to $1$ even though
-there are two temporal paths from $(1, t_1)$ to $(3, t_3)$ as shown in Figure \ref{fig:active}.
-% Indeed, Grindrod et al. \cite{gphe11} observe
-% that the $(i,j)$ entry of $A^{[1]}A^{[2]}\cdots A^{[n]}$ counts the number of
-% temporal paths of length $n$ from $(i, 1)$ to $(j, n)$, which
-% go through exactly one edge at each time stamp between
-% $t_1$ and $t_n$.
-% Intuitively, it seems that the $(i,j)$ entry of
-% the summation of all the products of the form
-% counts the number of all temporal paths between $i$ and $j$ at all
-% time stamps.
-% However, this is not true since not all temporal paths go through
-% exactly one edge at each time stamp.
-% In Figure , the summation of all the products of adjacency matrices
-% $A^{[1]}$, $A^{[2]}$, and $A^{[3]}$
-% \[
-% A^{[1]}A^{[2]}A^{[3]}  = 0
-% \]
-% The $(1,3)$ entry is equal to $1$.  But there are two temporal
-% paths from $(1, t_1)$ to $(3, t_3)$ as shown in Figure \ref{fig:active}.
-% missing components call it E
-% For adjacency matrices $A^{[s]}$ and $A^{[t]}$, where $s < t$, a nonzero entry
-% $(i,j)$ of $A^{[s]} A^{[t]}$ indicates a temporal path from
- $(i,s)$ to $(j,t)$ via
-% some node $k$ (which is unknown).
-% In general, a nonzero entry $(i,j)$ of the matrix $A^{[1]}A^{[2]}\cdots A^{[n]}$
-% indicates a temporal path $p$ that starts at $(i,1)$ and ends at  $(j,n)$.
-% Notice the fact that $p$ must go through at least one edge at each time stamp between $1$ and $n$.
-Notice $A^{[1]}A^{[2]} = 0$ since
-there is no temporal path from time stamp $t_1$ to $t_2$ that
-go through at least one edge at $t_1$. However,
-\[
-\langle(1, t_1), (1, t_2), (3,t_2)\rangle
-\]
-is a temporal path as shown in Figure \ref{fig:active}, suggesting we miss to count a temporal path from $(1, t_1)$ to $(3, t_3)$.
-Notice also that if we introduce ones on the diagonal of the matrices $\{A^{[i]}\}$,
-each node of the resulting evolving graph will become
-active at all timestamps.
-Hence matrix-matrix multiplication between consecutive
-adjacency matrices can not discover all possible temporal paths and should be avoided
-when temporal paths are the essential parts of the algorithm.
-
-Node activeness is the key concept for defining temporal paths which
-shapes how to traverse an evolving graph.
-Recall an active node $(v,t)$ is a node $v$ at timestamp $t$
-that connects to at least another node other that itself.
-It means the entity at that specific timestamp either
-receives recommendations or gives recommendations to others (or possibly both).
-This implies the entity is ``alive''. Here ``alive'' does not necessary mean physically
-alive. For the case of citation network that we shall discuss later, node activeness more or less means spiritually alive in the sense that the idea or knowledge of the entity is alive.
-We traverse an evolving graph only on the active nodes.
-
-For adjacency matrices $A^{[s]}$ and $A^{[t]}$, where $s < t$, a nonzero entry
-$(i,j)$ of $A^{[s]} A^{[t]}$ indicates a temporal path from $(i,s)$ to $(j,t)$ via
-some node $k$ (which is unknown).
-In general, a nonzero entry $(i,j)$ of the matrix $A^{[1]}A^{[2]}\cdots A^{[n]}$
-indicates a temporal path $p$ that starts at $(i,1)$ and ends at  $(j,n)$.
-Notice the fact that $p$ must go through at least one edge at each time stamp between $1$ and $n$.
-For the evolving graph in Figure \ref{fig:eg_shortest_path}, $A^{[1]}A^{[2]}$
-is a zero matrix since there is no temporal path from time stamp $t_1$ to $t_2$ that
-go through at least one edge at $t_1$. However,
-$\langle(1, t_1), (1, t_2), (3,t_2)\rangle$ is a temporal path as shown in Figure \ref{fig:active}.
-Notice also that if we introduce ones on the diagonal of the matrices $\{A^{[i]}\}$,
-each node of the resulting evolving graph will become
-active at all timestamps.
-Hence matrix-matrix multiplication between consecutive
-adjacency matrices can not discover all possible temporal paths and should be avoided
-when temporal paths are the essential parts of the algorithm.
-
-% But it provides no information about the nodes in between. In addition, the temporal path must pass at least
-%  one node at each timestamps. We believe this requirement is too restrictive. For example,
-% $A^{[t_1]}A^{[t_2]}A^{[t_3]}$ is a zero matrix. Let $\langle i,j,\ldots k \rangle$ represents the sets of all possible subsets of $\langle 1, 2, \ldots, n \rangle$, where $i<j<\cdots<k$, then
-% the nonzero entries of the matrix
-% \[
-% \sum_{i,j,\ldots,k}A^{[1]}A^{[i]}A^{[j]}\cdots A^{[k]}A^{[n]},
-% \]
-% indicates temporal paths  from temporal paths from timestamp $1$ to $n$.
-
 \section{Implementation in Julia}
@@
 \end{center}
-and available with the MIT "Expat" license.
+and available with the MIT ``Expat'' license.
 

### 2016-01-23 paper.tex
Fix up typos

@@
 We can use analyze the spread of information between the players using graph
-theory. In this process, the time ordering of events matters, which leads us to
-consider a graph that changes over time, i.e.\ a graph $G(t) = (V(t), E(t))$ whose
-vertex set $V$ and edge set $E$ evolve over on time, $t$.
+theory. In this process, the time ordering of events matters, and hence its
+graph representation $G(t) = (V(t), E(t))$ must be time dependent.
 Such a graph is called an ``evolving graph''~\cite{fkp89,bkmu12},
@@
 requires paths that also move monotonically forward in time, thus necessating
-a generalization of the ordinary BFS algorithm. An informal description of such
-a generalization was provided by \cite{tmml09}.
-
+a generalization of the ordinary BFS algorithm.
+
+An informal description of BFS generalized to evolving graphs can be found in
+\cite{tmml09}. However, it turns out that the proper generalization involves
+paths of nontrivial structure which sometimes contradict na\"ive generalizations
+of results derived from static graphs.
 In this paper, we present a complete description of the BFS algorithm for evolving graphs.
@@
 \subsection{Temporal Paths over Active Nodes}
+\label{sec:temporal-paths}
 
@@
 influence of nodes over social networks. We will also show later in
-%XXX FIX X-REF TO SECTION
+Section~\ref{sec:temp-paths-adjac}
 that the resulting structure of allowable temporal paths leads to nontrivial
@@
 
-The example presented above in Section~\ref{sec:breadth-first-search}
+The example presented above in Section~\ref{sec:temporal-paths}
 demonstrates how the notion of active nodes restrict the set of temporal paths
@@
 node allows greater generality to cases where the node sets change over time.
-Furthermore, we shall show later in
-%XXX X-REF
+Furthermore, we shall show later in \ref{sec:temp-paths-adjac}
 that the explicit bookkeeping of the time labels is essential for correctly

### 2016-01-23 paper.tex
Better title and abstract

@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{An Algorithm for Searching Evolving Graphs in Linear Algebraic Language}
+\title{Breadth First Searches Over Evolving Graphs}
 
@@
 \IEEEauthorblockN{Jiahao Chen}
-\IEEEauthorblockA{ Computer Science and Artificial Intelligence Laboratory,\\
+\IEEEauthorblockA{Computer Science and Artificial Intelligence Laboratory,\\
 Massachusetts Institute of Technology,\\
-Massachusetts, 02139-4307,USA\\
+Cambridge, Massachusetts, 02139-4307, USA\\
 jiahao@mit.edu}
@@
 University of Manchester,\\
-Manchester, M13 9PL, England\\
+Manchester, M13 9PL, England, UK\\
 weijian.zhang@manchester.ac.uk}
@@
 \begin{abstract}
-  % A generalization of the breadth-first search (BFS) algorithm on evolving graphs is
-  % discussed.
-  % The representation of this algorithm
-  % as algebraic operations is investigated.
-  % The duality of the vector-matrix multiplication
-  % and a BFS step is extended to the fact that a block vector-matrix multiplication
-  % is dual with the evolving-graph BFS.
-  % It is found that the algebraic representation allows for flexible forward and backward searching and can be used for mining citation networks.
-A generalized breadth-first search (BFS) algorithm is presented for
-searching evolving graphs.
-The method is described in both graph and linear algebra notation and
-the two algorithms are shown to be equivalent.
-A preliminary discussion on mining citation networks is also presented.
+We present a generalization of the breadth first search (BFS) algorithm on
+evolving graphs and prove its correctness. The BFS generates paths with nontrivial
+structure, which we term temporal paths. We also show with simple counterexamples
+that simple products of adjacency matrices do not have the correct combinatorial
+properties of enumerating temporal paths. On the contrary, the correct generalization
+of BFS requires careful tracking of active nodes which participate in the definition
+of edges at specific times. We discuss how the structure of temporal paths leads
+to nontrivial matrix structure in adjacency matrix representations of evolving
+graphs, as well as implications for real world applications
+such as the analysis of citation networks.
 \end{abstract}

### 2016-01-23 paper.tex
add figure to illustrate BFS

--- 2016-01-23 paper.tex
@@
 \definecolor{blue2}{RGB}{235,244,250}
-\definecolor{yellow1}{RGB}{255,249,234}
+\definecolor{yellow1}{RGB}{255,255,102}
 \definecolor{blue3}{RGB}{63,40,96}
@@
 $frontier$ list which have not yet been reached by the algorithm.
+
+\begin{figure}[h]
+ \begin{center}
+    \begin{tikzpicture}[scale=.36, line width =1.4pt]
+  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {2};
+  \node[circle,draw=red1, minimum size=0.2cm] (n8) at (-10,5) {3};
+  \node[circle,fill=green1, minimum size=0.2cm] (n10) at (-8,10) {1};
+
+  \node[circle,draw=red1, minimum size=0.2cm] (n6) at (3,7) {2};
+  \node[circle,fill=green1, draw=black, minimum size=0.2cm] (n4) at (-2,5) {3};
+  \node[circle,fill=green1, draw=black, minimum size=0.2cm] (n1) at (0,10) {1};
+
+   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {2};
+  \node[circle,fill=green1, minimum size=0.2cm] (n12) at (6,5)  {3};
+  \node[circle,draw=red1, minimum size=0.2cm] (n14) at (8,10) {1};
+
+  \foreach \from/\to in {n10/n7, n11/n12}
+   \draw[every edge,->] (\from) -- (\to);
+     \draw[dotted,->](n1) -- (n4);
+  %\draw[dotted,->](n10) -- (n1);
+  %\draw[dotted,->](n4) -- (n12);
+\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
+(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{ $t_2$};
+\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
+(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $t_1$};
+\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
+(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $t_3$};
+    \end{tikzpicture}
+   \begin{tikzpicture}[scale=.36, line width =1.4pt]
+  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {2};
+  \node[circle,draw=red1, minimum size=0.2cm] (n8) at (-10,5) {3};
+  \node[circle,fill=green1, minimum size=0.2cm] (n10) at (-8,10) {1};
+
+  \node[circle,draw=red1, minimum size=0.2cm] (n6) at (3,7) {2};
+  \node[circle,fill=green1, draw=black, minimum size=0.2cm] (n4) at (-2,5) {3};
+  \node[circle,fill=green1, minimum size=0.2cm] (n1) at (0,10) {1};
+
+   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {2};
+  \node[circle,fill=green1, draw=black, minimum size=0.2cm] (n12) at (6,5)  {3};
+  \node[circle,draw=red1, minimum size=0.2cm] (n14) at (8,10) {1};
+
+  \foreach \from/\to in {n1/n4, n10/n7, n11/n12}
+   \draw[every edge,->] (\from) -- (\to);
+    % \draw[dotted,->](n7) to[out=70, in=40] (n11);
+  \draw[dotted,->](n4) -- (n12);
+  % \draw[dotted,->](n11) -- (n12);
+\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
+(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{ $t_2$};
+\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
+(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $t_1$};
+\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
+(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $t_3$};
+    \end{tikzpicture}
+   \begin{tikzpicture}[scale=.36, line width =1.4pt]
+  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {2};
+  \node[circle,draw=red1, minimum size=0.2cm] (n8) at (-10,5) {3};
+  \node[circle,fill=green1, minimum size=0.2cm] (n10) at (-8,10) {1};
+
+  \node[circle,draw=red1, minimum size=0.2cm] (n6) at (3,7) {2};
+  \node[circle,fill=green1, minimum size=0.2cm] (n4) at (-2,5) {3};
+  \node[circle,fill=green1, minimum size=0.2cm] (n1) at (0,10) {1};
+
+   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {2};
+  \node[circle,fill=green1, draw = black, minimum size=0.2cm] (n12) at (6,5)  {3};
+  \node[circle,draw=red1, minimum size=0.2cm] (n14) at (8,10) {1};
+
+  \foreach \from/\to in {n1/n4, n10/n7, n11/n12}
+   \draw[every edge,->] (\from) -- (\to);
+    % \draw[dotted,->](n7) to[out=70, in=40] (n11);
+  %\draw[dotted,->](n10) -- (n7);
+   %\draw[dotted,->](n11) -- (n12);
+\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
+(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{ $t_2$};
+\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
+(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $t_1$};
+\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
+(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $t_3$};
+    \end{tikzpicture}
+\end{center}
+\caption{An evolving graph with 3 time stamps $t_1$, $t_2$ and $t_3$.
+The circled green filled nodes indicate a BFS process starting at active node
+$(1, t_2)$.}
+\label{fig:example}
+\end{figure}
 
@@
 \label{sec:representation}
-
-\subsection{Not All Temporal Paths Can Be Expressed by Products of Adjacency Matrices}
-\label{sec:temp-paths-adjac}
-
-For a static graph $G$ with adjacency matrix $A$, $(A^k)_{ij}$ counts the number
-of paths of length $k$ between node $i$ and node $j$. Na\"ively, one might want
-to generalize this result to evolving graphs by postulating that the $(i,j)$th
-entry of the discrete path sum
-\begin{align}
-S^{[t_n]} = A^{[t_1]}A^{[t_n]} + \sum_{t_1 \le t \le t_n} A^{[t_1]}A^{[t]}A^{[t_n]}
-+ \dots \nonumber \\
-+ \sum_{t_1 \le t \le t' \le \dots \le t_n}
-A^{[t_1]}A^{[t]}A^{[t']}\cdots A^{[t_n]}
-\label{eq:sum}
-\end{align}
-counts the number of temporal paths from $(i,t_1)$ to $(j,t_n)$.
-However, this postulate is incorrect.
-In the example of Figure~\ref{fig:eg_shortest_path},
-\[
-(S^{[t_3]})_{13} = \left(A^{[t_1]}A^{[t_2]}A^{[t_3]} + A^{[t_1]}A^{[t_3]}\right)_{13} = 1
-\]
-even though there are clearly two temporal paths from
-$(1, t_1)$ to $(3, t_3)$ as shown in Figure~\ref{fig:active}.
-% Indeed, Grindrod et al. \cite{gphe11} observe
-% that the $(i,j)$ entry of $A^{[1]}A^{[2]}\cdots A^{[n]}$ counts the number of
-% temporal paths of length $n$ from $(i, 1)$ to $(j, n)$, which
-% go through exactly one edge at each time stamp between
-% $t_1$ and $t_n$.
-% Intuitively, it seems that the $(i,j)$ entry of
-% the summation of all the products of the form
-% counts the number of all temporal paths between $i$ and $j$ at all
-% time stamps.
-% However, this is not true since not all temporal paths go through
-% exactly one edge at each time stamp.
-% In Figure , the summation of all the products of adjacency matrices
-% $A^{[1]}$, $A^{[2]}$, and $A^{[3]}$
-% \[
-% A^{[1]}A^{[2]}A^{[3]}  = 0
-% \]
-% The $(1,3)$ entry is equal to $1$.  But there are two temporal
-% paths from $(1, t_1)$ to $(3, t_3)$ as shown in Figure \ref{fig:active}.
-% missing components call it E
-% For adjacency matrices $A^{[s]}$ and $A^{[t]}$, where $s < t$, a nonzero entry
-% $(i,j)$ of $A^{[s]} A^{[t]}$ indicates a temporal path from
-% $(i,s)$ to $(j,t)$ via
-% some node $k$ (which is unknown).
-% In general, a nonzero entry $(i,j)$ of the matrix $A^{[1]}A^{[2]}\cdots A^{[n]}$
-% indicates a temporal path $p$ that starts at $(i,1)$ and ends at  $(j,n)$.
-% Notice the fact that $p$ must go through at least one edge at each time stamp between $1$ and $n$.
-
-The first term in the sum $S^{[t_3]}$ vanishes since $A^{[t_1]}A^{[t_2]} = 0$.
-Furthermore, the vanishing of $S^{[t_2]} = A^{[t_1]}A^{[t_2]}$ itself reflects the absence of any temporal path
-from $t_1$ to $t_2$ that goes through at least one edge at $t_1$. However,
-\begin{equation}
-\langle(1, t_1), (1, t_2), (3, t_2)\rangle~\label{eq:tpathex}
-\end{equation}
-is a clearly a valid temporal path as shown in Figure~\ref{fig:active} which
-cannot be expressed by a product of adjacency matrices.
-
-Sums $S^{[t]}$ of the form \eqref{eq:sum} produce an incorrect count of temporal
-paths because they do not capture temporal paths with subpaths of the form
-$\langle (v, s), (v, t) \rangle$, $s < t$. The temporal path \eqref{eq:tpathex}
-is counted only by the matrix product $M^{[t_1]} A^{[t_2]}$, where
-\[
-    M^{[t_1]} = \begin{pmatrix}1 & 0 & 0 \\ 0 & 1 & 0 \\ 0 & 0 & 0\end{pmatrix}.
-\]
-$M^{[t_1]}$ describes the forward time propagation of active nodes. It counts
-temporal paths with subpaths $\langle (i, t_1), (i, t_2) \rangle$
-for all active nodes $(i, t_1)$ at time $t_1$, but not paths containing the
-subsequence $\langle (3, t_1), (3, t_2) \rangle$, which cannot exist in temporal
-paths since $(3, t_1)$ is an inactive node.
-
-One might attempt to amend the sums $S^{[t]}$ in \eqref{eq:sum} by redefining
-the adjacency matrices to include ones along the diagonal. However, the resulting
-sum is still incorrect, as it counts paths with subsequences $\langle (3, t_1), (3, t_2) \rangle$
-and are hence not temporal paths.
-
-The simple example of Figure~\ref{fig:eg_shortest_path} provides several
-counterexamples which demonstrate why sums over products of adjacency matrices of
-the form \eqref{eq:sum} do not count temporal paths correctly. The interpretation
-from considering powers of $A$ for static graphs therefore does not extend to
-evolving graphs, and sums like \eqref{eq:sum} cannot form the basis of correct
-analyses of evolving graphs.
-
-% But it provides no information about the nodes in between. In addition, the temporal path must pass at least
-%  one node at each timestamps. We believe this requirement is too restrictive. For example,
-% $A^{[t_1]}A^{[t_2]}A^{[t_3]}$ is a zero matrix. Let $\langle i,j,\ldots k \rangle$ represents the sets of all possible subsets of $\langle 1, 2, \ldots, n \rangle$, where $i<j<\cdots<k$, then
-% the nonzero entries of the matrix
-% \[
-% \sum_{i,j,\ldots,k}A^{[1]}A^{[i]}A^{[j]}\cdots A^{[k]}A^{[n]},
-% \]
-% indicates temporal paths  from temporal paths from timestamp $1$ to $n$.
 
@@
 
+\section{Node Activeness and its Impact}
+\label{sec:matr-matr-mult}
+
+% \subsection{Aggregated Graphs}
+% \label{sec:aggregated-graphs}
+
+% % In this section, we answer the question why we need evolving graph.
+% %
+% At the beginning of the paper, we set up a game played by three people $A$, $B$,
+% and $C$ and we showed if $A$ and $B$
+
+\subsection{Temporal Paths and Adjacency Matrices}
+\label{sec:temp-paths-adjac}
+
+%walks on graphs
+Let $A$ be the corresponding adjacency matrix of a graph $G$.
+It is a well known result that the $(i,j)$ entry of $A^k$ counts the number of
+walks of length exactly $k$ between node $i$ and node $j$.
+It is natural to consider a generalization of this relationship to
+the case of evolving graphs.
+One may conclude that  the $(i,j)$ entry of the sum of all the product
+\begin{equation}
+\label{eq:sum}
+A^{[1]}A^{[t_i]}A^{[t_j]} \cdots A^{t_k]}A^{[n]},  \quad t_i \leq t_j \leq \cdots \leq t_k,
+\end{equation}
+counts the number of temporal paths from $(i,1)$ to $(j,n)$.
+However, if we look at it carefully. We observe some temporal paths are
+missed.
+In Figure \ref{fig:eg_shortest_path}, the $(1,3)$ entry of the sum of the product
+\[
+A^{[1]}A^{[2]}A^{[3]} + A^{[1]}A^{[3]}
+\]
+is equal to $1$ even though
+there are two temporal paths from $(1, t_1)$ to $(3, t_3)$ as shown in Figure \ref{fig:active}.
+% Indeed, Grindrod et al. \cite{gphe11} observe
+% that the $(i,j)$ entry of $A^{[1]}A^{[2]}\cdots A^{[n]}$ counts the number of
+% temporal paths of length $n$ from $(i, 1)$ to $(j, n)$, which
+% go through exactly one edge at each time stamp between
+% $t_1$ and $t_n$.
+% Intuitively, it seems that the $(i,j)$ entry of
+% the summation of all the products of the form
+% counts the number of all temporal paths between $i$ and $j$ at all
+% time stamps.
+% However, this is not true since not all temporal paths go through
+% exactly one edge at each time stamp.
+% In Figure , the summation of all the products of adjacency matrices
+% $A^{[1]}$, $A^{[2]}$, and $A^{[3]}$
+% \[
+% A^{[1]}A^{[2]}A^{[3]}  = 0
+% \]
+% The $(1,3)$ entry is equal to $1$.  But there are two temporal
+% paths from $(1, t_1)$ to $(3, t_3)$ as shown in Figure \ref{fig:active}.
+% missing components call it E
+% For adjacency matrices $A^{[s]}$ and $A^{[t]}$, where $s < t$, a nonzero entry
+% $(i,j)$ of $A^{[s]} A^{[t]}$ indicates a temporal path from
+ $(i,s)$ to $(j,t)$ via
+% some node $k$ (which is unknown).
+% In general, a nonzero entry $(i,j)$ of the matrix $A^{[1]}A^{[2]}\cdots A^{[n]}$
+% indicates a temporal path $p$ that starts at $(i,1)$ and ends at  $(j,n)$.
+% Notice the fact that $p$ must go through at least one edge at each time stamp between $1$ and $n$.
+Notice $A^{[1]}A^{[2]} = 0$ since
+there is no temporal path from time stamp $t_1$ to $t_2$ that
+go through at least one edge at $t_1$. However,
+\[
+\langle(1, t_1), (1, t_2), (3,t_2)\rangle
+\]
+is a temporal path as shown in Figure \ref{fig:active}, suggesting we miss to count a temporal path from $(1, t_1)$ to $(3, t_3)$.
+Notice also that if we introduce ones on the diagonal of the matrices $\{A^{[i]}\}$,
+each node of the resulting evolving graph will become
+active at all timestamps.
+Hence matrix-matrix multiplication between consecutive
+adjacency matrices can not discover all possible temporal paths and should be avoided
+when temporal paths are the essential parts of the algorithm.
+
+Node activeness is the key concept for defining temporal paths which
+shapes how to traverse an evolving graph.
+Recall an active node $(v,t)$ is a node $v$ at timestamp $t$
+that connects to at least another node other that itself.
+It means the entity at that specific timestamp either
+receives recommendations or gives recommendations to others (or possibly both).
+This implies the entity is ``alive''. Here ``alive'' does not necessary mean physically
+alive. For the case of citation network that we shall discuss later, node activeness more or less means spiritually alive in the sense that the idea or knowledge of the entity is alive.
+We traverse an evolving graph only on the active nodes.
+
+For adjacency matrices $A^{[s]}$ and $A^{[t]}$, where $s < t$, a nonzero entry
+$(i,j)$ of $A^{[s]} A^{[t]}$ indicates a temporal path from $(i,s)$ to $(j,t)$ via
+some node $k$ (which is unknown).
+In general, a nonzero entry $(i,j)$ of the matrix $A^{[1]}A^{[2]}\cdots A^{[n]}$
+indicates a temporal path $p$ that starts at $(i,1)$ and ends at  $(j,n)$.
+Notice the fact that $p$ must go through at least one edge at each time stamp between $1$ and $n$.
+For the evolving graph in Figure \ref{fig:eg_shortest_path}, $A^{[1]}A^{[2]}$
+is a zero matrix since there is no temporal path from time stamp $t_1$ to $t_2$ that
+go through at least one edge at $t_1$. However,
+$\langle(1, t_1), (1, t_2), (3,t_2)\rangle$ is a temporal path as shown in Figure \ref{fig:active}.
+Notice also that if we introduce ones on the diagonal of the matrices $\{A^{[i]}\}$,
+each node of the resulting evolving graph will become
+active at all timestamps.
+Hence matrix-matrix multiplication between consecutive
+adjacency matrices can not discover all possible temporal paths and should be avoided
+when temporal paths are the essential parts of the algorithm.
+
+% But it provides no information about the nodes in between. In addition, the temporal path must pass at least
+%  one node at each timestamps. We believe this requirement is too restrictive. For example,
+% $A^{[t_1]}A^{[t_2]}A^{[t_3]}$ is a zero matrix. Let $\langle i,j,\ldots k \rangle$ represents the sets of all possible subsets of $\langle 1, 2, \ldots, n \rangle$, where $i<j<\cdots<k$, then
+% the nonzero entries of the matrix
+% \[
+% \sum_{i,j,\ldots,k}A^{[1]}A^{[i]}A^{[j]}\cdots A^{[k]}A^{[n]},
+% \]
+% indicates temporal paths  from temporal paths from timestamp $1$ to $n$.
+
 \section{Implementation in Julia}
@@
 \end{center}
-and available with the MIT ``Expat'' license.
+and available with the MIT "Expat" license.
 

### 2016-01-23 paper.tex
Rewrite abstract

@@
 \begin{abstract}
-We present a generalization of the breadth first search (BFS) algorithm on
-evolving graphs and prove its correctness. The BFS generates paths with nontrivial
-structure, which we term temporal paths. We also show with simple counterexamples
-that simple products of adjacency matrices do not have the correct combinatorial
-properties of enumerating temporal paths. On the contrary, the correct generalization
-of BFS requires careful tracking of active nodes which participate in the definition
-of edges at specific times. We discuss how the structure of temporal paths leads
-to nontrivial matrix structure in adjacency matrix representations of evolving
-graphs, as well as implications for real world applications
-such as the analysis of citation networks.
+We generalize the breadth first search (BFS) algorithm to evolving graphs and
+prove its correctness.
+The BFS generates paths with nontrivial structure, which we term temporal paths.
+We also show with simple counterexamples that products of adjacency matrices alone
+do not have the correct combinatorics of enumerating temporal paths.
+Generalizing BFS to evolving graphs requires careful tracking of active nodes
+which participate in the definition of edges at specific times.
+We discuss how the structure of temporal paths is reflected in adjacency matrix
+representations of evolving graphs,
+as well as implications for real world applications such as the analysis of citation networks.
 \end{abstract}

### 2016-01-24 paper.tex
Make paper IEEE compliant
Change citation style and delete hyperlinks - boo
add IEEEtran bib style and clearify bfs example

@@
 \makeatother
-\usepackage[colorlinks,urlcolor=blue,citecolor=hotpink,linkcolor=blue]{hyperref}
 \usetikzlibrary{decorations.pathreplacing}
@@
 
+\usepackage{hyperref}
 % *** MISC UTILITY PACKAGES ***
@@
 Figure~\ref{fig:eg_shortest_path} starting from the root $(1, t_2)$.
+The procedure is shown in Figure \ref{fig:example}.
 The $frontier$ list is first initialized to $\{(1,t_2)\}$.
@@
 In the next iteration $k=2$, the only forward neighbor of $(3,t_2)$ is $(3,t_3)$,
-so $reached[(3,t_3)]=2$ and $nextfrontier = \{(3,t_3)\}$ and so on.
+so $reached[(3,t_3)]=2$ and $nextfrontier = \{(3,t_3)\}$.
 The algorithm terminates after $k=3$ after verifying that $(3,t_3)$ has no
@@
 % http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
+\bibliographystyle{IEEEtran}
 % argument is your BibTeX string definitions and bibliography database(s)
@@
 
-\bibliographystyle{myplain2-doi}
-
 \bibliography{strings,njhigham,paper}

### 2016-01-24 paper.tex
fix typo and community searching in application

@@
 we can determine $a$'s community by searching backward in time to find
-$T^-1(a, t_1)$, the authors that influenced $a$ at time $t$, and also searching
-forward to find $T(a, t_1)$.
+$T^{-1}(a, t)$, the authors that influenced $a$ at time $t$, and then searching
+forward to find $T(l_1, t_1) \cup T(l_1, t_2) \cup \cdots \cup T(l_k, t_k)$, where 
+$(l_1,t_1), (l_2, t_2), \ldots, (l_k, t_k)$ are the leaves of $T^{-1}(a,t)$.
 The backward search in time follows straightforwardly from the forward time traveral

### 2016-01-26 paper.tex
update acknowledgements
fix typo
Fix language and M matrix definition

@@
 
-We can use analyze the spread of information between the players using graph
+We can analyze the spread of information between the players using graph
 theory. In this process, the time ordering of events matters, and hence its
@@
 \begin{defn}
-An \textbf{active node} is a temporal node $(v,t)$ if there exists at least one
+A temporal node $(v,t)$ is an \textbf{active node} is there exists at least one
 edge $e\in E^{[t]}$ that connects $v\in V^{[t]}$ to another node $w\in V^{[t]}$,
@@
 node allows greater generality to cases where the node sets change over time.
-Furthermore, we shall show later in \ref{sec:temp-paths-adjac}
+Furthermore, we shall show later in Sec.~\ref{sec:temp-paths-adjac}
 that the explicit bookkeeping of the time labels is essential for correctly
@@
 the forward neighbors of $(1, t_1)$ are $(2, t_1)$ and $(1, t_2)$ and
-the forward neighbors of $(2, t_1)$ is $(2, t_3)$.
+the only forward neighbor of $(2, t_1)$ is $(2, t_3)$.
 The 2-forward neighbors of $(1, t_1)$ are
@@
 Define the set of temporal nodes
-$\tilde V_L^{[t]} = \{(v_1)_t = (v_1, t) | (v_1, v_2) \in E^{[t]}\}$, which
+$\tilde V_L^{[t]} = \{(v_1, t) | (v_1, v_2) \in E^{[t]}\}$, which
 consists of the active nodes at time $t$ which participate on the left side of an edge.
-Similarly, $\tilde V_R^{[t]} = \{(v_2)_t = (v_2, t) | (v_1, v_2) \in E^{[t]}\}$
+Similarly, $\tilde V_R^{[t]} = \{ (v_2, t) | (v_1, v_2) \in E^{[t]}\}$
 are the corresponding active nodes on the right side of an edge. Then
@@
 paths because they do not capture temporal paths with subpaths of the form
-$\langle (v, s), (v, t) \rangle$, $s < t$. The temporal path \eqref{eq:tpathex}
-is counted only by the matrix product $M^{[t_1]} A^{[t_2]}$, where
+$\langle (v, s), (v, t) \rangle$, $s < t$.
+One might attempt to amend the sums $S^{[t]}$ in \eqref{eq:sum} by redefining
+the adjacency matrices to include ones along the diagonal, hence allowing
+paths containing the sequence $\langle (i, t_1), (i, t_2) \rangle$. However, the resulting
+sum is still incorrect, as it counts paths with subsequences $\langle (3, t_1), (3, t_2) \rangle$
+and are hence not temporal paths. Instead, the temporal path \eqref{eq:tpathex}
+is counted only by the matrix product $M^{[t_1, t_2]} A^{[t_2]}$, where
 \[
-    M^{[t_1]} = \begin{pmatrix}1 & 0 & 0 \\ 0 & 1 & 0 \\ 0 & 0 & 0\end{pmatrix}.
+    M^{[t_1, t_2]} = \begin{pmatrix}1 & 0 & 0 \\ 0 & 0 & 0 \\ 0 & 0 & 0\end{pmatrix}.
 \]
-$M^{[t_1]}$ describes the forward time propagation of active nodes. It counts
-temporal paths with subpaths $\langle (i, t_1), (i, t_2) \rangle$
-for all active nodes $(i, t_1)$ at time $t_1$, but not paths containing the
-subsequence $\langle (3, t_1), (3, t_2) \rangle$, which cannot exist in temporal
-paths since $(3, t_1)$ is an inactive node.
-
-One might attempt to amend the sums $S^{[t]}$ in \eqref{eq:sum} by redefining
-the adjacency matrices to include ones along the diagonal. However, the resulting
-sum is still incorrect, as it counts paths with subsequences $\langle (3, t_1), (3, t_2) \rangle$
-and are hence not temporal paths.
+$M^{[t_1, t_2]}$ describes the forward time propagation of temporals nodes that
+are active at both times $t_1$ and $t_2$, i.e.\ it counts
+temporal paths that contain subsequences $\langle (i, t_1), (i, t_2) \rangle$,
+and both $(i, t_1)$ and $(i, t_2)$ are active nodes.
 
@@
 We thank N. J. Higham and V. \u Sego (Manchester) for helpful suggestions.
-W.\ Z.\ thanks the School of Mathematics at U. Manchester for travel funding and
+W.\ Z.\ thanks the School of Mathematics at U. Manchester for research \& travel funding and
 A. Edelman for arranging for a fruitful visit to MIT.
-
-%XXX ANY OTHER FUNDING
 

### 2016-01-27 paper.tex
Rewrite block adjacency matrix section

@@
 
-The proof of Theorem~\ref{thm:correct} provided a construction for representing
+The proof of Theorem~\ref{thm:correct} provides a construction for representing
 an evolving graph $G_n$ by a static graph $G$ with nodes corresponding to active
-nodes of $G_n$. The adjacency matrix $\mathbf A$ of $G$ then has block structure
-which is useful for understanding the nature of the $\odot$ operation.
+nodes of $G_n$. It turns out that the block structure of $G$ is useful for
+understanding the nature of the $\odot$ operation.
 
@@
 %
-where $c_i = (A^{[i]})^Te_k$. Summing resultant vectors that share the same
+where $c_1 = (A^{[1]})^Te_k$ and $c_i = (A^{[i]})^T \odot e_k$ for $i>1$.
+Summing resultant vectors that share the same
 time stamp, we obtain vectors whose nonzero elements have indexes labeled by
@@
 
-Define the block triangular adjacency matrix
+Compare this with the matrix
+
 \[
-\bm{A_n}^T =
-\begin{bmatrix}
-(A^{[1]})^T &   0            & \cdots  &  & 0 \\
-\sigma(A^{[2]})^T & (A^{[2]})^T & 0         &\cdots  & \vdots \\
-\sigma(A^{[3]})^T & \sigma(A^{[3]})^T & (A^{[3]})^T & \cdots \\
-   \vdots              &                              &      \vdots            &  &  \vdots \\
-\sigma(A^{[n]})^T &                  &       \cdots                      &  & (A^{[n]})^T
-\end{bmatrix},
+\bm M_n = \begin{bmatrix}
+A^{[t_1]} & M^{[t_1, t_2]} & ... & M^{[t_1, t_n]} \\
+        0 & A^{[t_2]}      & ... & M^{[t_2, t_n]} \\
+                           &   ... \\
+        0 & 0 & ... & A^{[t_n]}
+\end{bmatrix}
 \]
-where $\sigma(A^{[i]})^Tb = (A^{[i]})^T\odot b$, and the block vector
+%
+where $M^{[t_i, t_j]}$ is the matrix whose rows are labeled by $V^{[t_i]}$ and
+columns are labeled by $V^{[t_j]}$, and whose entries are
 \[
-\bm{b}^T = [b^T, 0, \cdots, 0  ].
+M^{[t_i, t_j]}_{uv} =
+\begin{cases}
+1 & \mbox{if $(u,v)\in E^\prime$,} \\
+0 & \mbox{otherwise.}
+\end{cases}
 \]
-Then, \eqref{eq:out_ne} is equivalent to the matrix--vector product
-$\bm{A_n}^T\bm{b}$ and the summation over \eqref{eq:5}-\eqref{eq:6} is equivalent
-to $\bm{A_n}^T(\bm{A_n}^T\bm{b})$.
+%
+The adjacency matrix blocks $A^{[t]}$ encode the edge set $\tilde E^{[t]}$,
+whereas the off-diagonal blocks $M^{[t_i, t_j]}$ together encode the edge set
+$E^\prime$ which capture temporal paths with subsequences of the form
+$\langle (v, t_i), (v, t_j) \rangle$. Then $\bm M_n$ is the adjacency matrix of
+the graph $(\cup_t V^{[t]}, E)$, which is the graph $G$ together with all the
+inactive nodes. From the definition, $\bm M_n$ has nonzero
+entries only in rows and columns that correspond to active nodes $V$, and so
+retaining only these rows and columns corresponding to $V$ produces the
+adjacency matrix $\bm A_n$ of $G = (V, E)$.
+
+The off-diagonal blocks $M^{[t_i, t_j]}$ provide an explicit matrix representation
+for the $\odot$ product in that $(M^{[t_i, t_j]})^T b = A^T \odot b$. The upper
+triangular structure of $\bm M_n$ (and hence $\bm A_n$) reflects the causal
+nature of temporal paths in that they cannot go backward in time.
+
+The BFS algorithm presented above can therefore be interpreted as
+computing the sequence of matrix--vector products $\bm b$, $\bm A^T \bm b$,
+$\bm (A^T)^2 \bm b$, formed by applying successive monomials of $\bm A^T$ to the
+block vector
+\[
+\bm b^T = [b^T, 0, \cdots, 0 ].
+\]
+
+For the example of Figure~\ref{fig:eg_shortest_path}, we have
+\begin{align*}
+V & = \{(1,\!t_1), (2,\!t_1), (1,\!t_2), (3,\!t_2), (2,\!t_3), (3,\!t_3)\},\\
+\cup_t\!E^{[t]} &= \{((1,\!t_1), (2,\!t_1)), ((1,\!t_2), (3,\!t_2)), ((2,\!t_3), (3,\!t_3))\},\\
+E'  &= \{((1,\!t_1), (1,\!t_2)), ((2,\!t_2), (2,\!t_3)), ((3,\!t_2), (3,\!t_3))\}.
+\end{align*}
+In the order specified for $V$, the adjacency matrix of $G$ is then
+\[
+\bm A_3 = \begin{bmatrix}
+0 & 1 & 1 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 & 1 & 0 \\
+0 & 0 & 0 & 1 & 0 & 0 \\
+0 & 0 & 0 & 0 & 0 & 1 \\
+0 & 0 & 0 & 0 & 0 & 1 \\
+0 & 0 & 0 & 0 & 0 & 0
+\end{bmatrix}
+\]
+
+Starting from the vector $\bm b = e_1$, the sequence of iterates is then
+\[
+\left\langle
+\begin{bmatrix}1 \\ 0 \\ 0 \\ 0 \\ 0 \\ 0 \end{bmatrix},
+\begin{bmatrix}0 \\ 1 \\ 1 \\ 0 \\ 0 \\ 0 \end{bmatrix},
+\begin{bmatrix}0 \\ 0 \\ 0 \\ 1 \\ 1 \\ 0 \end{bmatrix},
+\begin{bmatrix}0 \\ 0 \\ 0 \\ 0 \\ 0 \\ 2 \end{bmatrix},
+\begin{bmatrix}0 \\ 0 \\ 0 \\ 0 \\ 0 \\ 0 \end{bmatrix},
+\dots
+\right\rangle.
+\]
+
+We see that $\bm (A^T)^2$ correctly encodes the collection of products in
+\eqref{eq:5}-\eqref{eq:6}. Furthermore, $\bm (A^T)^3 \bm b$ correctly counts the
+two allowed temporal paths from $(1, t_1)$ to $(3, t_3)$, and that the
+off-diagonal structure encoded in $E'$ and the $M^{[t,t']}$ blocks are critical
+to obtaining the correct count.
 

### 2016-01-27 paper.tex
Add comment about nilpotence and termination

@@
 adjacency matrix with elements
-\[
+\begin{equation}
 A^{[t]}_{ij} =
@@
 \end{cases}
-\]
+\label{eqn:adjacency}
+\end{equation}
 
@@
 The off-diagonal blocks $M^{[t_i, t_j]}$ provide an explicit matrix representation
-for the $\odot$ product in that $(M^{[t_i, t_j]})^T b = A^T \odot b$. The upper
-triangular structure of $\bm M_n$ (and hence $\bm A_n$) reflects the causal
+for the $\odot$ product in that $(M^{[t_i, t_j]})^T b = (A^{[t_i]})^T \odot b$.
+The upper triangular structure of $\bm M_n$ (and hence $\bm A_n$) reflects the causal
 nature of temporal paths in that they cannot go backward in time.
@@
 
+%XXX Would be nice to have a figure for $G$.
 For the example of Figure~\ref{fig:eg_shortest_path}, we have
@@
 
-We see that $\bm (A^T)^2$ correctly encodes the collection of products in
-\eqref{eq:5}-\eqref{eq:6}. Furthermore, $\bm (A^T)^3 \bm b$ correctly counts the
+We see that $\bm (A_n^T)^2$ correctly encodes the collection of products in
+\eqref{eq:5}-\eqref{eq:6}. Furthermore, $\bm (A_3^T)^3 \bm b$ correctly counts the
 two allowed temporal paths from $(1, t_1)$ to $(3, t_3)$, and that the
@@
 
+Finally, we prove a simple lemma that the block adjacency matrix $\bm A_n$ is
+nilpotent whenever all the subgraphs $G^{[t]}$ of $G_n$ are acyclic.
+
+\begin{lemma}[Acyclicity implies nilpotence]
+Let $G_n = \langle G^{[t_i]} \rangle_{i=1}^n$ be an evolving directed graph and
+let all the directed graphs $G^{[t]}$ be acyclic. Then $\bm A_n$ is nilpotent.
+\label{thm:nilpotent}
+\end{lemma}
+
+\IEEEproof
+Since each directed graph $G^{[t]}$ is acyclic, its corresponding adjacency
+matrix $A^{[t]}$ is strictly upper triangular. As a result, $\bm A_n$ must be
+upper triangular.
+
+Furthermore, none of the graphs $G^{[t]}$ can have any self-edges,
+i.e.\ edges of the form $(u, u)$, and so all diagonal entries of
+$A^{[t]}$ must be zero.  Therefore all the diagonal entries of $\bm A_n$ by
+construction must be zero also.
+
+We have now proven that $\bm A_n$ is an upper triangular matrix whose diagonal
+entries are all zero. Therefore, $\bm A_n$ is nilpotent.
+\endproof
+
+Lemma~\ref{thm:nilpotent} also holds for acyclic undirected graphs so long as
+the corresponding adjacency matrix representation is encoded in an asymmetric
+fashion akin to \eqref{eqn:adjacency}.
+
 \subsection{The algebraic formulation of BFS on evolving graphs}
@@
 to write down an elegant algebraic formulation of BFS on evolving graphs, as
-presented in Algorithm~\ref{alg:bfsla}. Note that $\bm A^T$ need never be formed
+presented in Algorithm~\ref{alg:bfsla}. Note that $\bm A_n^T$ need never be formed
 explicitly, as only its matrix--vector product is required in the algorithm.
@@
  \Fn{ABFS($A_n, (v_1,t_1)$)}{
-   Form $\bm{A}^T$ from $A_n$.\\
+   Form $\bm A^T_n$ from $A_n$.\\
    $\bm{b}$ is a sparse vector of length $mn$. \\
@@
    \While{\nz{$\bm{b}$} $\ne \emptyset$}{
-       $\bm{b} = \bm{A}^T\bm{b}$ \\
+       $\bm{b} = \bm{A}^T_n\bm{b}$ \\
        \For{$k \in$ \nz{$\bm{b}$}}{
@@
 and the function \texttt{map(b)} maps a block vector's index (or indices) to their corresponding
-active node (or active nodes).}
+active nodes.}
 \label{alg:bfsla}
 \end{algorithm}
+
+\begin{thm}
+Algorithm~\ref{alg:bfsla} terminates.
+\end{thm}
+
+\IEEEproof
+Recall from Lemma~\ref{thm:nilpotent} that $\bm A_n$ is nilpotent, i.e.\
+there exists some positive integer $k$ for which $\bm A_n^k = 0$.
+Hence, after iteration $k$, $\bm b$ is assigned the value $\bm A_n^k \bm b = 0$.
+Therefore, Algorithm~\ref{alg:bfsla} must terminate after iteration $k$.
+\endproof
 

### 2016-01-27 paper.tex
Minor cleanup

@@
 The function \texttt{nonzeros(v)} returns the nonzero indices of the vector $v$,
-and the function \texttt{map(b)} maps a block vector's index (or indices) to their corresponding
+and the function \texttt{map(b)} maps a block vector's indices to their corresponding
 active nodes.}

### 2016-01-27 paper.tex
Fix typo in Algorithm 1
some minor changes

@@
 Here, we describe evolving graphs as a time-ordered sequence of graphs, similar
-to the study of metrized graphs by Tang and coworkers~\cite{ntmm13,tmml09,tsmm09,tmml10}
+to the study of metrized graphs by Tang and coworkers~\cite{ntmm13,tmml09,tsmm09,tmml10} and of 
 the dynamics of communities by Grindrod, Higham and coworkers~\cite{gphe11,grihig13}.
@@
 which in turn can have important consequences for the computation and interpretation
-of quantities which can be expressed as sums over paths, e.g.\ Katz centrality.
+of quantities that can be expressed as sums over paths, e.g.\ Katz centrality.
 
@@
 The \textbf{$k$-forward neighbors} of a temporal node $(v, t)$ are the temporal nodes
-that are the $k+1$st temporal node in every temporal path of length $k+1$ starting
+that are the $(k+1)$st temporal node in every temporal path of length $k+1$ starting
 from $(v, t)$. The \textbf{forward neighbors} of a temporal node $(v, t)$ are its
@@
      \For{$(v',t') \in$\nei{$(v,t)$}}{
-          \If{$(v',t') \notin d$}{
+          \If{$(v',t') \notin reached$}{
               $reached[(v',t')] = k$\\
@@
 S^{[t_n]} = A^{[t_1]}A^{[t_n]} + \sum_{t_1 \le t \le t_n} A^{[t_1]}A^{[t]}A^{[t_n]}
-+ \dots \nonumber \\
++ \cdots \nonumber \\
 + \sum_{t_1 \le t \le t' \le \dots \le t_n}
@@
 manipulation, and study of evolving graphs written in Julia~\cite{bkse12}.
-The implementation of the evolving graph BFS is
-freely available online at
+It is freely available online at 
 \begin{center}
@@
 \end{center}
-and available with the MIT ``Expat'' license.
+and available with the MIT ``Expat'' license. The package contains an implementation
+of the evolving graph BFS.
 

### 2016-01-27 paper.tex
add citation for different graph names
reorganize linear algebra section

@@
 Such a graph is called an ``evolving graph''~\cite{fkp89,bkmu12},
-``evolving network'' or %XXX WHO?
-``time-dependent graph''. %XXX WHO?
+``evolving network''~ \cite{bfgm08} or 
+``temporal graph''~\cite{tmml09}.
 
@@
 \label{sec:temp-paths-adjac}
+
+For each static graph $G^{[t]} = (V^{[t]}, E^{[t]})$ that constitutes the
+evolving graph $G_n$, define its corresponding
+$\left\vert V^{[t]}\right\vert \times \left\vert V^{[t]}\right\vert$
+adjacency matrix with elements
+\begin{equation}
+A^{[t]}_{ij} =
+\begin{cases}
+ 1 & \mbox{if $(i,j)\in E^{[t]}$,} \\
+ 0 & \mbox{otherwise.}
+\end{cases}
+\label{eqn:adjacency}
+\end{equation}
+
+We can then represent $G_n$ using the sequence of adjacency matrices
+$A_n = \langle A^{[1]}, A^{[2]}, \ldots, A^{n]}\rangle$. The example in
+Figure~\ref{fig:eg_shortest_path} can be represented as
+\[
+\left\langle
+  \begin{bmatrix}
+    0 & 1 & 0 \\
+    0 & 0 & 0 \\
+    0 & 0 & 0
+  \end{bmatrix},~
+ \begin{bmatrix}
+   0 & 0 & 1 \\
+   0 & 0 & 0 \\
+   0 & 0 & 0
+ \end{bmatrix},~
+ \begin{bmatrix}
+  0 & 0 & 0 \\
+  0 & 0 & 1 \\
+  0 & 0 & 0
+ \end{bmatrix}
+\right\rangle.
+\]
 
@@
 
-\subsection{Evolving graphs as sequences of adjacency matrices}
-
-For each static graph $G^{[t]} = (V^{[t]}, E^{[t]})$ that constitutes the
-evolving graph $G_n$, define its corresponding
-$\left\vert V^{[t]}\right\vert \times \left\vert V^{[t]}\right\vert$
-adjacency matrix with elements
-\begin{equation}
-A^{[t]}_{ij} =
-\begin{cases}
- 1 & \mbox{if $(i,j)\in E^{[t]}$,} \\
- 0 & \mbox{otherwise.}
-\end{cases}
-\label{eqn:adjacency}
-\end{equation}
-
-We can then represent $G_n$ using the sequence of adjacency matrices
-$A_n = \langle A^{[1]}, A^{[2]}, \ldots, A^{n]}\rangle$. The example in
-Figure~\ref{fig:eg_shortest_path} can be represented as
-\[
-\left\langle
-  \begin{bmatrix}
-    0 & 1 & 0 \\
-    0 & 0 & 0 \\
-    0 & 0 & 0
-  \end{bmatrix},~
- \begin{bmatrix}
-   0 & 0 & 1 \\
-   0 & 0 & 0 \\
-   0 & 0 & 0
- \end{bmatrix},~
- \begin{bmatrix}
-  0 & 0 & 0 \\
-  0 & 0 & 1 \\
-  0 & 0 & 0
- \end{bmatrix}
-\right\rangle.
-\]
-
-This algebraic representation allows us to exploit a graphical interpretation of
+\subsection{Defining algebraic forward neighbors}
+
+The algebraic representation of evolving graphs presented in Section \ref{sec:temp-paths-adjac} 
+allows us to exploit a graphical interpretation of
 matrix--vector products involving the adjacency matrix~\cite{kegi11}.
@@
 computing the sequence of matrix--vector products $\bm b$, $\bm A^T \bm b$,
-$\bm (A^T)^2 \bm b$, formed by applying successive monomials of $\bm A^T$ to the
+$(\bm A^T)^2 \bm b$, formed by applying successive monomials of $\bm A^T$ to the
 block vector

### 2016-01-27 paper.tex
Fix typos

@@
 winner can be described in terms of discrete paths through the evolving graph,
-and can be interpreted naturally as a graph traveral problem. Traversals of an
+and can be interpreted naturally as a graph traversal problem. Traversals of an
 ordinary (static) graph may be computed using well known methods such as the
 breadth-first search (BFS). However, a proper description of the dynamics
-requires paths that also move monotonically forward in time, thus necessating
+requires paths that also move monotonically forward in time, thus necessitating
 a generalization of the ordinary BFS algorithm.
@@
 adjacency lists, one for each active node in $G_n$. Then we have that the
-asympotic complexity of BFS on $G_n$ is the same as that for BFS on $G$, using
+asymptotic complexity of BFS on $G_n$ is the same as that for BFS on $G$, using
 the 1-1 construction of $G$ from $G_n$.
@@
 \]
-$M^{[t_1, t_2]}$ describes the forward time propagation of temporals nodes that
+$M^{[t_1, t_2]}$ describes the forward time propagation of temporal nodes that
 are active at both times $t_1$ and $t_2$, i.e.\ it counts
@@
 The matrix--vector product $\bm A^T \bm b$ encodes the forward neighbors of all
-the frontier nodes. Subequently, active nodes that have already been visited in
+the frontier nodes. Subsequently, active nodes that have already been visited in
 previous iterations are zeroed out of the new $\bm b$.
@@
 $(l_1,t_1), (l_2, t_2), \ldots, (l_k, t_k)$ are the leaves of $T^{-1}(a,t)$.
-The backward search in time follows straightforwardly from the forward time traveral
+The backward search in time follows straightforwardly from the forward time traversal
 presented above simply by reversing the time labels, e.g.\ by the transformation

### 2016-01-27 paper.tex
add figure to illustrate graph G

@@
 
-%XXX Would be nice to have a figure for $G$.
+\begin{figure}[h]
+ \begin{center}
+   \begin{tikzpicture}[scale=.36, line width =1.4pt]
+  \node[circle,fill=green1, minimum size=0.1cm, inner sep = 0pt] (n7) at (-5,7) 
+{\scriptsize $(2,t_1)$};
+  \node[circle,draw=red1, minimum size=0.2cm, inner sep = 0pt] (n8) at (-10,5) 
+{\scriptsize $(3,t_1)$};
+  \node[circle,fill=green1, minimum size=0.1cm, inner sep = 0pt] (n10) at (-8,10) {\scriptsize $(1,t_1)$};
+
+  \node[circle,draw=red1, minimum size=0.2cm, inner sep = 0pt] (n6) at (3,7) 
+{\scriptsize $(2,t_2)$};
+  \node[circle,fill=green1, minimum size=0.2cm,  inner sep = 0pt] (n4) at (-2,5) 
+{\scriptsize $(3,t_2)$};
+  \node[circle,fill=green1, minimum size=0.2cm,  inner sep = 0pt] (n1) at (0,10) 
+{\scriptsize $(1,t_2)$};
+
+   \node[circle,fill=green1, minimum size=0.2cm,  inner sep = 0pt] (n11) at (11,7) 
+{\scriptsize $(2,t_3)$};
+  \node[circle,fill=green1, minimum size=0.2cm,  inner sep = 0pt] (n12) at (6,5)  
+{\scriptsize $(3,t_3)$};
+  \node[circle,draw=red1, minimum size=0.2cm,  inner sep = 0pt] (n14) at (8,10) 
+{\scriptsize $(1,t_3)$};
+
+  \foreach \from/\to in {n10/n7, n1/n4, n11/n12}
+   \draw[every edge,->] (\from) -- (\to);
+     \draw[dotted,->](n7) to[out=80, in=40] (n11);
+  \draw[dotted,->](n10) to[out=10, in=180] (n1);
+   \draw[dotted,->](n4) to[out=-30,in=-150] (n12);
+    \end{tikzpicture}
+\end{center}
+\caption{The static graph representation of Figure \ref{fig:eg_shortest_path}. 
+The black lines are edges from the edge set $\cup_t E^{[t]}$ and the 
+doted lines are edges from the edge set $E'$.}
+\label{fig:static}
+\end{figure}
+
 For the example of Figure~\ref{fig:eg_shortest_path}, we have

### 2016-01-27 paper.tex
emphasiz node activeness
timestamp -> time stamp

@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{Breadth First Searches Over Evolving Graphs}
+\title{The Right Way to Search Evolving Graphs}
 
@@
 to the study of metrized graphs by Tang and coworkers~\cite{ntmm13,tmml09,tsmm09,tmml10} and of 
-the dynamics of communities by Grindrod, Higham and coworkers~\cite{gphe11,grihig13}.
+the dynamics of communities by Grindrod, Higham and coworkers~\cite{gphe11,grihig13}. However, we emphasize a concept called
+\emph{node activeness} and we only study the active nodes on an evolving graph. 
+In real-life problems, the set of active nodes is usually sparse at a given time stamp. 
+Our model, as a result, can be used to model highly dynamic large evolving graph problems, where other models fail to.
 

### 2016-01-28 paper.tex
change title
More language tweaks

--- 2016-01-27 paper.tex
@@
 
-The game described at the beginning can be encoded in terms of analyzing the
-microscopic dynamics of an evolving graph. The spread of information to the
-winner can be described in terms of discrete paths through the evolving graph,
-and can be interpreted naturally as a graph traversal problem. Traversals of an
-ordinary (static) graph may be computed using well known methods such as the
-breadth-first search (BFS). However, a proper description of the dynamics
-requires paths that also move monotonically forward in time, thus necessitating
-a generalization of the ordinary BFS algorithm.
-
-An informal description of BFS generalized to evolving graphs can be found in
-\cite{tmml09}. However, it turns out that the proper generalization involves
-paths of nontrivial structure which sometimes contradict na\"ive generalizations
-of results derived from static graphs.
-In this paper, we present a complete description of the BFS algorithm for evolving graphs.
-The algebraic formulation of this algorithm makes explicit several subtleties in the correct
-determination of relevant trajectories, which we term \textit{temporal paths},
-which in turn can have important consequences for the computation and interpretation
-of quantities that can be expressed as sums over paths, e.g.\ Katz centrality.
+The game described in the beginning can be encoded in an evolving graph.
+The spread of information to the winner can be described in terms of traversing
+this graph using discrete paths that step in both space and time.
+Traversals of an ordinary (static) graph may be computed using well known methods
+such as the breadth-first search (BFS). An informal description of BFS generalized
+to evolving graphs can be found in \cite{tmml09}.
+However, it turns out that na\"ive extensions can lead to incorrect descriptions of
+the resulting graph traversals. A proper treatment requires the notion of
+\emph{node activeness} to describe the set of paths that can only traverse time
+or edges. We term such paths \textit{temporal paths}.
+As a result, our treatment can be applied to any evolving graph, even those that
+are highly dynamic with arbitrary changes to the nodes and edges.
+
+In this paper, we explain how the BFS algorithm can be applied correctly on
+evolving graphs. We also provide a counterexample demonstrating that some na\"ive
+analyses in terms of discrete path sums over the time-independent adjacency
+matrices lead to incorrect results. The correct generalization to evolving graphs
+is formally equivalent to BFS over a particular static graph generated by connecting
+together active notes of the evolving graph. We also show that BFS over evolving
+graphs can be interpreted algebraically as computing a sequence of vectors formed
+by repeatedly multiplying a starting vector with the adjacency matrix of the
+equivalent static graph, and in so doing reveals interesting connections between
+properties of the BFS algorithm and properties of the adjacency matrix. Finally,
+we explain how BFS on evolving graphs may be applied to study dynamical processes
+over citation networks.
 
@@
 At each time stamp, the evolving graph is represented as a graph.
-The green filled nodes in the figure are active nodes.}
+The green filled circles represent active nodes while the red circles represent
+inactive nodes.}
 \label{fig:eg_shortest_path}
@@
 The key new idea in generalizing BFS to evolving graphs is to be able to compute
-paths that traverse multiple times. We call the relevant paths
-\textit{temporal paths}.
+paths that evolve forward in time and can only traverse the node space along
+existing edges. We call these paths \textit{temporal paths}.
 
@@
 $\langle (1, t_1), (1, t_2), (3, t_2), (3, t_3) \rangle$ and
-$\langle (1, t_1), (2, t_1), (2, t_3), (3, t_3) \rangle$
+$\langle (1, t_1), (2, t_1)$, $(2, t_3), (3, t_3) \rangle$
 are both examples of \textit{temporal paths} from $(1, t_1)$ to $(3, t_3)$,
@@
 \end{center}
-\caption{An evolving graph with 3 time stamps $t_1$, $t_2$ and $t_3$.
-The black dashed lines of the two figures indicate two temporal paths from $(1, t_1)$ to $(3,t_3)$.}
+\caption{Two temporal paths from $(1, t_1)$ to $(3,t_3)$ on the evolving graph
+shown in Figure~\ref{fig:eg_shortest_path}, shown in black dashed lines.
+The paths traverse only active nodes along edges, and are allowed to advance
+between the same node if it is active at different times.}
 \label{fig:active}
@@
 \end{center}
-\caption{An evolving graph with 3 time stamps $t_1$, $t_2$ and $t_3$.
-The circled green filled nodes indicate a BFS process starting at active node
-$(1, t_2)$.}
+\caption{Breadth-first search (BFS) on the evolving graph shown in
+Figure~\ref{fig:eg_shortest_path} starting from the root $(1, t_2)$.
+Note that the time $t_1$ does not participate in the BFS. Black circles indicate
+the active nodes forming the $frontier$ and $nextfrontier$ sets in Algorithm~\ref{alg:bfs},
+connected by the dotted black lines.}
 \label{fig:example}
@@
 We now establish a similar 1-1 correspondence of forward neighbors of an active
-node with a subset of $G$.
-By definition, the forward neighbors of some active node $(v, t) \in G_n$ are
+node with a subset of $G$. By induction, we can show that at iteration $k$,
+all new nodes populated into the key set of $reached$ by definition are of
+distance $k$ from the root. By definition,
+the forward neighbors of some active node $(v, t) \in G_n$ are
 active nodes of either the form $(v, t')$ for some $t' > t$ or $(u, t)$ for some
@@
 
-\subsection{Not All Temporal Paths Can Be Expressed by Products of Adjacency Matrices}
+\subsection{Not all temporal paths are counted by products of adjacency matrices}
 \label{sec:temp-paths-adjac}
@@
 and are hence not temporal paths. Instead, the temporal path \eqref{eq:tpathex}
-is counted only by the matrix product $M^{[t_1, t_2]} A^{[t_2]}$, where
-\[
+is counted by the matrix product $M^{[t_1, t_2]} A^{[t_2]}$, where
+\begin{equation}
     M^{[t_1, t_2]} = \begin{pmatrix}1 & 0 & 0 \\ 0 & 0 & 0 \\ 0 & 0 & 0\end{pmatrix}.
-\]
+\label{eq:m-example}
+\end{equation}
 $M^{[t_1, t_2]}$ describes the forward time propagation of temporal nodes that
@@
 for the $\odot$ product in that $(M^{[t_i, t_j]})^T b = (A^{[t_i]})^T \odot b$.
+An example of such an off-diagonal block was already provided in \eqref{eq:m-example}.
+These off-diagonal blocks represent traversal between active notes with the same
+node space labels but are still separated by time, and are essential for
+the correct enumeration of temporal paths.
 The upper triangular structure of $\bm M_n$ (and hence $\bm A_n$) reflects the causal
@@
 The BFS algorithm presented above can therefore be interpreted as
-computing the sequence of matrix--vector products $\bm b$, $\bm A^T \bm b$,
-$(\bm A^T)^2 \bm b$, formed by applying successive monomials of $\bm A^T$ to the
+computing the sequence of matrix--vector products $\bm b$, $\bm A_n^T \bm b$,
+$(\bm A_n^T)^2 \bm b$, ..., formed by applying successive monomials of $\bm A_n^T$ to the
 block vector
-\[
-\bm b^T = [b^T, 0, \cdots, 0 ].
-\]
+$\bm b^T = [b^T, 0, \cdots, 0 ]$
+where $b^T$ encodes the root in the space of active nodes $\tilde V^{[t_1]}$.
 
@@
 \end{center}
-\caption{The static graph representation of Figure \ref{fig:eg_shortest_path}. 
-The black lines are edges from the edge set $\cup_t E^{[t]}$ and the 
-doted lines are edges from the edge set $E'$.}
+\caption{Static graphs corresponding to the evolving graph example of
+Figure~\ref{fig:eg_shortest_path}. The green nodes are active nodes while the
+red nodes are inactive nodes.
+The black lines are edges in $\cup_t E^{[t]}$ and are encoded
+algebraically in the diagonal blocks $A^{[t]}$ of the adjacency matrix $\bm A_3$
+or $\bm M_3$.
+The dotted lines are edges in $E'$ and are encoded algebraically in
+the off-diagonal blocks $M^{[t_i, t_j]}$.
+The static graph $G$ constructed in the proof of Theorem~\ref{thm:correct} is
+formed by retaining all the edges shown and only the active nodes, and has the
+adjacency matrix $\bm A_3$. The graph containing all the edges and temporal nodes
+shown has adjacency matrix $\bm M_3$.}
 \label{fig:static}
@@
 
-We see that $\bm (A_n^T)^2$ correctly encodes the collection of products in
-\eqref{eq:5}-\eqref{eq:6}. Furthermore, $\bm (A_3^T)^3 \bm b$ correctly counts the
+We see that $(\bm A_n^T)^2$ correctly encodes the collection of products in
+\eqref{eq:5}-\eqref{eq:6}. Furthermore, $(\bm A_3^T)^3 \bm b$ correctly counts the
 two allowed temporal paths from $(1, t_1)$ to $(3, t_3)$, and that the
@@
  \SetKwFunction{intersect}{intersect}
- \SetKwFunction{map}{map}
+ \SetKwFunction{map}{activeNodes}
  \SetKwProg{Fn}{function}{}{end}
@@
    Form $\bm A^T_n$ from $A_n$.\\
-   $\bm{b}$ is a sparse vector of length $mn$. \\
-   $\bm{b}_s = 1$ \\
+   $\bm{b}_{v_1} = 1$ \\
    $k = 1$ \\
@@
       }
-      $activeNodes =$\map{$\bm{b}$}\\
-       \For{$activeNode \in activeNodes$}{
-           $reached[activeNode] = k$ \\
+       \For{$node \in \map{\bm{b}}$}{
+           $reached[node] = k$ \\
          }
@@
 frontier nodes encoding the $frontier$ set of Algorithm~\ref{alg:bfs}.
-The matrix--vector product $\bm A^T \bm b$ encodes the forward neighbors of all
+The matrix--vector product $\bm A_n^T \bm b$ encodes the forward neighbors of all
 the frontier nodes. Subsequently, active nodes that have already been visited in
@@
 This paper introduced the generalization of BFS to evolving graphs and presented
-an equivalent formulation using linear algebra. The concept temporal paths
+an equivalent formulation using linear algebra. The concept of temporal paths
 connecting active nodes turns out to be critical for the correct generalization

### 2016-01-28 paper.tex
Explain that you should never compute with the block matrix representation

@@
 
+The blocked matrix structure of the adjacency matrices presented here provide
+interesting relationships between their matrix properties and the algorithmic
+properties of BFS, made possible because of the reformulation of BFS as
+repeated power iterations of the adjacency matrix in Algorithm~\ref{alg:bfsla}.
+Note, however, that these matrices never need to be instantiated for practical
+computations. Rather, since Algorithm~\ref{alg:bfsla} only requires the matrix--
+vector product involving the adjacency matrix, the formulation of
+Algorithm~\ref{alg:bfs} provides an efficient way to exploit the block structure
+of $\bm A_n$. The $\odot$ operation provides an efficient way to compute the
+action of the off-diagonal products. Representing the diagonal blocks
+$A^{[t]}$ as sparse matrices product further reduces the cost of BFS by exploiting
+latent sparsity in graphs that show up in practical applications.
+
 \subsection{The algebraic formulation of BFS on evolving graphs}

### 2016-01-28 paper.tex
Fix grammar of subsection
Fix errors in typesetting

@@
 
-\subsection{Defining algebraic forward neighbors}
+\subsection{Defining forward neighbors algebraically}
 
@@
 Note, however, that these matrices never need to be instantiated for practical
-computations. Rather, since Algorithm~\ref{alg:bfsla} only requires the matrix--
-vector product involving the adjacency matrix, the formulation of
+computations. Rather, since Algorithm~\ref{alg:bfsla} only requires the m
+matrix--vector product involving the adjacency matrix, the formulation of
 Algorithm~\ref{alg:bfs} provides an efficient way to exploit the block structure
@@
       }
-       \For{$node \in \map{\bm{b}}$}{
+       \For{$node \in $\map{$\bm{b}$}}{
            $reached[node] = k$ \\

### 2016-01-28 paper.tex
change abstract

@@
 \begin{abstract}
-We generalize the breadth first search (BFS) algorithm to evolving graphs and
-prove its correctness.
-The BFS generates paths with nontrivial structure, which we term temporal paths.
-We also show with simple counterexamples that products of adjacency matrices alone
-do not have the correct combinatorics of enumerating temporal paths.
-Generalizing BFS to evolving graphs requires careful tracking of active nodes
-which participate in the definition of edges at specific times.
-We discuss how the structure of temporal paths is reflected in adjacency matrix
-representations of evolving graphs,
-as well as implications for real world applications such as the analysis of citation networks.
+We present a correctly verified breadth first search (BFS) algorithm for evolving graphs.
+Using simple examples, we show na\"ive matrix-matrix multiplication on time-dependent adjacency matrices miscounts the number of temporal paths and therefore should be avoided if temporal path is a component of the algorithm.
+We formulate our
+evolving graph BFS using the language of linear algebra and present an adjacency matrix representation of the equivalent static graph, which is used to study
+interesting properties about the original evolving graphs. 
+% We generalize the breadth first search (BFS) algorithm to evolving graphs and
+% prove its correctness.
+% The BFS generates paths with nontrivial structure, which we term temporal paths.
+% We also show with simple counterexamples that products of adjacency matrices alone
+% do not have the correct combinatorics of enumerating temporal paths.
+% Generalizing BFS to evolving graphs requires careful tracking of active nodes
+% which participate in the definition of edges at specific times.
+% We discuss how the structure of temporal paths is reflected in adjacency matrix
+% representations of evolving graphs,
+% as well as implications for real world applications such as the analysis of citation networks.
 \end{abstract}
@@
 evolving graphs. We also provide a counterexample demonstrating that some na\"ive
-analyses in terms of discrete path sums over the time-independent adjacency
+analyses in terms of discrete path sums over the time-dependent adjacency
 matrices lead to incorrect results. The correct generalization to evolving graphs

### 2016-01-28 paper.tex
change conclusion

@@
 
-This paper introduced the generalization of BFS to evolving graphs and presented
-an equivalent formulation using linear algebra. The concept of temporal paths
-connecting active nodes turns out to be critical for the correct generalization
-of BFS to evolving graphs. An efficient implementation of BFS on evolving
-graphs is now available as a open source Julia package.
-
+This paper contemplated one of basic archetypes of graph traversal algorithms
+for evolving graphs.
+Although rarely mentioned in previous papers,
+we showed the concept of node activeness is actually crucial for designing correct and efficient algorithms for evolving graphs.
+
+% This paper introduced the generalization of BFS to evolving graphs and presented
+% an equivalent formulation using linear algebra. The concept of temporal paths
+% connecting active nodes turns out to be critical for the correct generalization
+% of BFS to evolving graphs. An efficient implementation of BFS on evolving
+% graphs is now available as a open source Julia package.
 % This paper leaves important extensions open for further research. First, extensions

### 2016-01-28 paper.tex
update conclusion

@@
 \begin{defn}
-A temporal node $(v,t)$ is an \textbf{active node} is there exists at least one
+A temporal node $(v,t)$ is an \textbf{active node} if there exists at least one
 edge $e\in E^{[t]}$ that connects $v\in V^{[t]}$ to another node $w\in V^{[t]}$,
@@
 
-This paper contemplated one of basic archetypes of graph traversal algorithms
-for evolving graphs.
-Although rarely mentioned in previous papers,
-we showed the concept of node activeness is actually crucial for designing correct and efficient algorithms for evolving graphs.
-
+This paper contemplated one of basic archetypes of evolving graph traversal algorithms. The treatment interweaved graph theory with linear algebra. 
+Correct tracking of temporal paths on active nodes 
+provides the right way to search evolving graphs, 
+which can also be applied to other algorithms for evolving graphs.
 % This paper introduced the generalization of BFS to evolving graphs and presented

### 2016-01-28 paper.tex
update abstract

@@
 \begin{abstract}
-We present a correctly verified breadth first search (BFS) algorithm for evolving graphs.
-Using simple examples, we show na\"ive matrix-matrix multiplication on time-dependent adjacency matrices miscounts the number of temporal paths and therefore should be avoided if temporal path is a component of the algorithm.
-We formulate our
-evolving graph BFS using the language of linear algebra and present an adjacency matrix representation of the equivalent static graph, which is used to study
-interesting properties about the original evolving graphs. 
+Evolving graphs arise in many different contexts where the 
+interrelations between data elements change over time. 
+We present a breadth first search (BFS) algorithm for evolving graphs
+that can track (active) nodes correctly. 
+Using simple examples, we show na\"ive matrix-matrix multiplication on time-dependent adjacency matrices miscounts the number of temporal paths. 
+By mapping an evolving graph to an adjacency matrix of the equivalent static graph, 
+we  prove the properties of the BFS algorithm using the properties of the 
+adjacency matrix.
+We show an application of our algorithm in mining citation networks.
+
 % We generalize the breadth first search (BFS) algorithm to evolving graphs and

### 2016-01-28 paper.tex
fix typo

@@
 
-Figure~\ref{fig:eg_shortest_path} shows a small example of a evolving
+Figure~\ref{fig:eg_shortest_path} shows a small example of an evolving
 directed graph, $G_3 = \langle G^{[1]}, G^{[2]}, G^{[3]}\rangle$, consisting of

### 2016-01-29 paper.tex
use 10-point size
More edits

@@
-\documentclass[11pt,conference,compsocconf]{IEEEtran}
+\documentclass[10pt,conference,compsocconf]{IEEEtran}
 \usepackage{tikz}
@@
 that can track (active) nodes correctly.
-Using simple examples, we show na\"ive matrix-matrix multiplication on time-dependent adjacency matrices miscounts the number of temporal paths. 
+Using simple examples, we show na\"ive matrix-matrix multiplication on
+time-dependent adjacency matrices miscounts the number of temporal paths.
 By mapping an evolving graph to an adjacency matrix of the equivalent static graph,
 we prove the properties of the BFS algorithm using the properties of the
-adjacency matrix.
-We show an application of our algorithm in mining citation networks.
+adjacency matrix. Finally, demonstrate how the BFS over evolving graphs can be
+applied to mining citation networks.
 
@@
 Such a graph is called an ``evolving graph''~\cite{fkp89,bkmu12},
-``evolving network''~ \cite{bfgm08} or
-``temporal graph''~\cite{tmml09}.
+``evolving network''~\cite{bfgm08} or ``temporal graph''~\cite{tmml09}.
 
 Treatments of evolving graphs vary in their generality and focus.
-Kivel\"a \textit{et al.}~\cite{kabg14} treats time dependence as a special
+Kivel\"a \textit{et al.}~\cite{kabg14} treat time dependence as a special
 case of families of graphs with multiple interrelationships.
@@
 \emph{node activeness} to describe the set of paths that can only traverse time
-or edges. We term such paths \textit{temporal paths}.
+or edges, which we call \textit{temporal paths}.
 As a result, our treatment can be applied to any evolving graph, even those that
@@
 
-In this paper, we explain how the BFS algorithm can be applied correctly on
-evolving graphs. We also provide a counterexample demonstrating that some na\"ive
-analyses in terms of discrete path sums over the time-dependent adjacency
-matrices lead to incorrect results. The correct generalization to evolving graphs
-is formally equivalent to BFS over a particular static graph generated by connecting
-together active notes of the evolving graph. We also show that BFS over evolving
-graphs can be interpreted algebraically as computing a sequence of vectors formed
-by repeatedly multiplying a starting vector with the adjacency matrix of the
-equivalent static graph, and in so doing reveals interesting connections between
-properties of the BFS algorithm and properties of the adjacency matrix. Finally,
+In Section~\ref{sec:breadth-first-search}, we explain how the BFS algorithm can be applied correctly on
+evolving graphs. Section~\ref{sec:temporal-paths} provides an example showing that considering
+only products of the time-dependent adjacency matrices can lead to incorrect enumeration
+of temporal paths. We present and demonstate the BFS algorithm over evolving graphs
+in Section~\ref{sec:evolving-graph-bfs}, showing that it is formally equivalent to
+BFS over a particular static graph generated by connecting
+together active notes of the evolving graph. This static graph generates an algebraic
+representation of the BFS as power iteration of its adjacency matrix to a starting
+search node, as shown in Section~\ref{sec:bfsla}. The algebraic formulation also
+demonstrates interesting connections between properties of the BFS algorithm and
+the adjacency matrix. Finally in Section~\ref{sec:applications},
 we explain how BFS on evolving graphs may be applied to study dynamical processes
@@
 \end{center}
-\caption{An evolving graph with 3 time stamps $t_1$, $t_2$ and $t_3$.
+\caption{An evolving directed graph with 3 time stamps $t_1$, $t_2$ and $t_3$.
 At each time stamp, the evolving graph is represented as a graph.
 The green filled circles represent active nodes while the red circles represent
-inactive nodes.}
+inactive nodes. Directed edges are shown as black errors.}
 \label{fig:eg_shortest_path}
@@
 \end{center}
-\caption{Two temporal paths from $(1, t_1)$ to $(3,t_3)$ on the evolving graph
+\caption{The two temporal paths of length 4 from $(1, t_1)$ to $(3,t_3)$ on the evolving graph
 shown in Figure~\ref{fig:eg_shortest_path}, shown in black dashed lines.
@@
 The example presented above in Section~\ref{sec:temporal-paths}
-demonstrates how the notion of active nodes restrict the set of temporal paths
+demonstrates how active nodes restrict the set of temporal paths
 that need to be considered when traversing an evolving graph.
@@
 of temporal paths.
-Our notation generalizes the standard notations for static graphs presented in
-\cite{even12,kegi11}.
+Our notation generalizes that for static graphs presented in \cite{even12,kegi11}.
 
@@
 \begin{defn}
-The distance from a temporal node $(v, t)$ to a temporal node $(w, s)$ is
+The \textbf{distance} from a temporal node $(v, t)$ to a temporal node $(w, s)$ is
 the $k$ for which $(w, s)$ is a $k$-forward neighbor of $(v, t)$.
@@
 Similarly, $\tilde V_R^{[t]} = \{ (v_2, t) | (v_1, v_2) \in E^{[t]}\}$
-are the corresponding active nodes on the right side of an edge. Then
+contains the corresponding active nodes on the right side of an edge. Then
 $\tilde V^{[t]} = \tilde V_L^{[t]} \cup \tilde V_R^{[t]}$ is the set of active
@@
 The nonzero entries of the first vector represent forward neighbors that are on
-the same time stamp $t_1$, whereas all other nonzero entries represent forward
+the same time stamp $t_1$, whereas nonzero entries of the other vectors represent forward
 neighbors that are advanced in time but remain on the same node $k$.
@@
 
-Consider two iterations of the BFS on $G_n$ with root $(k, t_1)$. The second
-iteration requires computing the sequences
+Consider the second iteration of BFS on $G_n$ with root $(k, t_1)$, which
+requires computing the sequences
 
@@
 Note, however, that these matrices never need to be instantiated for practical
-computations. Rather, since Algorithm~\ref{alg:bfsla} only requires the m
+computations. Rather, since Algorithm~\ref{alg:bfsla} only requires the
 matrix--vector product involving the adjacency matrix, the formulation of
@@
 \subsection{The algebraic formulation of BFS on evolving graphs}
+\label{sec:bfsla}
 

### 2016-01-29 paper.tex
More comprehensive conclusion

@@
 only products of the time-dependent adjacency matrices can lead to incorrect enumeration
-of temporal paths. We present and demonstate the BFS algorithm over evolving graphs
+of temporal paths. We present and demonstrate the BFS algorithm over evolving graphs
 in Section~\ref{sec:evolving-graph-bfs}, showing that it is formally equivalent to
@@
 
-This paper contemplated one of basic archetypes of evolving graph traversal algorithms. The treatment interweaved graph theory with linear algebra.
-Correct tracking of temporal paths on active nodes
-provides the right way to search evolving graphs,
-which can also be applied to other algorithms for evolving graphs.
+The correct generalization of BFS to evolving graphs necessitates a careful
+enumeration of temporal paths which is not always captured by products of
+successive adjacency matrices from the snapshots. BFS over any evolving graph $G_n$
+is formally equivalent to BFS on a related static graph $G$ formed by connecting only
+active nodes across different times. The adjacency matrix $\bm A_n$ of $G$ correctly
+encodes causality of the temporal paths. When specialized to acyclic $G_n$, the matrix
+$\bm A_n$ is also nilpotent, thus providing a simple proof that BFS terminates.
+Nevertheless, the BFS can be computed efficiently without ever explicitly constructing
+$\bm A_n$, as only its action on a vector is required for the power iterations.
+The new concepts of activeness, temporal paths, and the explicit construction of
+an equivalent static graph $G$ that make possible a correct implementation of BFS
+to evolving graphs provide powerful new insignts into how similar graphical
+algorithms may be generalized correctly.
+
 % This paper introduced the generalization of BFS to evolving graphs and presented

### 2016-02-01 paper.tex
HYPERLINKS NOT ALLOWED BY IEEE
Add arXiv build target
Also fix typo
Fix embarrassing typo

@@
 
-\usepackage{hyperref}
+% NOT ALLOWED BY IEEE
+% \usepackage{hyperref}
 % *** MISC UTILITY PACKAGES ***
@@
 that can track (active) nodes correctly.
-Using simple examples, we show na\"ive matrix-matrix multiplication on
+Using simple examples, we show that na\"ive matrix-matrix multiplication on
 time-dependent adjacency matrices miscounts the number of temporal paths.
@@
 The green filled circles represent active nodes while the red circles represent
-inactive nodes. Directed edges are shown as black errors.}
+inactive nodes. Directed edges are shown as black arrows.}
 \label{fig:eg_shortest_path}

### 2016-02-18 paper.tex
Fix embarrassing typo
fix errors in the paper

@@
 The green filled circles represent active nodes while the red circles represent
-inactive nodes. Directed edges are shown as black arrows.}
+inactive nodes. Directed edges are shown as black errors.}
 \label{fig:eg_shortest_path}
@@
 \begin{thm}[Correctness of the evolving graph BFS]
-Let $G_n$ be an evolving graph and $(v_1, t_1)$ be an active node of $G_n$.
+Let $G_n$ be a directed evolving graph and $(v_1, t_1)$ be an active node of $G_n$.
 Then Algorithm~\ref{alg:bfs} discovers every active node that is reachable from
@@
 requires a new kind of matrix--vector product, $\odot$, defined by
+
 \[
@@
 \begin{cases}
-b & \mbox{if $A^Tb \ne 0$,} \\
+b & \mbox{if $A^Tb \ne 0$ or $Ab \ne 0$,} \\
 0 & \mbox{otherwise.}

### 2016-02-18 paper.tex
clarify figure 3

@@
  \begin{center}
-    \begin{tikzpicture}[scale=.36, line width =1.4pt]
+    \begin{tikzpicture}[scale=.31, line width =1.4pt]
+ \node[text width=3cm] at (-9.5, 7.5) {$k=1$};
   \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {2};
@@
     \end{tikzpicture}
-   \begin{tikzpicture}[scale=.36, line width =1.4pt]
+   \begin{tikzpicture}[scale=.31, line width =1.4pt]
+ \node[text width=3cm] at (-9.5, 7.5) {$k=2$};
   \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {2};
@@
     \end{tikzpicture}
-   \begin{tikzpicture}[scale=.36, line width =1.4pt]
+   \begin{tikzpicture}[scale=.31, line width =1.4pt]
+ \node[text width=3cm] at (-9.5, 7.5) {$k=3$};
   \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {2};
@@
 \caption{Breadth-first search (BFS) on the evolving graph shown in
-Figure~\ref{fig:eg_shortest_path} starting from the root $(1, t_2)$.
+Figure~\ref{fig:eg_shortest_path} starting from the root $(1, t_2)$ at iteration
+$k = 1, 2, 3$.
 Note that the time $t_1$ does not participate in the BFS. Black circles indicate

### 2016-02-18 paper.tex
add BFS experiment table

@@
 and available with the MIT ``Expat'' license. The package contains an implementation
-of the evolving graph BFS.
-
+of the evolving graph BFS. \texttt{IntEvolvingGraph}, a data type in EvolvingGraphs.jl,
+represents an evolving graph as adjacency lists. 
+For a random (directed) \texttt{IntEvolvingGraph} with $5$ time stamps (each edge is included in the graph with probability $0.5$), 
+Table \ref{tab:bfs} shows the time taken and 
+the total number of bytes allocated of Algorithm \ref{alg:bfs}.
+\begin{table}[h]
+\centering
+\begin{tabular}{c | c | c | c }
+\hline
+no. of nodes & no. of edges  & time(s) &  allocation(KB) \\\hline
+           10               &        233               & 0.000180   & 43  \\
+            15             &     510                  & 0.000285  &  73  \\
+            20             &   961                   & 0.000431 &  115 \\
+            25             &   1578                 & 0.000642 &   171 \\
+            30             &  2174                  & 0.000819 &  268 \\
+            35             &  3011                  & 0.001066 & 344
+\end{tabular}
+\caption{The time and the total number of bytes allocated for computing BFS
+on a random evolving graph with 5 time stamps. Each row shows the
+number of nodes, the number of edges, the time taken, and the memory requirements.}
+\label{tab:bfs}
+\end{table}
 %Julia is a high-level, high-performance dynamic programming language for technical
@@
 % the parameterizable type system and multiple dispatch in Julia.
-\texttt{MatrixList}, a data type in EvolvingGraphs.jl, represents an evolving graph
+\texttt{IntMatrixList}, a data type in EvolvingGraphs.jl, represents an evolving graph
 as a sequence of sparse matrices stored in compressed sparse column (CSC) format.
-Many basic graph functions are also implemented for \texttt{MatrixList}.
+We can convert an \texttt{IntMatrixList} to its corresponding block adjacency matrix
+using the function \texttt{spmatrix}.
 

### 2016-02-20 paper.tex
update experiments

--- 2016-02-18 paper.tex
@@
 \begin{thm}[Complexity of the evolving graph BFS]
+\label{thm:complexity}
 Let $G_n$ be an evolving graph represented using adjacency lists,
@@
 and available with the MIT ``Expat'' license. The package contains an implementation
-of the evolving graph BFS.
-
+of the evolving graph BFS. \texttt{IntEvolvingGraph}, a data type in EvolvingGraphs.jl,
+represents an evolving graph as adjacency lists.
+We generate a group of  random (directed) \texttt{IntEvolvingGraph}s with $2000$ nodes and $4$ time stamps, where each edge is included in the graph with
+probability $p$.
+ Figure \ref{fig:time} shows the plots of  number of edges against the computation time
+for running Algorithm \ref{alg:bfs} in Julia. All experiments are conducted 
+in a Linux system with Intel(R) Core(TM) i7-3537U CPU @ $2.00$GHz.
+The results show Algorithm \ref{alg:bfs} can be computed in linear time, which agrees
+with Theorem \ref{thm:complexity}. A summary of the experiments is provided in Table \ref{tab:bfs}.
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.44]{time.png}
+  \caption{Experiments of computing Algorithm \ref{alg:bfs} on a group of 
+random evolving 
+graphs with $2000$ nodes and $4$ time stamps. The x-axis shows the 
+number of edges of each evolving graph and the y-axis shows the corresponding
+computation time. The values in both axes increase linearly in the direction of 
+the arrow.}
+  \label{fig:time}
+\end{figure}
+\begin{table}[h]
+\centering
+\begin{tabular}{c | c | c | c }
+\hline
+probability $p$ & no. of edges  & time (s) &  memory (MB) \\\hline
+      0.01  &         160,156      & 0.098   & 18.388  \\
+      0.09  &       1,440,895        & 0.446  &  134.253  \\
+      0.15   &     2,399,774         & 0.743 & 221.511 \\
+      0.23   &    3,676,959          & 1.206 &   337.688 \\
+      0.31   &     4,956,969          & 1.504  &  494.833 \\
+\end{tabular}
+\caption{the computation time and the amount of memory allocated for 
+computing Algorithm \ref{alg:bfs} on a random evolving graph represented 
+adjacency lists with $4$ time stamps and $2000$ nodes.  
+Each edge is included in the graph with probability $p$. 
+This is a summery of Figure \ref{fig:time}.}
+\label{tab:bfs}
+\end{table}
 %Julia is a high-level, high-performance dynamic programming language for technical
@@
 % the parameterizable type system and multiple dispatch in Julia.
-\texttt{MatrixList}, a data type in EvolvingGraphs.jl, represents an evolving graph
-as a sequence of sparse matrices stored in compressed sparse column (CSC) format.
-Many basic graph functions are also implemented for \texttt{MatrixList}.
+% \texttt{IntMatrixList}, a data type in EvolvingGraphs.jl, represents an evolving graph
+% as a sequence of sparse matrices stored in compressed sparse column (CSC) format.
+% We can convert an \texttt{IntMatrixList} to its corresponding block adjacency matrix
+% using the function \texttt{spmatrix}.
 

### 2016-02-20 paper.tex
make figure looks nicer and update caption

@@
 \usepackage{bm}
+\usepackage{tcolorbox}
+\tcbuselibrary{skins}
+\usepackage{lipsum}
 \usepackage[linesnumbered]{algorithm2e}
@@
 
+ \begin{center}
 \begin{figure}[h]
- \begin{center}
+\begin{tcolorbox}[width=3.4in,
+                  %%enhanced,
+                  %%frame hidden,
+                  %interior hidden,
+                  boxsep=0pt,
+                  left=0pt,
+                  right=0pt,
+                  top=4pt,
+                  ]%%
     \begin{tikzpicture}[scale=.31, line width =1.4pt]
@@
     \end{tikzpicture}
+\end{tcolorbox}
+\begin{tcolorbox}[width=3.4in,
+                  %%enhanced,
+                  %%frame hidden,
+                  interior hidden,
+                  boxsep=0pt,
+                  left=0pt,
+                  right=0pt,
+                  top=4pt,
+                  ]%%
    \begin{tikzpicture}[scale=.31, line width =1.4pt]
@@
     \end{tikzpicture}
-\end{center}
+\end{tcolorbox}
 \caption{Breadth-first search (BFS) on the evolving graph shown in
@@
 \end{figure}
-
+\end{center}
 As a simple example, consider the BFS on the example graph in
@@
 number of edges of each evolving graph and the y-axis shows the corresponding
-computation time. The values in both axes increase linearly in the direction of 
-the arrow.}
+computation time. The values in the x, y-axis increase linearly from left to 
+right and from bottom to top respectively.}
   \label{fig:time}

### 2016-02-20 paper.tex
clarify the importance of this paper

@@
 are highly dynamic with arbitrary changes to the nodes and edges.
+
+Although it is well known that sparse matrix-vector product is equivalent to a 
+one-step BFS on the corresponding graph~\cite{kegi11},
+it is little known about the generalization of this duality for evolving graphs.
+In fact, previous works form the off-diagonal blocks wrongly~\cite{fehi15}. 
+We prove the generalization of this duality by finding
+the correct block sparse matrix representation of an evolving graph.
+This work forms the basis for many algebraic evolving graph algorithms.
 

### 2016-02-20 paper.tex
clarify algorithm 1

@@
 The restriction that temporal paths may only traverse active nodes reflects
-underlying structure in many real world applications, such as analyzing the
+underlying causal structure in many real world applications, such as analyzing the
 influence of nodes over social networks. We will also show later in
@@
 
+\subsection{Description of the Algorithm}
+\label{sec:descr-algor}
+
 The BFS on evolving graphs is described in Algorithm~\ref{alg:bfs}.
+Algorithm~\ref{alg:bfs} is identical to the BFS on static graphs expect 
+line $8$, where we visit the forward neighbors of a given temporal node
+in both space and time.
 Given an evolving graph $G_n$ and a root $(v_1, t_1)$,

### 2016-02-20 paper.tex
update the proof of theorem 1

@@
 \begin{thm}[Correctness of the evolving graph BFS]
-Let $G_n$ be a directed evolving graph and $(v_1, t_1)$ be an active node of $G_n$.
+Let $G_n$ be an evolving graph and $(v_1, t_1)$ be an active node of $G_n$.
 Then Algorithm~\ref{alg:bfs} discovers every active node that is reachable from
@@
 \IEEEproof
+Let's first consider the case when $G_n$ is directed.
 Define the set of temporal nodes
@@
 
-The node set $V$ and edge set $E$ now define a static graph $G=(V,E)$ that is in
-1-1 correspondence with the evolving graph $G_n$. The node set $V$ of $G$ is in
-1-1 correspondence with active nodes of $G_n$ while the edge set $E$ is in 1-1
+The node set $V$ and edge set $E$ now define a directed static graph $G=(V,E)$ 
+that is in 1-1 correspondence with the evolving graph $G_n$. The node set $V$ of $G$ is in 1-1 correspondence with active nodes of $G_n$ while the edge set $E$ is in 1-1
 correspondence with all temporal paths of length 2 on $G_n$.
@@
 correspondence with a node in $V$ that is a neighbor of $v_t \in V$.
+
+When $G_n$ is undirected, every edge in $\tilde E^{[t]}$ can be represented by 
+two edges in $G$: from an active node in $\tilde V_L^{[t]}$ to
+an active node in $\tilde V_R^{[t]}$ and the reverse.
+Every edge in $E'$ is in 1-1 correspondence with
+an edge in $G$ by causality. Therefore, the forward neighbors of an 
+active node is in 1-1 correspondence with a subset of $G$ and the analysis above
+follows.
 

### 2016-02-20 paper.tex
add note for causality
add complexity analysis of algebraic BFS

@@
 
-\subsection{Description of the Algorithm}
+\subsection{Description of the BFS algorithm}
 \label{sec:descr-algor}
@@
 \endproof
+
+Note the edge set $E$ is $\cup_t\tilde E^{[t]} \cup E'$, which contains edges that
+are causal structured.
 
@@
 to write down an elegant algebraic formulation of BFS on evolving graphs, as
-presented in Algorithm~\ref{alg:bfsla}. Note that $\bm A_n^T$ need never be formed
-explicitly, as only its matrix--vector product is required in the algorithm.
+presented in Algorithm~\ref{alg:bfsla}.
 
@@
 \endproof
+
+\subsection{Computational complexity analysis of the algebraic BFS}
+\label{sec:comp-compl-analys}
+
+Given an evolving graph $G_n = \langle G^{[1]}, G^{[2]}, \ldots, G^{[n]} \rangle$, let $V$ be the set of active nodes,
+$E = \cup_t \tilde E^{[t]} \cup E'$ be the set of edges, where
+$\tilde E^{[t]}$ represents the set of edges of $G^{[t]}$ and $E'$ represents the set
+of causal edges between time stamps. Let $\bm A_n$ be the corresponding
+adjacency matrix of $G_n$.
+
+Let's analyze the running time of Algorithm \ref{alg:bfsla}. 
+Suppose $\bm A_n$ is dense.
+Forming the block adjacency matrix $\bm A_n^T$ in line $3$
+requires $\Theta(|V|^2)$ operations. 
+We note, however, $\bm A_n^T$ need never be formed
+explicitly, as only its matrix-vector product is required in the algorithm.
+The matrix-vector product in line $7$ requires $O(|V|^2)$ operations.
+In the worst case, we need to reach all active nodes and reach only one 
+active node at each iteration. 
+So the number of iterations $k$ of the while loop at line $6$ is bounded by
+$|E|$, i.e., $k = \Theta(|E|)$. Hence, the total running time is 
+$\Theta(|E||V|^2)$. This is much worse than Algorithm \ref{alg:bfs}, which
+has linear running time.
 

### 2016-02-20 paper.tex
more analysis of algebraic BFS
update experiments

@@
 
+If $\bm A_n$ is sparse, forming $\bm A_n$ costs $\Theta(|E|)$ flops. 
+The gaxpy operation in line $7$ costs 
+$2m$ flops, where $m$ is the number of none zeros of $\bm A_n$ \cite[Sec.11]{golo13}. 
+Here $m$ is equal to $|E|$. Thus, in the worst case, the total running time is 
+$\Theta(|E|^2)$.
+
+% Storing $\bm A_n$ requires $\Theta(|V|^2)$ memory and if $\bm A_n$ is sparse, 
+% we need $O(|V|+|E|)$ memory.
+
 \section{Implementation in Julia}
@@
 represents an evolving graph as adjacency lists.
-We generate a group of  random (directed) \texttt{IntEvolvingGraph}s with $2000$ nodes and $4$ time stamps, where each edge is included in the graph with
+We generate a group of  random (directed) \texttt{IntEvolvingGraph}s with $8,000$ active nodes and $4$ time stamps, where each edge is included in the graph with
 probability $p$.
  Figure \ref{fig:time} shows the plots of  number of edges against the computation time
-for running Algorithm \ref{alg:bfs} in Julia. All experiments are conducted 
+for running Algorithm \ref{alg:bfs}
+in Julia. All experiments are conducted 
 in a Linux system with Intel(R) Core(TM) i7-3537U CPU @ $2.00$GHz.
@@
   \includegraphics[scale=0.44]{time.png}
-  \caption{Experiments of computing Algorithm \ref{alg:bfs} on a group of 
-random evolving 
-graphs with $2000$ nodes and $4$ time stamps. The x-axis shows the 
+  \caption{Experiments of running Algorithm \ref{alg:bfs}
+on a group of random evolving graphs with $8,000$ active nodes and $4$ time stamps. The x-axis shows the
 number of edges of each evolving graph and the y-axis shows the corresponding
@@
 \caption{the computation time and the amount of memory allocated for 
-computing Algorithm \ref{alg:bfs} on a random evolving graph represented 
-adjacency lists with $4$ time stamps and $2000$ nodes.  
-Each edge is included in the graph with probability $p$. 
-This is a summery of Figure \ref{fig:time}.}
+computing Algorithm \ref{alg:bfs}. This is a summery of Figure \ref{fig:time}.  
+Each edge is included in the graph with probability $p$. }
 \label{tab:bfs}

### 2016-02-22 paper.tex
use pdf for timing figure
clarify the generation of random evolving graphs

@@
 represents an evolving graph as adjacency lists.
-We generate a group of  random (directed) \texttt{IntEvolvingGraph}s with $8,000$ active nodes and $4$ time stamps, where each edge is included in the graph with
-probability $p$.
+We generate a group of  random (directed) \texttt{IntEvolvingGraph}s
+with $8,000$ active nodes and $4$ time stamps.
+At each time stamp $i$, an edge is included in $G^{[i]} \in G_4$ with probability $p$.
  Figure \ref{fig:time} shows the plots of  number of edges against the computation time
@@
   \centering
-  \includegraphics[scale=0.44]{time.png}
+  \includegraphics[scale=0.44]{time.pdf}
   \caption{Experiments of running Algorithm \ref{alg:bfs}

### 2016-02-22 paper.tex
add book section

@@
 Although it is well known that sparse matrix-vector product is equivalent to a 
-one-step BFS on the corresponding graph~\cite{kegi11},
+one-step BFS on the corresponding graph~\cite[Sec. 1.1]{kegi11},
 it is little known about the generalization of this duality for evolving graphs.

### 2016-02-22 paper.tex
One last round of changes

--- 2016-02-22 paper.tex
@@
 % correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
+\hyphenation{}
 
@@
 \begin{abstract}
-Evolving graphs arise in many different contexts where the
-interrelations between data elements change over time.
-We present a breadth first search (BFS) algorithm for evolving graphs
-that can track (active) nodes correctly.
-Using simple examples, we show that na\"ive matrix-matrix multiplication on
-time-dependent adjacency matrices miscounts the number of temporal paths.
-By mapping an evolving graph to an adjacency matrix of the equivalent static graph,
-we prove the properties of the BFS algorithm using the properties of the
-adjacency matrix. Finally, demonstrate how the BFS over evolving graphs can be
-applied to mining citation networks.
+Evolving graphs arise in problems where interrelations between data change over
+time. We present a breadth first search (BFS) algorithm for evolving graphs
+that computes the most direct influences between nodes at two different times.
+Using simple examples, we show that na\"ive unfoldings of adjacency matrices
+miscount the number of temporal paths.
+By mapping an evolving graph to an adjacency matrix of an equivalent static graph,
+we prove that our generalization of the BFS algorithm is correct.
+Finally, we demonstrate how the BFS over evolving graphs can be applied to mine
+citation networks.
 
@@
 
-Although it is well known that sparse matrix-vector product is equivalent to a 
-one-step BFS on the corresponding graph~\cite{kegi11},
-it is little known about the generalization of this duality for evolving graphs.
-In fact, previous works form the off-diagonal blocks wrongly~\cite{fehi15}. 
+It is well known that sparse matrix-vector product is equivalent to a
+one-step BFS on the corresponding graph~\cite[Sec. 1.1]{kegi11}.
+However, little is known about the generalization of this duality for evolving graphs.
 We prove the generalization of this duality by finding
 the correct block sparse matrix representation of an evolving graph.
-This work forms the basis for many algebraic evolving graph algorithms.
 
@@
 This definition of a temporal path differs from that of the dynamic walk
-in \cite{gphe11,grihig13} in its explicit requirement of containing only active
-nodes. The definition implies that if either or both end points of a temporal
+in \cite{gphe11,grihig13} in that \textbf{causal edges}, i.e.\ sequences of the
+form $\langle (v, t), (v, t\prime)\rangle$ are included explicitly in temporal
+paths but are only implicitly included in dynamic walks and are not counted
+toward the length of dynamic walks.
+Our definition implies that if either or both end points of a temporal
 path are inactive, then the entire temporal path must be the empty sequence
@@
 
+Our definition of distance, again, differs from the definition of distance
+in the formulation of \cite{gphe11,grihig13} in that we explicitly count causal
+edges toward the distance. It also differs from the notion of temporal distance
+in the work of Tang and coworkers~\cite{tmml10}, which is the number of time
+steps between $t$ and $s$ (inclusive). In this respect, our formulation of the
+BFS on evolving graphs differs from these other works by minimizing a different
+notion of distance over an evolving graph.
+
 Note that this notion of distance is not a metric, since the distance from $(v, t)$
@@
 The BFS on evolving graphs is described in Algorithm~\ref{alg:bfs}.
-Algorithm~\ref{alg:bfs} is identical to the BFS on static graphs expect 
+Algorithm~\ref{alg:bfs} is identical to the BFS on static graphs except for
 line $8$, where we visit the forward neighbors of a given temporal node
@@
 
-Similarly, define an edge set $E' = \{(u_s,v_t)|u_s=(u,s)\in V,
+Similarly, define the set of \textbf{causal edges} $E' = \{(u_s,v_t)|u_s=(u,s)\in V,
 v_t=(v,t)\in V, v=u, s<t\}$, which consists of temporal nodes that connect
@@
 $\langle (v,s), (v,t) \rangle$.
-Define also $\tilde E^{[t]} = \{(e, t)|e\in E^{[t]}\}$, which are simply the edge
-sets in $G_n$ with time labels. Then $E = \bigcup_t \tilde E^{[t]} \cup E'$ is
-the set of all edges representing all allowed temporal paths of length 2.
+Define also the set of \textbf{static edges at time $t$},
+$\tilde E^{[t]} = \{(e, t)|e\in E^{[t]}\}$, which are simply the edge
+sets in $G_n$ with time labels, and the set of \textbf{static edges} $\tilde E$,
+being simply the union over all times, $\bigcup_t \tilde E^{[t]}$.
+Then $E = \tilde E \cup E'$ is the set of all edges representing all allowed
+temporal paths of length 2.
 
@@
 We now establish a similar 1-1 correspondence of forward neighbors of an active
-node with a subset of $G$. By induction, we can show that at iteration $k$,
-all new nodes populated into the key set of $reached$ by definition are of
-distance $k$ from the root. By definition,
+node with a subset of $G$. By induction, all new nodes populated into the key
+set of $reached$ at iteration $k$ are of distance $k$ from the root. By definition,
 the forward neighbors of some active node $(v, t) \in G_n$ are
 active nodes of either the form $(v, t')$ for some $t' > t$ or $(u, t)$ for some
-$u\ne v$. Clearly, the former are elements of $E' \subseteq E$ while the latter
-are elements of $E^{[t]} \subseteq \bigcup_s \tilde E^{[s]} \subseteq E$.
+$u\ne v$. In other words, they are connected either by a causal edge or a static edge.
+Clearly, the former are elements of $E' \subseteq E$ while the latter
+are elements of $\tilde E \subseteq E$.
 Thus each forward neighbor of an active node $(v, t) \in G_n$ is in 1-1
@@
 
-\begin{thm}[Complexity of the evolving graph BFS]
+\begin{thm}[Computational complexity of the evolving graph BFS]
 \label{thm:complexity}
@@
 constructed from $G_n$ using the 1-1 correspondences defined in the proof of
-Theorem~\ref{thm:correct}. Then the asymptotic complexity of
+Theorem~\ref{thm:correct}. Then the asymptotic computational complexity of
 Algorithm~\ref{alg:bfs} is $O(|E| + |V|)$.
@@
 
-Note the edge set $E$ is $\cup_t\tilde E^{[t]} \cup E'$, which contains edges that
-are causal structured.
+Note that in the theorems in this section construct an equivalent static graph
+$G$ corresponding to the evolving graph $G_n$. However, $G$ contains more edges
+than the union of all the static parts of $G_n$, as we also add causal edges
+$E'$. To our knowledge, our formulation of the BFS represents the first attempt
+to include these edges explicitly in the treatment of evolving graphs.
 
@@
 
-\subsection{Not all temporal paths are counted by products of adjacency matrices}
+\subsection{The importance of causal edges}
 \label{sec:temp-paths-adjac}
@@
 $\left\vert V^{[t]}\right\vert \times \left\vert V^{[t]}\right\vert$
-adjacency matrix with elements
+one-sided adjacency matrix with elements
 \begin{equation}
@@
 Sums $S^{[t]}$ of the form \eqref{eq:sum} produce an incorrect count of temporal
-paths because they do not capture temporal paths with subpaths of the form
-$\langle (v, s), (v, t) \rangle$, $s < t$.
+paths because they do not capture temporal paths with causal edges, i.e.\ subpaths
+of the form $\langle (v, s), (v, t) \rangle$, $s < t$.
 One might attempt to amend the sums $S^{[t]}$ in \eqref{eq:sum} by redefining
@@
 evolving graphs, and sums like \eqref{eq:sum} cannot form the basis of correct
-analyses of evolving graphs.
+analyses of evolving graphs as they neglect the combinatorics associated with
+the causal edge set $E'$.
 
@@
 \]
+The condition $A^Tb \ne 0$ ensures that the product is nonzero in components
+involving left active nodes $\cup_t \tilde V_L^{[t]}$, and the condition
+$Ab \ne 0$ is the analogue for right active nodes $\cup_t \tilde V_R^{[t]}$.
 The forward neighbors of a temporal node $(k, t_1)$ in $A_n$ can then be determined
@@
 %
-The adjacency matrix blocks $A^{[t]}$ encode the edge set $\tilde E^{[t]}$,
-whereas the off-diagonal blocks $M^{[t_i, t_j]}$ together encode the edge set
-$E^\prime$ which capture temporal paths with subsequences of the form
+The adjacency matrix blocks $A^{[t]}$ encode the static edge set $\tilde E$,
+whereas the off-diagonal blocks $M^{[t_i, t_j]}$ together encode the causal edge
+set $E^\prime$, which capture temporal paths with subsequences of the form
 $\langle (v, t_i), (v, t_j) \rangle$. Then $\bm M_n$ is the adjacency matrix of
@@
 
-Finally, we prove a simple lemma that the block adjacency matrix $\bm A_n$ is
+Finally, we note that some results regarding the BFS on evolving graphs can be
+derived easily using properties of the block adjacency matrix $\bm A_n$. For
+example, we can prove a simple lemma that the block adjacency matrix $\bm A_n$ is
 nilpotent whenever all the subgraphs $G^{[t]}$ of $G_n$ are acyclic.
@@
 \IEEEproof
+Recall from the definition of the matrix $\bm A_n$ that it is block upper
+triangular, reflecting causality.
 Since each directed graph $G^{[t]}$ is acyclic, its corresponding adjacency
@@
 repeated power iterations of the adjacency matrix in Algorithm~\ref{alg:bfsla}.
-Note, however, that these matrices never need to be instantiated for practical
+Note, however, that these matrices need never be instantiated for practical
 computations. Rather, since Algorithm~\ref{alg:bfsla} only requires the
@@
 action of the off-diagonal products. Representing the diagonal blocks
-$A^{[t]}$ as sparse matrices product further reduces the cost of BFS by exploiting
+$A^{[t]}$ as sparse matrices further reduces the cost of BFS by exploiting
 latent sparsity in graphs that show up in practical applications.
@@
 \IEEEproof
+First, we prove that the BFS terminates in the case of acyclic evolving graphs.
 Recall from Lemma~\ref{thm:nilpotent} that $\bm A_n$ is nilpotent, i.e.\
@@
 Therefore, Algorithm~\ref{alg:bfsla} must terminate after iteration $k$.
+
+For evolving graphs with cycles, lines 9-11 of Algorithm \ref{alg:bfsla} enforce
+that the BFS visits each active node at most once. Since the $k$th block of
+$\bm b$ is zeroed out if an active node has already been visited, the subgraph
+traversed in the BFS cannot have cycles. Thus all that is required is the
+previous result that the BFS on an acyclic graph terminates.
 \endproof
@@
 
-Given an evolving graph $G_n = \langle G^{[1]}, G^{[2]}, \ldots, G^{[n]} \rangle$, let $V$ be the set of active nodes,
-$E = \cup_t \tilde E^{[t]} \cup E'$ be the set of edges, where
-$\tilde E^{[t]}$ represents the set of edges of $G^{[t]}$ and $E'$ represents the set
-of causal edges between time stamps. Let $\bm A_n$ be the corresponding
-adjacency matrix of $G_n$.
-
-Let's analyze the running time of Algorithm \ref{alg:bfsla}. 
-Suppose $\bm A_n$ is dense.
-Forming the block adjacency matrix $\bm A_n^T$ in line $3$
-requires $\Theta(|V|^2)$ operations. 
-We note, however, $\bm A_n^T$ need never be formed
-explicitly, as only its matrix-vector product is required in the algorithm.
-The matrix-vector product in line $7$ requires $O(|V|^2)$ operations.
-In the worst case, we need to reach all active nodes and reach only one 
-active node at each iteration. 
-So the number of iterations $k$ of the while loop at line $6$ is bounded by
-$|E|$, i.e., $k = \Theta(|E|)$. Hence, the total running time is 
-$\Theta(|E||V|^2)$. This is much worse than Algorithm \ref{alg:bfs}, which
-has linear running time.
-
-If $\bm A_n$ is sparse, forming $\bm A_n$ costs $\Theta(|E|)$ flops. 
-The gaxpy operation in line $7$ costs 
-$2m$ flops, where $m$ is the number of none zeros of $\bm A_n$ \cite[Sec.11]{golo13}. 
-Here $m$ is equal to $|E|$. Thus, in the worst case, the total running time is 
-$\Theta(|E|^2)$.
+The complexity of the algebraic BFS of Algorithm \ref{alg:bfsla} is significantly
+more complicated that that of Algorithm \ref{alg:bfs}. While the latter uses the
+usual adjacency list representation for graphs, the computational cost of the
+former depends critically on the actual representation of the matrices.
+Furthermore, the average case analysis is complicated by the expected fill-in of
+the vector $\bm b$, which influences the cost of the matrix-vector product on line 7
+and the expected number of iterations of the \textbf{while} loop beginning on line6.
+
+While a full complexity analysis is beyond the scope of this paper, it is
+straightforward to present worst-case results for dense and compressed sparse
+column (CSC) matrices.
+
+\begin{lemma}[Number of iterations]
+In the worst case, the number of iterations in the \textbf{while} loop of
+Algorithm~\ref{alg:bfsla} is $k = O(|E|)$.
+\end{lemma}
+
+\IEEEproof
+In the worst case, the BFS must traverse every active node, and only one new
+active node is discovered in each iteration. The number of active
+nodes is bounded above by the cardinality of the full edge set, $|E|$.
+\endproof
+
+The average case analysis for the number of iterations is considerably more
+complicated and is beyond the scope of this paper.
+
+\begin{thm}[Dense matrices]
+Suppose $\bm A_n$ is represented as a dense matrix. Then the computational
+complexity of Algorithm~\ref{alg:bfsla} is $O(k |V|^2)$, which in the worst case is
+$O(|E| |V|^2)$.
+\end{thm}
+
+\IEEEproof
+Since $\bm A_n$ is a $|V| \times |V|$ matrix, the matrix-vector product
+$\bm A_n \bm b$ takes $O(|V|^2)$ operations to compute. Thus the cost of
+Algorithm~\ref{alg:bfsla} is $O(k |V|^2) = O(|E| |V|^2)$ in the worst case.
+\endproof
+
+It is clear that practical implementations of the BFS should never construct
+the full matrix $\bm A_n$ in memory. What happens if we use a sparse blocked
+representation?
+
+\begin{thm}[Block diagonal sparse matrices]
+Suppose $\bm A_n$ is represented by a collection of compressed sparse column
+matrices for each diagonal block $A^{[t]}$. Then the
+computational complexity of Algorithm~\ref{alg:bfsla} is $O(k(|\tilde E| + |V|))$, which in
+the worst case is $O(|E|(|\tilde E| + |V|))$.
+\end{thm}
+
+\IEEEproof
+The gaxpy operation for CSC matrices costs $2n_{nz}$ flops, where $n_{nz}$ is the
+number of stored values in the matrix. The cost of each diagonal subblock calculation
+$A^{[t]} b_t$ is therefore $O(|E^{[t]}|)$, since $A^{[t]}$ by construction has
+nonzero entries only when a static edge exists.
+
+The off-diagonal products
+$M^{[t,t\prime]} b_{t\prime}$ can be computed in $O(|V^{[t]}| + |E^{[t]}|)$ time
+for all $t\prime \ge t$
+since it can be implemented by the $\cdot$ operation, which constructs either a
+zero vector or keeps the same vector.
+The cost of checking the condition $(A^{[t]})^T b_{t^\prime} \ne 0$ is
+$O(|E^{t}|)$ in the worst case since all that is required is to check whether or
+not each column of $A$ is empty.
+Similarly, checking the condition $A^{[t]} b_{t^\prime} \ne 0$ reduces to
+checking if each row of $A$ is empty, and thus is of cost $O(|V^{t}|)$.
+
+Thus, the cost of multiplying one block row of $\bm A^T$ (for some time $t$)
+with $\bm b$ is $O(|V^{[t]}| + |E^{[t]}|)$. Summing over all times yields the
+desired result.
+\endproof
+
+We can see that even implementing the BFS algebraically using CSC matrices is
+insufficient to reduce the running time to linear, which can be achieved for
+the adjacency list representation in Algorithm \ref{alg:bfs}.
+This result strongly suggests that additional work is needed to produce true
+algorithmic equivalence at the computational level.
 
@@
 and available with the MIT ``Expat'' license. The package contains an implementation
-of the evolving graph BFS. \texttt{IntEvolvingGraph}, a data type in EvolvingGraphs.jl,
+of the evolving graph BFS of Algorithm~\ref{alg:bfs}. \texttt{IntEvolvingGraph}, a data type in EvolvingGraphs.jl,
 represents an evolving graph as adjacency lists.
+
+We now present some simple timing data to show that our implementation of
+Algorithm~\ref{alg:bfs} is indeed linear scaling in computational cost.
+
 We generate a group of  random (directed) \texttt{IntEvolvingGraph}s
@@
 Figure \ref{fig:time} shows the plots of number of edges against the computation time
-for running Algorithm \ref{alg:bfs}
-in Julia. All experiments are conducted 
-in a Linux system with Intel(R) Core(TM) i7-3537U CPU @ $2.00$GHz.
+for running Algorithm \ref{alg:bfs} in Julia. All experiments are conducted
+on a Linux system with Intel(R) Core(TM) i7-3537U CPU @ $2.00$GHz.
+
 The results show Algorithm \ref{alg:bfs} can be computed in linear time, which agrees
@@
 
-In this section, we show that evolving graph formalism presented above can be
-used to capture the structure of citation networks. Consider
+Evolving graphs have found many applications to analyzing networks that change
+over time \cite{gphe11,grihig13}. In this section, we focus specifically on
+citation networks, and show that evolving graph formalism presented above can be
+used to capture the dynamical structure of citation networks. Consider
 the evolving graph $G_n=\langle G^{[t]} \rangle_t$ such that $G^{[t]}$ has
@@
 
+We are currently investigating the use of our evolving graph BFS on citation
+networks.
+
 \section{Conclusion}
@@
 The correct generalization of BFS to evolving graphs necessitates a careful
-enumeration of temporal paths which is not always captured by products of
-successive adjacency matrices from the snapshots. BFS over any evolving graph $G_n$
-is formally equivalent to BFS on a related static graph $G$ formed by connecting only
-active nodes across different times. The adjacency matrix $\bm A_n$ of $G$ correctly
-encodes causality of the temporal paths. When specialized to acyclic $G_n$, the matrix
-$\bm A_n$ is also nilpotent, thus providing a simple proof that BFS terminates.
-Nevertheless, the BFS can be computed efficiently without ever explicitly constructing
-$\bm A_n$, as only its action on a vector is required for the power iterations.
-The new concepts of activeness, temporal paths, and the explicit construction of
-an equivalent static graph $G$ that make possible a correct implementation of BFS
-to evolving graphs provide powerful new insignts into how similar graphical
+enumeration of temporal paths. The structure associated with causal edges $E'$
+turns out to be of vital importance, and cannot be capture simply by products of
+successive adjacency matrices, which by construction can only capture the
+topologies of the static edges $\tilde E$. Only by considering both causal edges
+and static edges can we show that BFS over any evolving graph $G_n$ computes the
+correct result for our notion of distance.
+The new concepts of activeness, temporal paths, and causal edges make possible a
+correct implementation of BFS to evolving graphs and we expect that these
+ideas will continue to provide powerful new insights into how similar graphical
 algorithms may be generalized correctly.
 
-% This paper introduced the generalization of BFS to evolving graphs and presented
-% an equivalent formulation using linear algebra. The concept of temporal paths
-% connecting active nodes turns out to be critical for the correct generalization
-% of BFS to evolving graphs. An efficient implementation of BFS on evolving
-% graphs is now available as a open source Julia package.
+Furthermore, we show that BFS on evolving graphs admits an algebraic formulation
+that easily provides nontrivial results, such as termination of the algorithm.
+However, our current understanding tells us that the BFS over evolving graphs
+is most efficiently computed in the adjacency list representation, thus never
+forming explicit matrix-vector products. Further work is needed to elucidate
+more efficient formulations of the algebraic BFS for evolving graphs.
+
 % This paper leaves important extensions open for further research. First, extensions

### 2016-02-22 paper.tex
Remove table

@@
 with $8,000$ active nodes and $4$ time stamps.
-At each time stamp $i$, an edge is included in $G^{[i]} \in G_4$ with probability $p$.
+At each time stamp $i$, a static edge is included in $G^{[i]} \in G_4$ with
+probability $p$. Note that in this experiment, we do not have direct control
+over the full edge set $E$, only the static edge set $\tilde E$. However, the
+causal edge set $E'$ can grow at most by $O(\tilde E)$, so it suffices to demonstrate
+linear scaling in $|\tilde E|$.
 Figure \ref{fig:time} shows the plots of number of edges against the computation time
@@
 The results show Algorithm \ref{alg:bfs} can be computed in linear time, which agrees
-with Theorem \ref{thm:complexity}. A summary of the experiments is provided in Table \ref{tab:bfs}.
+with the result of Theorem \ref{thm:complexity}.
+
 \begin{figure}[h]
@@
   \includegraphics[scale=0.44]{time.pdf}
-  \caption{Experiments of running Algorithm \ref{alg:bfs}
-on a group of random evolving graphs with $8,000$ active nodes and $4$ time stamps. The x-axis shows the
-number of edges of each evolving graph and the y-axis shows the corresponding
-computation time. The values in the x, y-axis increase linearly from left to
-right and from bottom to top respectively.}
+  \caption{Experimental run time and memory use of Algorithm \ref{alg:bfs}
+on a collection of random evolving graphs with $8,000$ active nodes and $4$ time
+stamps, showing linear scaling in $|\tilde E|$. The horizontal axis shows the
+number of edges ($|\tilde E|$, including both static and causal edges) of each evolving
+graph, while the vertical axis shows the corresponding
+computation time.}
   \label{fig:time}
 \end{figure}
-\begin{table}[h]
-\centering
-\begin{tabular}{c | c | c | c }
-\hline
-probability $p$ & no. of edges  & time (s) &  memory (MB) \\\hline
-      0.01  &         160,156      & 0.098   & 18.388  \\
-      0.09  &       1,440,895        & 0.446  &  134.253  \\
-      0.15   &     2,399,774         & 0.743 & 221.511 \\
-      0.23   &    3,676,959          & 1.206 &   337.688 \\
-      0.31   &     4,956,969          & 1.504  &  494.833 \\
-\end{tabular}
-\caption{the computation time and the amount of memory allocated for
-computing Algorithm \ref{alg:bfs}. This is a summery of Figure \ref{fig:time}.
-Each edge is included in the graph with probability $p$. }
-\label{tab:bfs}
-\end{table}
 %Julia is a high-level, high-performance dynamic programming language for technical

### 2016-02-26 paper.tex
fix typo

--- 2016-02-22 paper.tex
@@
 the vector $\bm b$, which influences the cost of the matrix-vector product on line 7
-and the expected number of iterations of the \textbf{while} loop beginning on line6.
+and the expected number of iterations of the \textbf{while} loop beginning on line 6.
 

### 2016-03-04 paper.tex
add new experiments and update comments

@@
 We generate a group of random (directed) \texttt{IntEvolvingGraph}s
-with $8,000$ active nodes and $4$ time stamps.
-At each time stamp $i$, a static edge is included in $G^{[i]} \in G_4$ with
-probability $p$. Note that in this experiment, we do not have direct control
+with $1.0\times 10^5$ active nodes and $10$ time stamps.
+The first random \texttt{IntEvolvingGraph} in the group has 
+$1.0 \times 10^8$ static edges.
+We consecutively add $5.0\times 10^7$
+random static edges to this \texttt{IntEvolvingGraph}. For example, 
+the second random \texttt{IntEvolvingGraph} in the group has about $1.5 \times 10^8$ static edges and the third
+has about $2.0 \times 10^8$ static edges.
+% At each time stamp $i$, a static edge is included in $G^{[i]} \in G_4$ with
+% probability $p$. 
+Note that in this experiment, we do not have direct control
 over the full edge set $E$, only the static edge set $\tilde E$. However, the
@@
 for running Algorithm \ref{alg:bfs} in Julia. All experiments are conducted
-on a Linux system with Intel(R) Core(TM) i7-3537U CPU @ $2.00$GHz.
-
+on a Linux system with Intel(R) Xeon(R) CPU E7-8850 @ $2.00$GHz.
 The results show Algorithm \ref{alg:bfs} can be computed in linear time, which agrees
 with the result of Theorem \ref{thm:complexity}.
+By doing this experiment, we find that it takes longer and longer time to 
+add new static edges, since we also need to include
+all the causal edges introduced by these new static edges.
 
@@
   \centering
-  \includegraphics[scale=0.44]{time.pdf}
+  \includegraphics[scale=0.58]{time.pdf}
   \caption{Experimental run time and memory use of Algorithm \ref{alg:bfs}
-on a collection of random evolving graphs with $8,000$ active nodes and $4$ time
+on a collection of random evolving graphs with $1.0\times 10^5$ active nodes and $10$ time
 stamps, showing linear scaling in $|\tilde E|$. The horizontal axis shows the
-number of edges ($|\tilde E|$, including both static and causal edges) of each evolving
+number of edges ($|\tilde E|$, including only the causal edges) of each evolving
 graph, while the vertical axis shows the corresponding

### 2016-03-04 paper.tex
fix figure caption

@@
 stamps, showing linear scaling in $|\tilde E|$. The horizontal axis shows the
-number of edges ($|\tilde E|$, including only the causal edges) of each evolving
+number of edges ($|\tilde E|$, including only the static edges) of each evolving
 graph, while the vertical axis shows the corresponding

### 2016-03-04 paper.tex
update new timing figure

@@
 
-We generate a group of random (directed) \texttt{IntEvolvingGraph}s
+We generate a sequence of random (directed) \texttt{IntEvolvingGraph}s
 with $1.0\times 10^5$ active nodes and $10$ time stamps.
-The first random \texttt{IntEvolvingGraph} in the group has 
+The first \texttt{IntEvolvingGraph} in the sequence has about
 $1.0 \times 10^8$ static edges.
-We consecutively add $5.0\times 10^7$
+We consecutively add new
 random static edges to this \texttt{IntEvolvingGraph}. For example, 
-the second random \texttt{IntEvolvingGraph} in the group has about $1.5 \times 10^8$ static edges and the third
-has about $2.0 \times 10^8$ static edges.
+the second random \texttt{IntEvolvingGraph} in the sequence 
+has about $1.5 \times 10^8$ static edges and the third
+has $1.8 \times 10^8$ static edges.
 % At each time stamp $i$, a static edge is included in $G^{[i]} \in G_4$ with
@@
 with the result of Theorem \ref{thm:complexity}.
-By doing this experiment, we find that it takes longer and longer time to 
-add new static edges, since we also need to include
-all the causal edges introduced by these new static edges.
 

### 2016-03-04 paper.tex
remove memory use info

@@
   \includegraphics[scale=0.58]{time.pdf}
-  \caption{Experimental run time and memory use of Algorithm \ref{alg:bfs}
+  \caption{Experimental run time of Algorithm \ref{alg:bfs}
 on a collection of random evolving graphs with $1.0\times 10^5$ active nodes and $10$ time

### 2016-03-04 paper.tex
Final round of edits

@@
 By mapping an evolving graph to an adjacency matrix of an equivalent static graph,
-we prove that our generalization of the BFS algorithm is correct.
+we prove that our generalization of the BFS algorithm correctly accounts for
+paths that traverse both space and time.
 Finally, we demonstrate how the BFS over evolving graphs can be applied to mine
@@
 to the study of metrized graphs by Tang and coworkers~\cite{ntmm13,tmml09,tsmm09,tmml10} and of
-the dynamics of communities by Grindrod, Higham and coworkers~\cite{gphe11,grihig13}.
-
-The game described in the beginning can be encoded in an evolving graph.
+community dynamics by Grindrod, Higham and coworkers~\cite{gphe11,grihig13}.
+
+The game described above can be encoded in an evolving graph.
 The spread of information to the winner can be described in terms of traversing
@@
 However, it turns out that na\"ive extensions can lead to incorrect descriptions of
-the resulting graph traversals. A proper treatment requires the notion of
+the resulting graph traversals by accounting for traversals of edges in space,
+but not necessarily in time. A proper treatment requires the notions of
 \emph{node activeness} to describe the set of paths that can only traverse time
-or edges, which we call \textit{temporal paths}.
+or edges, which we call \textit{temporal paths}, as well as \emph{causal edges}
+which connect active nodes with the same identity across different times.
 As a result, our treatment can be applied to any evolving graph, even those that
@@
 It is well known that sparse matrix-vector product is equivalent to a
-one-step BFS on the corresponding graph~\cite[Sec. 1.1]{kegi11}.
-However, little is known about the generalization of this duality for evolving graphs.
-We prove the generalization of this duality by finding
-the correct block sparse matrix representation of an evolving graph.
-
-In Section~\ref{sec:breadth-first-search}, we explain how the BFS algorithm can be applied correctly on
-evolving graphs. Section~\ref{sec:temporal-paths} provides an example showing that considering
-only products of the time-dependent adjacency matrices can lead to incorrect enumeration
-of temporal paths. We present and demonstrate the BFS algorithm over evolving graphs
-in Section~\ref{sec:evolving-graph-bfs}, showing that it is formally equivalent to
-BFS over a particular static graph generated by connecting
-together active notes of the evolving graph. This static graph generates an algebraic
+one-step BFS on the corresponding (static) graph~\cite[Sec. 1.1]{kegi11}.
+In this paper, we demonstrate a correct corresponding result for an evolving
+graph by constructing a block triangular matrix representation of the graph that
+takes into account both static and causal connections between active nodes.
+In Section~\ref{sec:breadth-first-search}, we explain how the BFS algorithm can
+be applied to evolving graph to enumerate paths that traverse edges across time
+and space. Section~\ref{sec:temporal-paths} provides an example showing that considering
+only products of the time-dependent adjacency matrices fails to enumerate certain
+temporal paths. We present and demonstrate the BFS algorithm over evolving graphs
+in Section~\ref{sec:evolving-graph-bfs}, showing its formal equivalence to
+BFS over a particular static graph generated by adding causal edges that connect
+active nodes. This static graph generates an algebraic
 representation of the BFS as power iteration of its adjacency matrix to a starting
@@
 demonstrates interesting connections between properties of the BFS algorithm and
-the adjacency matrix. Finally in Section~\ref{sec:applications},
+the adjacency matrix. We describe in Section~\ref{sec:implementation-julia} an
+implementation of the algorithm in Julia. Finally in Section~\ref{sec:applications},
 we explain how BFS on evolving graphs may be applied to study dynamical processes
@@
 
-The simple example of Figure~\ref{fig:eg_shortest_path} provides several
-counterexamples which demonstrate why sums over products of adjacency matrices of
-the form \eqref{eq:sum} do not count temporal paths correctly. The interpretation
-from considering powers of $A$ for static graphs therefore does not extend to
-evolving graphs, and sums like \eqref{eq:sum} cannot form the basis of correct
-analyses of evolving graphs as they neglect the combinatorics associated with
-the causal edge set $E'$.
+The simple example of Figure~\ref{fig:eg_shortest_path} demonstrates why sums
+over products of adjacency matrices of the form \eqref{eq:sum} do not count
+temporal paths correctly: they neglect the combinatorics associated with
+the causal edge set $E'$. In the next section, we show how to account for these
+causal edges by introducing a new matrix--vector product $\ocirc$.
 
@@
 red nodes are inactive nodes.
-The black lines are edges in $\cup_t E^{[t]}$ and are encoded
+The black lines are edges in the static edge set $\tilde E$ and are encoded
 algebraically in the diagonal blocks $A^{[t]}$ of the adjacency matrix $\bm A_3$
 or $\bm M_3$.
-The dotted lines are edges in $E'$ and are encoded algebraically in
+The dotted lines are edges in the causal edge set $E'$ and are encoded algebraically in
 the off-diagonal blocks $M^{[t_i, t_j]}$.
@@
 V & = \{(1,\!t_1), (2,\!t_1), (1,\!t_2), (3,\!t_2), (2,\!t_3), (3,\!t_3)\},\\
-\cup_t\!E^{[t]} &= \{((1,\!t_1), (2,\!t_1)), ((1,\!t_2), (3,\!t_2)), ((2,\!t_3), (3,\!t_3))\},\\
+\tilde E &= \{((1,\!t_1), (2,\!t_1)), ((1,\!t_2), (3,\!t_2)), ((2,\!t_3), (3,\!t_3))\},\\
 E'  &= \{((1,\!t_1), (1,\!t_2)), ((2,\!t_2), (2,\!t_3)), ((3,\!t_2), (3,\!t_3))\}.
@@
 
-We see that $(\bm A_n^T)^2$ correctly encodes the collection of products in
-\eqref{eq:5}-\eqref{eq:6}. Furthermore, $(\bm A_3^T)^3 \bm b$ correctly counts the
+We see that $(\bm A_n^T)^2$ encodes all the products in
+\eqref{eq:5}-\eqref{eq:6}, including both the ordinary matrix--vector product
+and the $\ocirc$ product. Furthermore, $(\bm A_3^T)^3 \bm b$ correctly counts the
 two allowed temporal paths from $(1, t_1)$ to $(3, t_3)$, and that the
@@
 We generate a sequence of random (directed) \texttt{IntEvolvingGraph}s
-with $1.0\times 10^5$ active nodes and $10$ time stamps.
+with $10^5$ active nodes and $10$ time stamps.
 The first \texttt{IntEvolvingGraph} in the sequence has about
-$1.0 \times 10^8$ static edges.
+$10^8$ static edges.
 We consecutively add new
@@
 for running Algorithm \ref{alg:bfs} in Julia. All experiments are conducted
-on a Linux system with Intel(R) Xeon(R) CPU E7-8850 @ $2.00$GHz.
+on a single core of a Linux system with an Intel(R) Xeon(R) E7-8850 CPU running
+at 2.00 GHz clock speed.
 The results show Algorithm \ref{alg:bfs} can be computed in linear time, which agrees

### 2016-03-04 paper.tex
Fix typo: ocirc -> odot

@@
 the causal edge set $E'$. In the next section, we show how to account for these
-causal edges by introducing a new matrix--vector product $\ocirc$.
+causal edges by introducing a new matrix--vector product $\odot$.
 
@@
 \eqref{eq:5}-\eqref{eq:6}, including both the ordinary matrix--vector product
-and the $\ocirc$ product. Furthermore, $(\bm A_3^T)^3 \bm b$ correctly counts the
+and the $\odot$ product. Furthermore, $(\bm A_3^T)^3 \bm b$ correctly counts the
 two allowed temporal paths from $(1, t_1)$ to $(3, t_3)$, and that the

### 2016-03-04 paper.tex
clarify causal edges in the experiment

@@
 Note that in this experiment, we do not have direct control
-over the full edge set $E$, only the static edge set $\tilde E$. However, the
-causal edge set $E'$ can grow at most by $O(\tilde E)$, so it suffices to demonstrate
+over the full edge set $E$, only the static edge set $\tilde E$. 
+When we add new static edges, new causal edges may be added as well
+(if the corresponding temporal nodes were not active before).
+However, the number of newly introduced causal edges for each active node 
+is bounded by the number of time stamps, so it suffices to demonstrate
 linear scaling in $|\tilde E|$.
@@
   \caption{Experimental run time of Algorithm \ref{alg:bfs}
-on a collection of random evolving graphs with $1.0\times 10^5$ active nodes and $10$ time
+on a collection of random evolving graphs with $10^5$ active nodes and $10$ time
 stamps, showing linear scaling in $|\tilde E|$. The horizontal axis shows the

### 2016-03-04 paper.tex
Describe Julia machine better
various minor fixes

@@
 In Section~\ref{sec:breadth-first-search}, we explain how the BFS algorithm can
-be applied to evolving graph to enumerate paths that traverse edges across time
+be applied to an evolving graph to enumerate paths that traverse edges across time
 and space. Section~\ref{sec:temporal-paths} provides an example showing that considering
@@
 
-The node set $V$ and edge set $E$ now define a directed static graph $G=(V,E)$
+The node set $V$ and edge set $E$ now define a static directed graph $G=(V,E)$
 that is in 1-1 correspondence with the evolving graph $G_n$. The node set $V$ of $G$ is in 1-1 correspondence with active nodes of $G_n$ while the edge set $E$ is in 1-1
@@
 
-We can then represent $G_n$ using the sequence of adjacency matrices
+We can then represent $G_n$ using a sequence of adjacency matrices
 $A_n = \langle A^{[1]}, A^{[2]}, \ldots, A^{n]}\rangle$. The example in
@@
 for running Algorithm \ref{alg:bfs} in Julia. All experiments are conducted
-on a single core of a Linux system with an Intel(R) Xeon(R) E7-8850 CPU running
-at 2.00 GHz clock speed.
+on a single core of a Linux system with 1TB of RAM and 80 cores of Intel(R)
+Xeon(R) E7-8850s running at 2.00 GHz clock speed.
 The results show Algorithm \ref{alg:bfs} can be computed in linear time, which agrees
